### 2013-02-18 document.tex
start article
Signed-off-by: loopasam <samuel.croset@gmail.com>

% Template for PLoS
% Version 1.0 January 2009
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{plos2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW

%% END MACROS SECTION

\begin{document}

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf{Title}
}
% Insert Author names, affiliations and corresponding author email.
\\
Author1$^{1}$, 
Author2$^{2}$, 
Author3$^{3,\ast}$
\\
\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
\\
\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
\\
$\ast$ E-mail: Corresponding author@institute.edu
\end{flushleft}

% Please keep the abstract between 250 and 300 words
\section*{Abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLoS ONE authors please skip this step. 
% Author Summary not valid for PLoS ONE submissions.   
\section*{Author Summary}

\section*{Introduction}

pouet pouet blablabalba \cite{Krotzsch2012} for

% Results and Discussion can be combined.
\section*{Results}

\subsection*{Subsection 1}

\subsection*{Subsection 2}

\section*{Discussion}

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

%\section*{References}
% The bibtex filename
\bibliography{document}

\section*{Figure Legends}
%\begin{figure}[!ht]
%\begin{center}
%%\includegraphics[width=4in]{figure_name.2.eps}
%\end{center}
%\caption{
%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
%should be left justified, as specified by the options to the caption 
%package.
%}
%\label{Figure_label}
%\end{figure}

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

\end{document}

### 2013-02-19 document.tex
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 {\Large
-\textbf{Title}
+\textbf{The Functional Therapeutic Chemical Classification System}
 }
@@
 \\
-Author1$^{1}$, 
-Author2$^{2}$, 
-Author3$^{3,\ast}$
+Samuel Croset$^{1,\ast}$, 
+John Overington$^{2}$, 
+Dietrich Rebholz-Schuhmann$^{3}$
 \\
-\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
+\bf{1} Samuel Croset Dept/Program/Center, Institution Name, City, State, Country
 \\
-\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
+\bf{2} John Overington Dept/Program/Center, Institution Name, City, State, Country
 \\
-\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
+\bf{3} Dietrich Rebholz-Schuhmann Dept/Program/Center, Institution Name, City, State, Country
 \\
-$\ast$ E-mail: Corresponding author@institute.edu
+$\ast$ E-mail: Corresponding croset@ebi.ac.uk
 \end{flushleft}
@@
 \section*{Abstract}
+
+The abstract of the paper should be succinct; it must not exceed 300 words. 
+Authors should mention the techniques used without going into methodological detail 
+and should summarize the most important results. While the abstract is conceptually divided into 
+three sections (Background, Methodology/Principal Findings, and Conclusions/Significance), please 
+do not apply these distinct headings to the abstract within the article file. Please do 
+not include any citations and avoid specialist abbreviations.
 
@@
 
+We ask that all authors of research articles include a 150–200 word non-technical summary 
+of the work as part of the manuscript to immediately follow the abstract. This text is subject to editorial 
+change, should be written in the first-person voice, and should be distinct from the scientific abstract. Aim to 
+highlight where your work fits within a broader context; present the significance or possible implications of your work 
+simply and objectively; and avoid the use of acronyms and complex terminology wherever possible. The goal is to make your 
+findings accessible to a wide audience that includes both scientists and non-scientists. Authors may benefit from consulting 
+with a science writer or press officer to ensure they effectively communicate their findings to a general audience. Examples 
+are available at:
+
 \section*{Introduction}
 
-pouet pouet blablabalba \cite{Krotzsch2012} for
+pouet \cite{Krotzsch2012} pouet
+
+The introduction should put the focus of the manuscript into a broader context. 
+As you compose the introduction, think of readers who are not experts in this field. 
+Include a brief review of the key literature. If there are relevant controversies or disagreements in 
+the field, they should be mentioned so that a non-expert reader can delve into these issues further. 
+The introduction should conclude with a brief statement of the overall aim of the experiments and a comment 
+about whether that aim was achieved.
 
@@
 \section*{Results}
+
+The results section should provide details of all of the experiments that 
+are required to support the conclusions of the paper. There is no specific word limit for 
+this section, but details of experiments that are peripheral to the main thrust of the article and that 
+detract from the focus of the article should not be included. The section may be divided into subsections, 
+each with a concise subheading. Large datasets, including raw data, should be submitted as supporting files; 
+these are published online alongside the accepted article. The results section should be written in the past tense.
 
@@
 
+The discussion should spell out the major conclusions of the work along with some 
+explanation or speculation on the significance of these conclusions. How do the 
+conclusions affect the existing assumptions and models in the field? How can future research 
+build on these observations? What are the key experiments that must be done? The discussion should 
+be concise and tightly argued. The results and discussion may be combined into one section, if desired.
+
 % You may title this section "Methods" or "Models". 
@@
 
+This section should provide enough detail for reproduction of the findings. Protocols 
+for new methods should be included, but well-established protocols may simply be referenced. While we do 
+encourage authors to submit all appendices, detailed protocols, or details of the algorithms for newer or less 
+well-established methods, please do so as Supporting Information files. These are not included in the typeset manuscript, 
+but are downloadable and fully searchable from the HTML version of the article.
+
 % Do NOT remove this, even if you are not including acknowledgments
 \section*{Acknowledgments}
+
+People who contributed to the work but do not fit the criteria for authors should be listed in the 
+Acknowledgments, along with their contributions. You must also ensure that anyone named in the Acknowledgments 
+agrees to being so named.
+
+Details of the funding sources that have supported the work should be confined to the funding statement 
+provided in the online submission system. Do not include them in the Acknowledgments.
 
@@
 
+The aim of the figure legend should be to describe the key messages of the figure, but 
+the figure should also be discussed in the text. An enlarged version of the figure and its full legend 
+will often be viewed in a separate window online, and it should be possible for a reader to understand the figure 
+without switching back and forth between this window and the relevant parts of the text. Each legend should have 
+a concise title of no more than 15 words that can stand alone, without the use of figure part labels. The overall 
+legend itself should be succinct, while still explaining all figure parts, symbols and abbreviations. Avoid lengthy descriptions of methods.
+
 \section*{Tables}
@@
 
+Tables should be included at the end of the manuscript file and cited sequentially in the text. 
+All tables should have a concise title. Footnotes can be used to explain abbreviations. 
+Citations should be indicated using the same style as outlined above. Tables should not occupy more than one 
+printed page; larger tables can be published as online supporting information. Tables must be cell-based; 
+do not use picture elements, text boxes, tabs, or returns in tables. Please ensure that all tables conform to our 
+Guidelines for Figure and Table Preparation when preparing them.
+
 \end{document}

### 2013-03-22 document.tex
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 
-pouet \cite{Krotzsch2012} pouet
-
-The introduction should put the focus of the manuscript into a broader context. 
+The introduction \cite{Krotzsch2012} should put the focus of the manuscript into a broader context. 
 As you compose the introduction, think of readers who are not experts in this field. 
@@
 about whether that aim was achieved.
+
+Functional Therapeutic Chemical Classification System (FTC) is a taxonomy (ontology) representing the mode of action of approved
+drugs in the human body.
+It can be pictured as a toolbox classifying the therapeutic agents based on their functionality in the human body.
+FTC categories are automatically built based on Gene Ontology terms.
+FTC categories are also equivalent to arbitrary definitions such as drug that perturbs some (protein that involved-in some (regulation of 
+biological process)).
+A knowledge base created by integrating the content of DrugBank and some GO annotations assists to automatically
+assign compounds into FTC categories based on the satisfaction of the definition.
+The FTC is evaluated against the Anatomical Therapeutic Chemical Classification System by assigning equivalences between classes.
+The knowledge base can be queried in order to retrieve implicit information about the involvement of approved drugs into biological
+processes.
+Drug repurposing hypotheses can be formulated out of the evaluation.
+The FTC can serve as starting point for analysis or projects requiring a structured vocabulary representing the mode of action.
+The source code of the FTC as well as the data are publicly available for free to anyone.
+
+The indication of a drug depends on the effects induced by the molecule when administrated in a biological system.
+For instance, a molecule such as the Ximelagatran is indicated as anticoagulant because
+it directly inhibits the catalytic activity of the thrombin, 
+enzyme itself involved in the blood coagulation.
+In order to speak of this therapeutic effect (anticoagulant), people sometimes refer to 
+the **mode of action** of an active compound. 
+Wikipedia gives us the following definition for it:
+> A mode of action describes a functional or anatomical change, at the 
+> cellular level, resulting from the exposure of a living organism to a substance.
+
+Characterising the mode of action of drugs is interesting for the following reasons:
+It is a central link between the molecular structure of a drug and its indication towards a disease. Drugs with similar
+mode of actions could be used to treat similar conditions.
+It abstracts away from the chemical structure and allows to reason over the effects of the compound in the body. The mode of action
+of a drug is a feature against which compounds can be classified.
+It can captures polypharmacology. Most of the drugs are not magic bullets
+and produce various side effects. A drug can have several mode of actions, representing the different processes the drug is perturbing directly
+or indirectly.
+A representation of the mode of action helps to better understand and analyse the behaviour of an agent in the human body.
+
+Although the **mode of action** is an old and useful notion in the drug discovery 
+community (see number of results for thisPubMed search, no resources were
+previously fully dedicated to it. The aim of the FTC is to fill this gap.
+
+The FTC can be seen as a toolbox designed to fix dysfunctioning biological systems.
+It features over 20'000 compartments (categories) inside which approved 
+drugs are automatically classified, based on facts from various databases.
+FTC categories represent the mode of action of compounds in the human system. 
+The resource can be used for various tasks where a structured vocabulary is needed, such as compound annotation, drug repurposing analysis and 
+anything else you can think of!
 

### 2013-08-15 document.tex
Signed-off-by: loopasam <samuel.croset@gmail.com>
first dump of text
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-02-18 document.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[10pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-
-%% END MACROS SECTION
+\documentclass{bioinfo}
+\copyrightyear{2005}
+\pubyear{2005}
 
 \begin{document}
+\firstpage{1}
 
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{Title}
+\title[short Title]{The Functional Therapeutic Chemical Classification System}
+\author[Croset \textit{et~al}]{Samuel Croset\,$^{1, \footnote{to whom correspondence should be addressed}}$, John Overington\,$^{1}$ and Dietrich Rebholz-Schuhmann\,$^{1}$
 }
-% Insert Author names, affiliations and corresponding author email.
-\\
-Author1$^{1}$, 
-Author2$^{2}$, 
-Author3$^{3,\ast}$
-\\
-\bf{1} Author1 Dept/Program/Center, Institution Name, City, State, Country
-\\
-\bf{2} Author2 Dept/Program/Center, Institution Name, City, State, Country
-\\
-\bf{3} Author3 Dept/Program/Center, Institution Name, City, State, Country
-\\
-$\ast$ E-mail: Corresponding author@institute.edu
-\end{flushleft}
+\address{$^{1}$European Molecular Biology Laboratory
+European Bioinformatics Institute (EMBL-EBI), Wellcome Trust Genome Campus, Hinxton, Cambridge CB10 1SD,
+United Kingdom\\}
 
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
+\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
 
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-\section*{Author Summary}
+\editor{Associate Editor: XXXXXXX}
 
-\section*{Introduction}
+\maketitle
 
-pouet pouet blablabalba \cite{Krotzsch2012} for
+\begin{abstract}
 
-% Results and Discussion can be combined.
-\section*{Results}
+\section{Motivation:}
+Drug repurposing is the discovery of new indications for compounds already approved and used in a clinical setting. 
+Recently, some computational approaches have been developed to unveil new opportunities in a systematic fashion, by taking 
+into consideration gene expression signatures or chemical features for instance. 
+We present here a novel method based on knowledge integration using semantic technologies. 
+In order to computationally generate repurposing hypotheses, we used the Web Ontology Language (OWL) to 
+mathematically define the meaning of over 20'000 terms formally describing various modes and mechanisms of action.
 
-\subsection*{Subsection 1}
+\section{Results:}
+Based on an integration of public data, we have automatically assigned over a thousand of approved drugs into 
+these mode of action categories. The resulting new resource is called the Functional Therapeutic Chemical Classification System (FTC) 
+and was further evaluated against the content of the traditional Anatomical Therapeutic Chemical Classification System (ATC). 
+We illustrate how the new classification can be used to generate drug repurposing hypotheses, using Alzheimer’s disease as a use-case.
 
-\subsection*{Subsection 2}
+\section{Availability:} A web application built on the top of the resource is freely available at {{https://www.ebi.ac.uk/chembl/ftc}}. The 
+source code of the project is available at {{https://github.com/loopasam/ftc}}.
 
-\section*{Discussion}
+\section{Supplementary Information:}
+Supplementary data are available at Bioinformatics online.
 
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis" 
-\section*{Materials and Methods}
+\section{Contact:} \href{croset@ebi.ac.uk}{croset@ebi.ac.uk}
+\end{abstract}
 
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgments}
+\section{Motivation}
 
-%\section*{References}
-% The bibtex filename
-\bibliography{document}
+Drug repurposing is the application of known drugs and compounds to new indications (i.e., new diseases) (cite wikipedia and better source). 
+When administered in a living organism, a compound can indeed play various roles and affect different biological processes. 
+Accurately identifying these different functions helps to predict the potential side-effects a drug could have and can also lead to 
+interesting repurposing opportunities (http://www.ncbi.nlm.nih.gov/pubmed/23340113). For instance, sildenafil was initially developed 
+against angina and has been repurposed towards erectile dysfunction during the clinical trials (cite). Approved compounds are 
+attractive because they have been extensively studied and have by definition already successfully passed clinical trials, where 
+most drugs fail because of safety or efficacy issues.
+There is an increasing number of approaches to predict repurposing opportunities using computational methods (cite). Most 
+methods operate on the profiles of physicochemical descriptors derived from molecular structures (cite). Other methods characterise 
+the drugs on more abstract levels, such as the gene expression signature (cite) or via the reported side-effects (cite). These approaches 
+have in common to look for similarities within existing drugs and forward similar compounds as repurposing hypotheses.
+A feature of particular interest to describe drugs is the mode of action (MoA). The MoA describes a functional or anatomical change, 
+at the cellular level, resulting from the exposure of a living organism to a substance (cite wikipedia or better source). For instance 
+terms such as transcriptional regulation agent or anticoagulant define MoAs and characterise the roles of a certain type of drugs. The 
+MoA abstracts over the relations between molecular functions, protein targets and drug activities; it is the central concept linking a 
+chemical structure to a set of biological effects. Intuitively, the indication of a drug logically depends on its MoA.
+Despite its widespread use in drug discovery, the MoA has not been used yet as a descriptor for repurposing analyses. One reason for this 
+might be the challenge of formally defining MoAs. Indeed, MoAs are terms or categories, it is not possible to represent 
+them straightforwardly with values and numbers like one can do for a 3D molecular structure or for a gene expression profile. 
+Nonetheless, the meaning of a concept can be formalised with controlled vocabularies and ontologies (cite); originating from description logics, such frameworks help to define the mathematical meaning of symbols and strings of characters.
+In an ontology or knowledge base, concepts are organised and linked following the logical type of relation they have between them.
+In the Gene Ontology (GO) for example (cite), biological processes and molecular functions are manually curated in this resource and the meaning of terms is specified by the relations linking two GO terms.
+MoA definitions are present in other classifications such as the Medical Subject Headings (cite) or the Chemical Entities of 
+Biological Interest (cite) for example. The Anatomical Therapeutic Chemical Classification System (ATC) also describes to some 
+extent the action of drugs at the anatomical level. All these resources are valuable for the community as a source of carefully 
+manually curated information. Moreover, the categories described in these classification systems are sometimes used to annotate 
+drugs: For instance the compound sildenafil has been manually annotated as vasodilator agent (CHEBI:35620 or MeSH:D27.505.954.411.918).
+The classifications mentioned previously are not specially designed for drug repurposing; they purposefully report only the 
+well-known and major MoAs of chemical compounds. The pharmacological spectrum of each drug is not necessarily well covered, yet 
+it would be the best way to predict new indications. In our context, an ideal knowledge base would feature the known MoAs of a 
+drug as well as some predicted ones to be tested in experiments. The MoA categories should derive and scale over primary molecular 
+evidences exposed in biomedical databases in an automated way.
+To address the lack of systematic MoA annotations, we have implemented the Functional Therapeutic Chemical 
+Classification System (FTC), presented here in this manuscript. The FTC is automatically built by leveraging the content 
+of various biomedical databases using description logics and automated reasoning. Over 20’000 new MoA categories are defined 
+in the resource and further populated with approved drugs using the Web Ontology Language (OWL) in combination with a reasoner. 
+Most of the drugs are present in multiple FTC categories, reflecting the various roles a compound can play inside a biological system. 
+The resource was evaluated against the ATC, traditional classification scheme introduced before. We present as well some preliminary 
+analyses over the data, by looking at the relation between the MoA and the indication of a compound using semantic similarity. 
+Finally, we illustrate also how the FTC can be used as a pharmacological toolbox to analyse drug repurposing for Alzheimer’s disease.
 
-\section*{Figure Legends}
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
+ Text Text Text Text. Figure~\ref{fig:01} shows that the above method  Text 
+   Text Text.  \citep{Bag01} wants to know about text follows
 
-\section*{Tables}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
+\section{Results}
 
+The knowledge base behind the FTC is built by integrating information coming from various sources. 
+The GO terms serve as template to create the FTC categories describing the MoAs; DrugBank provides the 
+known links between drugs and their protein targets. Uniprot maps targets to their respective GO annotations. 
+Drugs are further assigned into mode of action categories based of the OWL constructs and axioms defined in the 
+FTC (see Availability and Implementation). A reasoner, a program capable of understanding such axioms, performed this task. 
+The process to build the FTC is summarised in Figure 0. It takes around 5 seconds on a modern 
+laptop (4 cores processor with 4 Gb of memory) to classify the knowledge base using the ELK reasoner (cite). 
+Other OWL reasoners have been tried (e.g, Hermit (cite), Pellet (cite), etc...) but none of them were suitable for this purpose; 
+the classification time was always superior to a few minutes (data not shown). The FTC has a taxonomic structure as illustrated 
+on Figure 1, which arises when the reasoner classifies the knowledge base. Categories can have multiple parents and multiple children. 
+The reader can browse and access the content of the FTC online at {{https://www.ebi.ac.uk/chembl/ftc}}.
+
+Figure 1: Parent categories to the FTC class Pro-fibrinolysis agent (FTC\_P0042730). The classification is a direct 
+acyclic graph where categories are describing increasingly specific concepts. Arrows entails subclass relationships 
+between the terms (is a relation).
+
+In total there are 1280 approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets. 
+The FTC introduces 23'353 new categories describing the mode and mechanism of action of therapeutic compounds. 4289 of these 
+categories are related to biological processes and 19'064 to the molecular functions. A summary of the metrics behind the latest 
+build is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}. Out of all FTC classes, 1432 
+categories (\textgreater 6\%) directly contain at least one approved drug. This number increases up to 2532 (\textgreater 11\%) 
+when direct and 
+indirect drugs are considered. FTC categories not containing drugs (e.g, FTC\_A0001771 - Anti-immunological synapse 
+formation agent) represent modes and mechanisms of action for which no approved compounds exist yet or that have not been identified 
+as such in the FTC.
+
+\subsection{Evaluation}
+The content of the FTC has been evaluated against the information present in the ATC. The motivation behind 
+this step is to have an estimation of how the automatically-built classification (FTC) compares to a manually curated 
+one (ATC). In the context of this work, we have considered the ATC as a gold standard against which the FTC 
+has been assessed. In practice, these two resources have different goals; therefore the evaluation should be 
+interpreted accordingly. The full methodology behind the evaluation is described in the section “Evaluation methodology”. 
+Briefly, 68 classes present in the FTC have an equivalent meaning to a combination of classes present in the ATC. 
+This semantic equivalence was manually asserted and called evaluation point (see “Evaluation points” from Availability and Implementation). 
+The set of drugs present in both classes for each evaluation points were then compared and some metrics are derived 
+from this comparison. For example, the FTC category Anti-hydrogen:potassium-exchanging ATPase activity agent (FTC\_A0008900) has 
+been manually asserted as equivalent to the ATC category Proton pump inhibitors (A02BC). A summary of this 
+evaluation point is furthermore available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/FTC\_A0008900}}.
+Out of the 1280 DrugBank compounds present in the FTC, 1134 are also present in the ATC, therefore only those were 
+considered. The evaluation points cover a total of 471 DrugBank compounds. Out of these, 277 compounds are true positives, 
+meaning that given an evaluation point, they are present in both FTC and ATC categories. The proton pump inhibitor 
+evaluation point is such an example where all the drugs (omeprazole, esomeprazole, pantoprazole, lansoprazole, rabeprazole) are 
+present both in the FTC category and in the corresponding ATC class. The total number of compounds presents in an ATC class but 
+not in the corresponding FTC category (false negatives) is 36. Finally, 279 compounds are present in a FTC class but not in the 
+corresponding ATC class; such agents are the false positives. From these values, it is possible to derive a 
+recall of 88\%; this percentage indicates that the automatic build of the FTC covers a good portion of the content 
+already present in the ATC. The precision of 50\% shows that the FTC contains for a given MoA many more drugs than the 
+equivalent ATC categories.
+
+\subsection{Exploration}
+The FTC was designed to assist drug repurposing analyses by explicitly representing the poly-pharmacology of 
+approved drugs. In this section we exemplify how the resource can be used to perform different types of analysis.
+Polypharmacology spectrum
+With more information on a drug’s molecular targets and their physiological roles, the more opportunities 
+exist to re-orient a drug into doing something new. The therapeutic agents described in the FTC can have several MoAs, 
+representing the intrinsic polypharmacology of the approved compounds. Figure 2 illustrates the poly-pharmacology 
+spectrum by showing the distribution of number of MoAs per compound. This number ranges from one to more 
+than eighty for certain drugs, when indirect parents are considered. Not all the MoAs are biologically relevant, 
+some FTC categories are particularly abstract (e.g, Anti-biological process agent) yet they represent discrete 
+categories to which the drug belongs with an explicit and clear meaning. These discrete MoAs are a good starting 
+point to understand what a compound can do when administered in a human system. The human readable definitions 
+associated to each FTC category furthermore help the user to choose the adequate MoA for the desired outcome.
+
+Figure 2: Distribution of the direct and indirect number of MoAs per drug. Means are indicated with a solid line. 
+When considering indirect categories, the distribution range is wider.
+Drugs with similar mode of actions have similar indications
+The list of MoAs attributed to a drug can be exploited as a descriptor for the therapeutic agent. 
+The tree structure of the FTC can be used to derive some similarity metrics over the MoAs. The underlying 
+heuristic is to assume that the closer two entities are in a taxonomy or ontology, the more similar they are. 
+We used a straightforward approach derived from the Jaccard index (see method for more details) in order to compare 
+approved drugs based on the similarity of their MoAs. For instance, the similarity between two compounds present in 
+the same FTC category is 1 (maximum). The similarity between an anti-blood coagulant and pro-blood coagulant is 0.29, 
+reflecting the fact that such compounds are dissimilar in regards to the outcome of their biological effect. As the MoA is 
+intuitively expected to be the central concept leading to the indication of the drug, we expected that on average, drugs with 
+similar MoAs would be indicated towards similar therapeutic areas. The heat map presented in Figure 3 shows a pairwise 
+comparison of all the drugs of the FTC based on their relative MoA similarity. The compounds are further grouped by 
+therapeutic indications as defined by the ATC.
+
+Figure 3: Pairwise comparison of mode of action similarities. Only the first level of 
+the ATC is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
+The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black). Drugs are grouped according 
+to their first ATC level (colours on the side). For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
+which appears as “B” (dark orange) on the plot. The average similarity of drugs present in the same therapeutic category 
+is significantly higher on average when separately compared to all other indications.
+
+The heat map displayed in Figure 3 reveals some square patches around the central diagonal. 
+The overall similarity appears higher when compounds from the same ATC group are considered. 
+A significance analysis (cf details covered in sections “Semantic similarity” 
+and “Mode of action similarity against indication” from Availability and Implementation) 
+revealed that the average MoA similarity of compounds belonging to the same ATC category is 
+significantly higher than when compounds belonging to different categories are compared. Indeed, for 
+each category, the p-value was inferior to 0.05 based on 20’000 random permutations over the similarity values. 
+This result supports the idea that drugs with similar MoAs have similar indications. Note that the mean of the 
+similarity values was considered for the statistical analysis; some outliers are also present in the map, which 
+can be interpreted as repurposing hypotheses. These outliers have indeed similar MoAs, yet they belong to totally 
+different therapeutic areas and are used for different purposes according to the ATC. We are currently further analysing 
+such cases in a systematic fashion. Supplementary Figure S1A present similar association behavior when two levels of the ATC are 
+considered (no statistical significance performed). Supplementary Figure S1B re-uses the same data as Figure 1 (one ATC level) 
+but with a clustering function apply to it (hierarchical clustering - manhattan distance) in order to reveal 
+functional clusters of drugs. Finally, Figure S1C shows the distribution when compounds are sorted based on 
+their identifiers; no patterns are identifiable in this case. Taken together, these results emphasize that the MoAs as 
+defined in the FTC are indeed on average associated with the therapeutic indication of a drug.
+
+Supplementary Figure S1: Pairwise comparison of mode of action similarities. The compounds are grouped 
+by ATC code (2 levels - legend not shown) on panel A. A hierarchical 
+clustering step (based on manhattan distance) was applied on panel B. The 
+dendrogram reflects the original structure of the FTC. Panel C shows the compounds 
+sorted by DrugBank identifiers only (≈ random). The colours displayed on panel B and C 
+correspond to one level ATC categories and are indexed in the legend panel.
+Drug repurposing hypotheses for Alzheimer’s disease
+In this section we provide examples how actual drug repurposing hypotheses can be derived from the FTC. 
+The approach presented here makes use of the FTC categories as analogs to compartments of a toolbox helping to 
+find drugs to treat Alzheimer’s disease. Five FTC categories containing drugs are directly related to the biological 
+processes of the neurodegenerative condition: Anti-amyloid precursor protein biosynthetic process agent (FTC\_A0042983), 
+Anti-tau-protein kinase activity agent (FTC\_A0050321), Anti-tau protein binding agent (FTC\_A0048156), Anti-beta-amyloid 
+binding agent (FTC\_A0001540) and Pro-beta-amyloid binding agent (FTC\_P0001540). We have then considered the drugs present 
+inside each of these classes as potential candidates. Figure 4 shows these drugs, which have been further manually grouped 
+based on the overall similarity of their actions (letters on Figure 4).
+
+Figure 4: FTC categories (letters A, F, H, I and J) describing some of the modes and mechanisms of 
+action that could impact Alzheimer’s disease. Drugs classified in these FTC categories are listed and further 
+manually grouped based on their mode of action similarities (subgroups B, C, D, E and G).
+
+The subgroups B, C and D are inhibitors of the cholinergic system and some of them, such as galantamine (DB00674), are already 
+investigated to treat Alzheimer’s disease and other related dementias (pubchem). This class of agent is in line with the 
+cholinergic hypothesis (cite PMID:10071091), stating that Alzheimer’s disease could be caused by dysfunctions in the processing of acetylcholine.
+The subgroup E is exclusively composed of barbiturates (central nervous system depressants).
+The presence of this pharmacological class of compounds as an Alzheimer treatment is more surprising, as very little literature reports on it.
+Further investigations reveal that the neuronal acetylcholine receptor subunit alpha-7 (P36544), a common off-target of barbiturates, binds beta-amyloids with high affinity.
+As beta-amyloids are themselves strongly involved in the pathology (cite http://europepmc.org/abstract/MED/10681545), barbiturates could affect the state of the condition, similarly to cholinergic inhibitors.
+The group F contains four compounds; nicotine and varenicline have been grouped because of similar pharmacology (group G). 
+Nicotine has been shown to improve some of the symptoms of Alzheimer’s disease (PMID:1410164), it is therefore expected to 
+find this molecule in the predictions. Varenicline possess a pharmacology related to that of nicotine, which would explain 
+the presence of the drug in this category. The two remaining drugs of group F are respectively pralidoxime and dipivefrin; 
+little information is available regarding their potential action against the disease, yet these compounds seem linked to the 
+cholinergic hypothesis and could be considered for experimental testing.
+The groups H and I contain respectively one molecule each. These compounds have been classified as agents perturbing some of 
+the physiological function of the Tau protein, key actor in Alzheimer's disease (cite). Vorinostat (group H) is currently 
+indicated for the treatment of cutaneous manifestations in patients with T-cell lymphoma (drugbank), yet a study has 
+shown in vivo (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
+deficit (cite http://www.nature.com/npp/journal/v35/n4/full/npp2009197a.html). The presence of lithium (group I) was 
+unexpected and suspected of being a false positive at first glance. However a recent study demonstrated a long-term 
+protective effect for the ion in regards to Alzheimer’s disease (PMID:22500970 and 21525519). The last group (J) contains 
+ezetimibe and hesperetin. These two compounds are primarily used as cholesterol lowering agents (statins). As cholesterol 
+metabolism in the brain appears to be related to dementia, statins are believed to prevent or improve the symptoms of the 
+patients. Although early studies (PMID:17877925) have failed to clearly show a beneficial effect, the investigation is still open.
+From the examples briefly presented above, reported and confirmed by the literature, the FTC appears to be suitable to 
+identify real repurposing hypotheses tailored to a disease. The MoAs help to retrieve the compounds which might impact 
+the treatment of a given set of symptoms.
+
+\section{Discussion}
+
+The FTC is a novel classification for approved drugs, which can be used as a starting point to 
+generate drug repurposing hypotheses. We have guided the reader through the various ways the resource 
+can be explored. No experimental validations have been presented in this manuscript, yet references to 
+supporting literature was provided in order to legitimate the claims. The main goal of this document is to 
+introduce the resource to the community; deeper analyses are being performed over the classification at the time of writing.
+
+\subsection{Biological assumptions}
+An asset of the FTC is its ability to handle efficiently categorical data: Classes and relationships 
+are accurately defined, in order to classify compounds based on the semantics of their relations. The properties 
+linking drugs to their respective protein targets (positive and negative perturbations) are however simplistic. At the 
+time being, no consideration is given regarding the binding strength between the drug and the proteins, yet it is a key 
+factor to derive potent and specific activities in the human body. This is also the case for other types of numerical data, 
+such as the dosage; the FTC can predict a role for a drug, yet it cannot provide any information about the concentration or 
+the administration route necessary to obtain the potential effects. The current relations between targets and their involvement 
+in biological processes are also not a fully accurate representation of the biological phenomenon. In a cell, specific domains 
+of the protein could mediate different functions. Only one of such activity types can sometimes be inhibited 
+by a drug (cite felix paper), yet we are assuming in the FTC that as long as a drug affects a protein, 
+it can therefore alter all its known functions.
+These limitations come from the semantics behind the axioms structuring the classification, themselves based on the 
+information available from the databases (e.g. GOA, DrugBank, etc...). Despite entailing not entirely accurately the 
+biochemical reality, the axioms help to generate a larger number of hypotheses, the primary goal of the FTC. The dosage 
+issue is partially addressed by the regulator pattern (see section in Availability and Implementation): It should be 
+easier to experimentally adjust the concentration of the compounds classified as pro or anti biological process agents 
+in order to modulate a physiological effect.
+The predictions generated by the FTC depend on the resolution of the curated information released by 
+the original data providers. Erroneous or missing information will lead to misclassification by the reasoner. 
+Some expected outcomes are also missing from the predictions; sildenafil for instance was expected to be classified 
+as pro-penile erection agent (FTC\_A0043084), yet the lack of appropriate GO annotation prevents it. After discussion 
+with the GOA curation team, a manual annotation can only be asserted based on published experimental results. No document 
+was found to support the involvement of the cGMP-specific 3',5'-cyclic phosphodiesterase (O76074 - sildenafil’s main target) 
+in the negative regulation of penile erection (GO:0060407), therefore no annotation can be made. Further work could be done 
+in this direction, by trying to automatically infer more annotations or by using the electronically generated ones, in order 
+to generate broader yet potentially less plausible repurposing hypotheses.
+
+\subsection{Interpreting the evaluation}
+The high recall value (88\%) supports the idea behind the automated build of the FTC: The data from different 
+repositories funded and curated in parallel, can be integrated to automatically create a new resource. This new resource (FTC) 
+contains most of the known information present in an external gold standard (ATC) and relies on description logics to 
+leverage the native information. Here we have assessed the content of the FTC against the ATC, knowing that these two 
+taxonomies have diverging goals. During the evaluation, equivalences have been manually asserted between classes, 
+which are assumed to have fairly similar meaning and containing similar sets of compounds. These manual assertions 
+are however a weakness, as they are themselves not evaluated (free parameter). The presence or absence of a link was 
+determined only by one curator and any mistake can influence consequently the recall and precision values. In this regards, 
+the evaluation should be considered more as a safety control rather than a formal assessment of a predictive method.
+The false positives derived from the evaluation can also be considered as drug repurposing hypotheses: These drugs 
+can indeed be interpreted as suitable for the ATC category, yet not indexed as such. However, these predictions 
+should be interpreted with caution, as it is currently impossible to distinguish a false positive from a 
+reprofiling opportunity. These considerations do not interfere with the repurposing predictions generated 
+based on semantic similarities or discrete categories as presented in the Exploration section. Finally, 
+note that the ATC/FTC equivalences are open and editable online, any modification will be automatically 
+incorporated in the next release of the resource. It is also possible to evaluate the FTC against a different 
+taxonomy, like the Medical Subject Headings for example, which can be subject to future work.
+
+\begin{figure}[!tpb]%figure1
+%\centerline{\includegraphics{fig01.eps}}
+\caption{Caption, caption.}\label{fig:01}
+\end{figure}
+
+\begin{figure}[!tpb]%figure2
+%\centerline{\includegraphics{fig02.eps}}
+\caption{Caption, caption.}\label{fig:02}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%
+%     please remove the " % " symbol from \centerline{\includegraphics{fig01.eps}}
+%     as it may ignore the figures.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Conclusion}
+
+This manuscript introduces the FTC, a free, open and extensible resource, which can assist drug repurposing 
+initiatives or enhance computational studies making use of the concept of mode of action. The construction of 
+the classification relies on description logics as a means to define the mode and mechanism of action of approved 
+drugs. We shown the validity of the approach by comparing the content of the FTC to another well implemented taxonomy, 
+the ATC. We further illustrate the tight relationship between the MoA and the indication of a drug and shown how the 
+resource can serve as a starting point to formulate hypotheses for Alzheimer's disease. The work leverages the 
+semantics of distinct databases, working in parallel on different thematics. The platform will be further used 
+to generate predictions in a systematic fashion, which can then be experimentally tested in the laboratory for validation.
+
+\section*{Acknowledgement}
+Text Text Text Text Text Text  Text Text.  \citealp{Boffelli03} might want to know about  text text text text
+
+\paragraph{Funding\textcolon} Text Text Text Text Text Text  Text Text.
+
+%\bibliographystyle{natbib}
+%\bibliographystyle{achemnat}
+%\bibliographystyle{plainnat}
+%\bibliographystyle{abbrv}
+%\bibliographystyle{bioinformatics}
+%
+%\bibliographystyle{plain}
+%
+%\bibliography{Document}
+
+\begin{thebibliography}{}
+\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03} Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.
+
+\bibitem[Bag {\it et~al}., 2001]{Bag01} Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.
+
+\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
+Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
+in nigral dopaminergic neurnol cell: correlation with the known
+pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
+Res.}, \textbf{110}(Suppl. 1), 76--84.
+
+\bibitem[Lehmann, 1986]{Leh86}
+Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.
+
+\bibitem[Crenshaw and Jones, 2003]{Cre03}
+Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
+cancer management: one tumor, one chip. \textit{Bioinformatics},
+doi:10.1093/bioinformatics/btn000.
+
+\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
+Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
+(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
+???--???.
+
+\bibitem[Bardet, 1920]{Bar20}
+Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
+polydactylie et retinite pigmentaire (contribution a l'etude des
+formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
+institution, Paris, France.
+
+\end{thebibliography}
 \end{document}

### 2013-08-15 document.tex
figures inserted
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 FTC (see Availability and Implementation). A reasoner, a program capable of understanding such axioms, performed this task. 
-The process to build the FTC is summarised in Figure 0. It takes around 5 seconds on a modern 
+The process to build the FTC is summarised in Figure 0.
+
+\begin{figure*}[!tpb]%figure1
+\centerline{\includegraphics{fig1.png}}
+\caption{Parent categories to the FTC class Pro-fibrinolysis agent (FTC\_P0042730). The classification is a direct 
+acyclic graph where categories are describing increasingly specific concepts. Arrows entails subclass relationships 
+between the terms (is a relation).}\label{fig:01}
+\end{figure*}
+
+It takes around 5 seconds on a modern 
 laptop (4 cores processor with 4 Gb of memory) to classify the knowledge base using the ELK reasoner (cite). 
@@
 
-Figure 1: Parent categories to the FTC class Pro-fibrinolysis agent (FTC\_P0042730). The classification is a direct 
+\begin{figure}[!tpb]%figure2
+\centerline{\includegraphics{fig2.png}}
+\caption{Parent categories to the FTC class Pro-fibrinolysis agent (FTC\_P0042730). The classification is a direct 
 acyclic graph where categories are describing increasingly specific concepts. Arrows entails subclass relationships 
-between the terms (is a relation).
+between the terms (is a relation).}\label{fig:02}
+\end{figure}
 
@@
 approved drugs. In this section we exemplify how the resource can be used to perform different types of analysis.
-Polypharmacology spectrum
+
+\subsubsection{Polypharmacology spectrum}
 With more information on a drug’s molecular targets and their physiological roles, the more opportunities 
@@
 
-Figure 2: Distribution of the direct and indirect number of MoAs per drug. Means are indicated with a solid line. 
-When considering indirect categories, the distribution range is wider.
-Drugs with similar mode of actions have similar indications
+\begin{figure*}[!tpb]%figure3
+\centerline{\includegraphics{fig3.png}}
+\caption{Distribution of the direct and indirect number of MoAs per drug. 
+Means are indicated with a solid line. When considering indirect categories, the distribution range is wider.}\label{fig:03}
+\end{figure*}
+
+\subsubsection{Drugs with similar mode of actions have similar indications}
 The list of MoAs attributed to a drug can be exploited as a descriptor for the therapeutic agent. 
@@
 
-Figure 3: Pairwise comparison of mode of action similarities. Only the first level of 
+\begin{figure*}[!tpb]%figure4
+\centerline{\includegraphics{fig4.png}}
+\caption{Pairwise comparison of mode of action similarities. Only the first level of 
 the ATC is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
@@
 which appears as “B” (dark orange) on the plot. The average similarity of drugs present in the same therapeutic category 
-is significantly higher on average when separately compared to all other indications.
+is significantly higher on average when separately compared to all other indications.}\label{fig:04}
+\end{figure*}
 
@@
 significantly higher than when compounds belonging to different categories are compared. Indeed, for 
-each category, the p-value was inferior to 0.05 based on 20’000 random permutations over the similarity values. 
+each category, the p-value was inferior to 0.05 based on 20'000 random permutations over the similarity values. 
 This result supports the idea that drugs with similar MoAs have similar indications. Note that the mean of the 
@@
 
-Supplementary Figure S1: Pairwise comparison of mode of action similarities. The compounds are grouped 
-by ATC code (2 levels - legend not shown) on panel A. A hierarchical 
-clustering step (based on manhattan distance) was applied on panel B. The 
-dendrogram reflects the original structure of the FTC. Panel C shows the compounds 
-sorted by DrugBank identifiers only (≈ random). The colours displayed on panel B and C 
-correspond to one level ATC categories and are indexed in the legend panel.
-Drug repurposing hypotheses for Alzheimer’s disease
+\subsection{Drug repurposing hypotheses for Alzheimer’s disease}
 In this section we provide examples how actual drug repurposing hypotheses can be derived from the FTC. 
@@
 processes of the neurodegenerative condition: Anti-amyloid precursor protein biosynthetic process agent (FTC\_A0042983), 
-Anti-tau-protein kinase activity agent (FTC\_A0050321), Anti-tau protein binding agent (FTC\_A0048156), Anti-beta-amyloid 
-binding agent (FTC\_A0001540) and Pro-beta-amyloid binding agent (FTC\_P0001540). We have then considered the drugs present 
+Anti-tau protein kinase activity agent (FTC\_A0050321), Anti-tau protein binding agent (FTC\_A0048156), Anti-beta amyloid 
+binding agent (FTC\_A0001540) and Pro-beta amyloid binding agent (FTC\_P0001540). We have then considered the drugs present 
 inside each of these classes as potential candidates. Figure 4 shows these drugs, which have been further manually grouped 
@@
 
-Figure 4: FTC categories (letters A, F, H, I and J) describing some of the modes and mechanisms of 
+\begin{figure}[!tpb]%figure5
+\centerline{\includegraphics{fig5.png}}
+\caption{FTC categories (letters A, F, H, I and J) describing some of the modes and mechanisms of 
 action that could impact Alzheimer’s disease. Drugs classified in these FTC categories are listed and further 
-manually grouped based on their mode of action similarities (subgroups B, C, D, E and G).
+manually grouped based on their mode of action similarities (subgroups B, C, D, E and G).}\label{fig:05}
+\end{figure}
 
@@
 
-\begin{figure}[!tpb]%figure1
-%\centerline{\includegraphics{fig01.eps}}
-\caption{Caption, caption.}\label{fig:01}
-\end{figure}
-
-\begin{figure}[!tpb]%figure2
-%\centerline{\includegraphics{fig02.eps}}
-\caption{Caption, caption.}\label{fig:02}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%     please remove the " % " symbol from \centerline{\includegraphics{fig01.eps}}
-%     as it may ignore the figures.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 \section{Conclusion}
-
 This manuscript introduces the FTC, a free, open and extensible resource, which can assist drug repurposing 
@@
 to generate predictions in a systematic fashion, which can then be experimentally tested in the laboratory for validation.
+
+\section{Availability and Implementation}
+The code behind the creation of the resource is entirely open and available 
+at {{https://github.com/loopasam/ftc}}. The web application built on the top of the FTC can be 
+find at {{https://www.ebi.ac.uk/chembl/ftc}} and the documentation can be accessed at {{https://github.com/loopasam/ftc/wiki}}. The full
+description of the methodology used to generate the classification is available as supplementary material on Bioinformatics website.
 

### 2013-08-15 document.tex
supp material on the way
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-03-22 document.tex
@@
-% Template for PLoS
-% Version 1.0 January 2009
-%
-% To compile to pdf, run:
-% latex plos.template
-% bibtex plos.template
-% latex plos.template
-% latex plos.template
-% dvipdf plos.template
-
-\documentclass[10pt]{article}
-
-% amsmath package, useful for mathematical formulas
-\usepackage{amsmath}
-% amssymb package, useful for mathematical symbols
-\usepackage{amssymb}
-
-% graphicx package, useful for including eps and pdf graphics
-% include graphics with the command \includegraphics
-\usepackage{graphicx}
-
-% cite package, to clean up citations in the main text. Do not remove.
-\usepackage{cite}
-
-\usepackage{color}
-
-% Use doublespacing - comment out for single spacing
-%\usepackage{setspace} 
-%\doublespacing
-
-% Text layout
-\topmargin 0.0cm
-\oddsidemargin 0.5cm
-\evensidemargin 0.5cm
-\textwidth 16cm 
-\textheight 21cm
-
-% Bold the 'Figure #' in the caption and separate it with a period
-% Captions will be left justified
-\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}
-
-% Use the PLoS provided bibtex style
-\bibliographystyle{plos2009}
-
-% Remove brackets from numbering in List of References
-\makeatletter
-\renewcommand{\@biblabel}[1]{\quad#1.}
-\makeatother
-
-% Leave date blank
-\date{}
-
-\pagestyle{myheadings}
-%% ** EDIT HERE **
-
-%% ** EDIT HERE **
-%% PLEASE INCLUDE ALL MACROS BELOW
-
-%% END MACROS SECTION
+\documentclass{bioinfo}
+\copyrightyear{2005}
+\pubyear{2005}
 
 \begin{document}
+\firstpage{1}
 
-% Title must be 150 characters or less
-\begin{flushleft}
-{\Large
-\textbf{The Functional Therapeutic Chemical Classification System}
-}
-% Insert Author names, affiliations and corresponding author email.
-\\
-Samuel Croset$^{1,\ast}$, 
-John Overington$^{2}$, 
-Dietrich Rebholz-Schuhmann$^{3}$
-\\
-\bf{1} Samuel Croset Dept/Program/Center, Institution Name, City, State, Country
-\\
-\bf{2} John Overington Dept/Program/Center, Institution Name, City, State, Country
-\\
-\bf{3} Dietrich Rebholz-Schuhmann Dept/Program/Center, Institution Name, City, State, Country
-\\
-$\ast$ E-mail: Corresponding croset@ebi.ac.uk
-\end{flushleft}
+\title[Supplementary information - The Functional Therapeutic Chemical Classification System]{Supplementary information - 
+The Functional Therapeutic Chemical Classification System}
 
-% Please keep the abstract between 250 and 300 words
-\section*{Abstract}
+\section*{Availability and Implementation}
 
-The abstract of the paper should be succinct; it must not exceed 300 words. 
-Authors should mention the techniques used without going into methodological detail 
-and should summarize the most important results. While the abstract is conceptually divided into 
-three sections (Background, Methodology/Principal Findings, and Conclusions/Significance), please 
-do not apply these distinct headings to the abstract within the article file. Please do 
-not include any citations and avoid specialist abbreviations.
+This section describes the building mechanism behind the FTC. 
+First, a list of categories describing the mode and mechanism of action of drugs is defined. 
+Then in a second step the newly created categories are automatically populated with approved compounds. 
+Finally, the FTC is evaluated and repurposing hypotheses can be generated.
 
-% Please keep the Author Summary between 150 and 200 words
-% Use first person. PLoS ONE authors please skip this step. 
-% Author Summary not valid for PLoS ONE submissions.   
-\section*{Author Summary}
+\section{Source code}
+The code behind the creation of the resource is entirely open and available 
+at {{https://github.com/loopasam/ftc}}. The web application built on the top of the 
+FTC can be find at {{https://www.ebi.ac.uk/chembl/ftc}} and the documentation can be 
+accessed at {{https://github.com/loopasam/ftc/wiki}}. The reader should be familiar with 
+description logics and the Web Ontology Language (OWL) to fully understand the construction of the 
+knowledge base. An introduction to description logics from the perspective of the biomedical scientist is 
+available on the wiki at {{https://github.com/loopasam/ftc/wiki/Description-Logics}}. The FTC implementation 
+relies mostly on Brain (cite) and the web application builds on the top of the Play! framework (cite). 
+Classification tasks use the ELK reasoner (cite). The computer hosting the web application has 8Go of memory 
+with 4 processors, this architecture allows fast parallel reasoning, thanks to ELK’s design. More functionalities 
+will be added to the web application following user requirements (lean implementation).
 
-We ask that all authors of research articles include a 150–200 word non-technical summary 
-of the work as part of the manuscript to immediately follow the abstract. This text is subject to editorial 
-change, should be written in the first-person voice, and should be distinct from the scientific abstract. Aim to 
-highlight where your work fits within a broader context; present the significance or possible implications of your work 
-simply and objectively; and avoid the use of acronyms and complex terminology wherever possible. The goal is to make your 
-findings accessible to a wide audience that includes both scientists and non-scientists. Authors may benefit from consulting 
-with a science writer or press officer to ensure they effectively communicate their findings to a general audience. Examples 
-are available at:
+\section{FTC categories creation}
+The mode of action categories present in the FTC are defined based on the terms coming from the 
+Gene Ontology (GO). Both the molecular function and biological process branches are used for this purpose, yet handled differently.
 
-\section*{Introduction}
+\subsection{Categories related to biological processes}
+All the biological processes featured by the GO are looked-up one by one. All the 
+time a process is linked to another process (X) via a positive or negative regulation link, two FTC 
+classes are created: Anti-X agent and Pro-X agent. For instance the GO term positive regulation of blood 
+coagulation is linked to the term blood coagulation via a positively regulates relation, therefore two FTC 
+categories Anti-blood coagulation agent and Pro-blood coagulation agent are created. The identifiers of the 
+new FTC classes are also derived from the GO term used to create the class pattern. The GO numeric identifier 
+is re-used and the letter A or P is appended before to emphasize the anti or pro pattern. From the example presented 
+previously, the FTC class Anti-blood coagulation has FTC\_A0007596 as identifier, because the GO term blood coagulation is 
+referenced by GO:0007596. Following the same logic, FTC\_P0007596 is the identifier of the class Pro-blood coagulation. The 
+design choice for identifiers and labels allows the FTC to fully rely on the high quality work provided by the GO curation 
+team and scale over it.
 
-The introduction \cite{Krotzsch2012} should put the focus of the manuscript into a broader context. 
-As you compose the introduction, think of readers who are not experts in this field. 
-Include a brief review of the key literature. If there are relevant controversies or disagreements in 
-the field, they should be mentioned so that a non-expert reader can delve into these issues further. 
-The introduction should conclude with a brief statement of the overall aim of the experiments and a comment 
-about whether that aim was achieved.
+\subsection{Categories related to molecular functions}
+The mode and mechanism of actions related to molecular functions are created in the following manner: All the time 
+a molecular function (Y) is encountered then two FTC categories are created, as for the processes: Anti-Y agent and 
+Pro-Y agent. The identifiers are assigned the same way as described before. For instance, out of the GO term androgen 
+receptor activity (GO:0004882) two FTC classes are derived: Pro-androgen receptor activity agent (FTC\_P0004882) and 
+Anti-androgen receptor activity agent (FTC\_A0004882).
 
-Functional Therapeutic Chemical Classification System (FTC) is a taxonomy (ontology) representing the mode of action of approved
-drugs in the human body.
-It can be pictured as a toolbox classifying the therapeutic agents based on their functionality in the human body.
-FTC categories are automatically built based on Gene Ontology terms.
-FTC categories are also equivalent to arbitrary definitions such as drug that perturbs some (protein that involved-in some (regulation of 
-biological process)).
-A knowledge base created by integrating the content of DrugBank and some GO annotations assists to automatically
-assign compounds into FTC categories based on the satisfaction of the definition.
-The FTC is evaluated against the Anatomical Therapeutic Chemical Classification System by assigning equivalences between classes.
-The knowledge base can be queried in order to retrieve implicit information about the involvement of approved drugs into biological
-processes.
-Drug repurposing hypotheses can be formulated out of the evaluation.
-The FTC can serve as starting point for analysis or projects requiring a structured vocabulary representing the mode of action.
-The source code of the FTC as well as the data are publicly available for free to anyone.
+\section{Equivalent definitions}
+FTC classes are generated as presented in the previous section. Up to this point, these categories 
+are only individual tokens with a human readable label as well as an identifier. The next step is going to 
+assign equivalent definitions to each FTC class. A reasoner can understand such definitions and will automatically 
+classify the knowledge base accordingly; drugs will then be assigned to FTC categories and the taxonomic structure 
+arises after the reasoning step. Equivalent definitions are written as OWL class expressions using the entities of 
+the knowledge base. The core classes and properties are presented in the supplementary material, this section assumes 
+that the reader is already familiar with the FTC specifications (summarised at {{https://github.com/loopasam/ftc/wiki/Knowledge-Base}}). 
+There are two types of equivalences: The first one captures perturbation of regulatory 
+biological processes (so called regulatory patterns) and the second one handles the perturbed functions (functional patterns).
 
-The indication of a drug depends on the effects induced by the molecule when administrated in a biological system.
-For instance, a molecule such as the Ximelagatran is indicated as anticoagulant because
-it directly inhibits the catalytic activity of the thrombin, 
-enzyme itself involved in the blood coagulation.
-In order to speak of this therapeutic effect (anticoagulant), people sometimes refer to 
-the **mode of action** of an active compound. 
-Wikipedia gives us the following definition for it:
-> A mode of action describes a functional or anatomical change, at the 
-> cellular level, resulting from the exposure of a living organism to a substance.
+\subsection{Regulatory pattern}
+Some of the FTC categories are created from the biological processes present in the 
+GO (cf previous sections); these categories have two arbitrary equivalent definitions, representing the 
+two possible ways a compound might impact the biological process. Anti-biological process agent FTC categories 
+contain the drugs that negatively perturb a target involved in the positive regulation of the biological process. 
+The anti categories also feature the compounds that positively perturb a negative regulator of the same process. 
+The pro categories are equivalent to the opposite pattern. Figure 5 illustrates the equivalent definitions for the 
+FTC class Anti-blood coagulation agent (FTC_A0007596).
 
-Characterising the mode of action of drugs is interesting for the following reasons:
-It is a central link between the molecular structure of a drug and its indication towards a disease. Drugs with similar
-mode of actions could be used to treat similar conditions.
-It abstracts away from the chemical structure and allows to reason over the effects of the compound in the body. The mode of action
-of a drug is a feature against which compounds can be classified.
-It can captures polypharmacology. Most of the drugs are not magic bullets
-and produce various side effects. A drug can have several mode of actions, representing the different processes the drug is perturbing directly
-or indirectly.
-A representation of the mode of action helps to better understand and analyse the behaviour of an agent in the human body.
+\begin{figure}[!tpb]%figureS1
+\centerline{\includegraphics{figS1.png}}
+\caption{Example of equivalent definitions (gray boxes) for the FTC category Anti-blood coagulation 
+agent (FTC\_A0007596). After the data integration step, a reasoner will look the definitions and identify 
+which drugs comply with the FTC definitions. Definitions are expressed using the Web Ontology Language (OWL), 
+serialised here using the Manchester syntax.}\label{fig:S01}
+\end{figure}
 
-Although the **mode of action** is an old and useful notion in the drug discovery 
-community (see number of results for thisPubMed search, no resources were
-previously fully dedicated to it. The aim of the FTC is to fill this gap.
+\subsection{Functional pattern}
+The FTC categories generated from the GO molecular functions are also equivalent to a logical 
+definition. Anti FTC categories dealing with molecular activities are asserted as equals to the drugs 
+that negatively perturb the function. Pro categories are equivalent to the drugs that positively perturb the 
+function of interest. A summary of the patterns definitions is available on the 
+online wiki at {{https://github.com/loopasam/ftc/wiki/Mode-of-Action}}.
 
-The FTC can be seen as a toolbox designed to fix dysfunctioning biological systems.
-It features over 20'000 compartments (categories) inside which approved 
-drugs are automatically classified, based on facts from various databases.
-FTC categories represent the mode of action of compounds in the human system. 
-The resource can be used for various tasks where a structured vocabulary is needed, such as compound annotation, drug repurposing analysis and 
-anything else you can think of!
+\section{Data integration}
+At this stage, the knowledge base contains the created FTC classes associated with their 
+logical definitions, as well as the GO and the core FTC entities. The knowledge base is then 
+further populated with some information coming from various public databases. Only manually curated 
+information is considered. This section illustrates how the core entities interact with the different data types 
+in order to create the axioms.
 
-% Results and Discussion can be combined.
-\section*{Results}
+\subsection{DrugBank}
+The DrugBank database is a unique bioinformatics and cheminformatics resource 
+that combines detailed drug data with comprehensive drug target information (cite). The approved 
+drugs (small molecules and biotherapeutics) acting on proteins are extracted from the database and 
+imported in the FTC knowledge base. In order to be selected, a compound must firstly be approved and 
+secondly have a pharmacological action on at least one human protein target present in Uniprot (cite). 
+The protein targets all have at least one manually asserted GO annotation for a biological process or a 
+molecular function. DrugBank links compounds to targets via actions. The DrugBank actions are somehow 
+structured and consistent: Concepts such as inhibitor or agonist are reused throughout the database for 
+example, yet they are not strictly formalised as a controlled vocabulary. These actions are manually 
+standardised to the core properties of the FTC according to their meaning: For instance the action 
+antagonist is mapped to the FTC negatively-perturbs.
+Compounds coming from DrugBank are represented as OWL classes and asserted 
+as subclasses of DrugBank compound (FTC\_C2). Protein targets are described as classes 
+too and subclasses of the core class protein. Each DrugBank compound is then connected to 
+its target via the following axiom pattern: drug SubClassOf perturbs some protein. 
+E.g. Ximelagatran SubClassOf negatively-perturbs some Prothrombin.
 
-The results section should provide details of all of the experiments that 
-are required to support the conclusions of the paper. There is no specific word limit for 
-this section, but details of experiments that are peripheral to the main thrust of the article and that 
-detract from the focus of the article should not be included. The section may be divided into subsections, 
-each with a concise subheading. Large datasets, including raw data, should be submitted as supporting files; 
-these are published online alongside the accepted article. The results section should be written in the past tense.
+\subsection{Gene Ontology annotations}
+The GO annotation program aims to provide high-quality GO annotations to 
+proteins in UniProt (cite). In the context of the FTC, such annotations are used 
+to create axioms linking protein targets to molecular functions and biological processes. 
+Each protein annotated with a function creates an axiom such 
+as protein SubClassOf has-function some molecular function. 
+Each protein annotated to a biological process creates an axiom such 
+as protein SubClassOf involved-in some biological process. 
+E.g. Prothrombin SubClassOf involved-in some positive regulation of blood coagulation. 
+Each protein can be involved in multiple processes and capable of realizing multiple functions.
 
-\subsection*{Subsection 1}
+\section{Knowledge base classification}Supplementary Material
+The knowledge base is fully built at this step and contains core classes, mode 
+of actions descriptions alongside the actions of approved DrugBank compounds on 
+protein targets in Uniprot. The proteins are linked to their molecular functions and involvement in 
+biological processes via the GO annotations. The logical specifications of the FTC are there to glue 
+the different data together and to explicitly express the logical links between resources. The FTC 
+knowledge base follows an OWL2 EL profile, which enable the use of fast and parallelised reasoners such 
+as ELK (cite). During the classification process, the reasoner checks whether the mode of action equivalent 
+definitions are satisfied or not and assigns drugs inside the corresponding FTC categories. The tree structure 
+of FTC appears also at this step from the logical definitions.
 
-\subsection*{Subsection 2}
+\section{Evaluation methodology}
+As the classification of therapeutic agents is done in an automated way, it 
+is important to evaluate the results generated against a known resource which will be 
+considered as gold standard. The assessment of the FTC is done against another similar 
+classification, the Anatomical Therapeutic Chemical Classification System (ATC) (cite). 
+The ATC has been developed to serve as a tool for drug utilization research in order to 
+improve quality of drug use. In this resource, the information is manually curated, and drugs 
+are assigned to categories based of their legally approved indications. The goal of the ATC differs 
+from the one of the FTC, yet the two resources are sharing some very similar concepts, which can be 
+used for the evaluation. Categories of both classifications contain approved drugs with a DrugBank 
+identifier, meaning that some of the drugs indexed in the FTC are also present in the ATC. From that, 
+it is possible to define some evaluation points, which will help to assess the automated classification process.
 
-\section*{Discussion}
+\subsection{Evaluation Points}
+An evaluation point is defined as an equivalence between a class from the FTC with one or 
+more classes from the ATC. The idea is to look at the set of drugs contained in both side of the 
+equivalence and estimate the overlap. This is illustrated in the Figure 6. Evaluation points are defined 
+by hand and not themselves evaluated. The full list of evaluation points as well as a summary of the results 
+are available online at https://www.ebi.ac.uk/chembl/ftc/evaluation/. Each evaluation point has a series of 
+true/false positive and false negative drugs associated with it.
 
-The discussion should spell out the major conclusions of the work along with some 
-explanation or speculation on the significance of these conclusions. How do the 
-conclusions affect the existing assumptions and models in the field? How can future research 
-build on these observations? What are the key experiments that must be done? The discussion should 
-be concise and tightly argued. The results and discussion may be combined into one section, if desired.
+Figure 6: Example of evaluation point. Classes from the ATC are manually 
+mapped to classes from the FTC. The drugs present in each side of the equivalences 
+are compared, and some metrics are derived from it (false positive/negative and true positive). 
+A summary of the evaluation is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.
 
-% You may title this section "Methods" or "Models". 
-% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
-% authors may use "Analysis" 
-\section*{Materials and Methods}
+\subsection{True Positives}
+Drugs that are present in both the FTC and the equivalent ATC class(es) are called true positives. 
+These compounds reflect that the automated classification was capable of retrieving correctly the 
+information present in the gold standard (ATC).
 
-This section should provide enough detail for reproduction of the findings. Protocols 
-for new methods should be included, but well-established protocols may simply be referenced. While we do 
-encourage authors to submit all appendices, detailed protocols, or details of the algorithms for newer or less 
-well-established methods, please do so as Supporting Information files. These are not included in the typeset manuscript, 
-but are downloadable and fully searchable from the HTML version of the article.
+\subsection{False Negatives}
+These drugs are present in the ATC class(es) but not in the corresponding FTC class. 
+The automated classification failed to retrieve these compounds. The smaller the number of 
+false negatives is, the better the FTC is at recalling drugs. A small number of false negatives 
+means that if a drug is present in the ATC (gold standard), then it is likely that this drug will 
+also be correctly categorised in the FTC.
 
-% Do NOT remove this, even if you are not including acknowledgments
-\section*{Acknowledgments}
+\subsection{False Positives}
+The false positives are the drugs present in the FTC category of the evaluation point but not in the corresponding ATC classes. 
+A high number of false positives means that the FTC is over-assigning compounds to classes.
+The false positives relates to the accuracy of the classification. In the context of this work, 
+some false positives could also be considered as drug repurposing opportunities.
 
-People who contributed to the work but do not fit the criteria for authors should be listed in the 
-Acknowledgments, along with their contributions. You must also ensure that anyone named in the Acknowledgments 
-agrees to being so named.
+\subsection{Precision}
+Precision is the probability that a randomly selected drug from the FTC is present in the equivalent ATC classes. 
+The value is standardised as a percentage and corresponds to the formula: True Positive / (True Positive + False Positive).
 
-Details of the funding sources that have supported the work should be confined to the funding statement 
-provided in the online submission system. Do not include them in the Acknowledgments.
+\subsection{Recall}
+Recall is the probability that a randomly selected drug from the ATC has been assigned to the correct 
+corresponding class in the FTC. The value is standardised as a percentage and corresponds 
+to the formula: True Positive / (True Positive + False Negative).
 
-%\section*{References}
-% The bibtex filename
-\bibliography{document}
+\section{Semantic similarity}
+The semantic similarity measure performed over the FTC is a derivative of the Jaccard index (cite). 
+It is probably best understood as an example: If we consider two classes A and B, the semantic similarity 
+between these classes corresponds to the number of OWL superclasses (direct and indirect, obtained with a reasoner) that 
+are shared by A and B (intersection) divided by the number of superclasses of A or B (union). The index ranges 
+from 0 (totally different) to 1 (identical).
 
-\section*{Figure Legends}
-%\begin{figure}[!ht]
-%\begin{center}
-%%\includegraphics[width=4in]{figure_name.2.eps}
-%\end{center}
-%\caption{
-%{\bf Bold the first sentence.}  Rest of figure 2  caption.  Caption 
-%should be left justified, as specified by the options to the caption 
-%package.
-%}
-%\label{Figure_label}
-%\end{figure}
+\section{Mode of action similarity against indication}
+A statistical analysis was performed over the data presented on Figure 3. 
+When two compounds are randomly taken, they have on average a higher mode of action similarity when 
+they are assigned to the same ATC category (one ATC level). In order to estimate whether this observation was 
+due to chance only, we formulated the following null hypothesis (H0): For a pair of drug A and B, it does not 
+matter to which ATC category they belong to, the average similarity is always the same. 
+The alternative hypothesis (H1) was: For a pair of drug A and B, if A and B have the same ATC code, 
+we expect on average to obtain a higher similarity value than if A and B have different ATC codes. 
+A permutation test was then performed for each ATC category. For example, we started with the ATC 
+category A (first row on Figure 3), looked at the similarity values when pairs of compounds both belong 
+to the category A (top right corner square) and compared it to the similarity values when the pair of compounds 
+belong to different categories (A and B, A and C and so forth). For each comparison we obtain two distributions of 
+values (not shown). On average the similarity values are always higher when the two compounds 
+belong to the same category (A/A versus A/B for instance). A permutation test (n = 20'000) was 
+performed in order to see whether this observation was due to chance only. We were able to reject the null 
+hypothesis for a significance level of 0.05 all the times. As the test is performed multiple times, 
+we applied a Bonferroni correction for multiple hypothesis testing, which changed the significance level accordingly. 
+The choice for a permutation test was driven by the fact that MoA similarity values do not follow any type 
+of standard distribution (data not shown). For the significance levels set, the null hypothesis was rejected for 
+all first level ATC categories.
 
-The aim of the figure legend should be to describe the key messages of the figure, but 
-the figure should also be discussed in the text. An enlarged version of the figure and its full legend 
-will often be viewed in a separate window online, and it should be possible for a reader to understand the figure 
-without switching back and forth between this window and the relevant parts of the text. Each legend should have 
-a concise title of no more than 15 words that can stand alone, without the use of figure part labels. The overall 
-legend itself should be succinct, while still explaining all figure parts, symbols and abbreviations. Avoid lengthy descriptions of methods.
+\section{Knowledge base specification}
+This section presents the scaffold of the knowledge base underlying the FTC. 
+The logic structuring the FTC comes essentially from a set of core OWL properties (rich RBox). 
+Some of these properties originate from the GO. When necessary some new ones have also been introduced. 
+In order to understand how these properties interact, first will be presented the fundamental classes present 
+at the top of the FTC classification.
 
-\section*{Tables}
-%\begin{table}[!ht]
-%\caption{
-%\bf{Table title}}
-%\begin{tabular}{|c|c|c|}
-%table information
-%\end{tabular}
-%\begin{flushleft}Table caption
-%\end{flushleft}
-%\label{tab:label}
-% \end{table}
+\subsection{Core classes}
+The high level concepts covered by the FTC are represented as OWL classes and enumerated below. 
+Some of these core classes are coming from external ontologies, in which case the original URI is preserved.
 
-Tables should be included at the end of the manuscript file and cited sequentially in the text. 
-All tables should have a concise title. Footnotes can be used to explain abbreviations. 
-Citations should be indicated using the same style as outlined above. Tables should not occupy more than one 
-printed page; larger tables can be published as online supporting information. Tables must be cell-based; 
-do not use picture elements, text boxes, tabs, or returns in tables. Please ensure that all tables conform to our 
-Guidelines for Figure and Table Preparation when preparing them.
+\begin{itemize}
+\item \textbf{molecular function} \\
+Identifier: http://purl.obolibrary.org/obo/GO\_0003674 \\
+Label: molecular function \\
+Definition: As defined by the Gene Ontology. Elemental activities, such as catalysis or binding, describing 
+the actions of a gene product at the molecular level. A given gene product may exhibit one or more molecular functions.
+\end{itemize}
 
+\begin{itemize}
+\item \textbf{biological process} \\
+Identifier: http://purl.obolibrary.org/obo/GO\_0008150 \\
+Label: biological process \\
+Definition: As defined by the Gene Ontology: Any process specifically pertinent to the functioning of 
+integrated living units: cells, tissues, organs, and organisms. A process is a collection of molecular 
+events with a defined beginning and end.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Protein} \\
+Identifier: http://purl.uniprot.org/core/Protein \\
+Label: protein \\
+Definition: As defined by Uniprot. Description of a protein.
+Comment: Gene products present inside the FTC are all human proteins. Uniprot URIs are used.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Drug} \\
+Identifier: http://schema.org/Drug \\
+Label: drug \\
+Definition: As defined by schema.org. A chemical or biologic substance, used as a medical therapy, that has a 
+physiological effect on an organism.
+Comment: In the context of the FTC, DrugBank chemicals are considered for their role as therapeutic agent 
+rather than for their chemical structure. Mode of action classes are representing agent types and are subclasses of this concept.
+\end{itemize}
+
+therapeutic agent
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_C1
+Label: therapeutic agent
+Definition: Role of a drug capable of producing a therapeutic effect.
+
+DrugBank compound
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_C2
+Label: DrugBank compound
+Definition: Drug coming from DrugBank.
+
+\subsection{Core properties}
+The expressivity of the FTC comes mostly for the properties (RBox). Below is a list of the 
+OWL object properties present inside the FTC knowledge base.
+part-of
+Identifier: http://purl.obolibrary.org/obo/BFO_0000050
+Characteristic: Transitive
+Label: part-of
+Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#partof
+has-part
+Identifier: http://purl.obolibrary.org/obo/BFO_0000051
+Characteristic: Transitive
+Label: has-part
+Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology-ext.relations.shtml#haspart
+regulates
+Identifier: http://purl.obolibrary.org/obo/RO_0002211
+Chained property: regulates o part-of -> regulates
+Label: regulates
+Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
+negatively-regulates
+Identifier: http://purl.obolibrary.org/obo/RO_0002212
+SubPropertyOf: regulates
+Label: negatively-regulates
+Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
+positively-regulates
+Identifier: http://purl.obolibrary.org/obo/RO_0002213
+SubPropertyOf: regulates
+Label: positively-regulates
+Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
+involved-in
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R1
+Label: involved-in
+Domain: protein
+Range: biological process
+Definition: Entails the participation of a protein in a biological process
+has-function
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R2
+Label: has-function
+Domain: protein
+Range: molecular function
+Definition: Describes the molecular function a protein can realize.
+perturbs
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R3
+Label: perturbs
+Domain: drug
+Range: protein
+Definition: Specific biochemical interaction through which a drug substance 
+will affect the activity of a protein (mechanism of action). The property refers to the specific molecular 
+targets to which the drug binds, such as an enzyme or receptor.
+negatively-perturbs
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R4
+Label: negatively-perturbs
+SubPropertyOf: perturbs
+Definition: Specific biochemical interaction through which a drug substance will decrease the activity of a protein. 
+The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
+positively-perturbs
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R5
+Label: positively-perturbs
+SubPropertyOf: perturbs
+Definition: Specific biochemical interaction through which a drug substance will increase the activity of a protein. 
+The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
+
+%\bibliographystyle{natbib}
+%\bibliographystyle{achemnat}
+%\bibliographystyle{plainnat}
+%\bibliographystyle{abbrv}
+%\bibliographystyle{bioinformatics}
+%
+%\bibliographystyle{plain}
+%
+%\bibliography{Document}
+
+\begin{thebibliography}{}
+\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03} Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.
+
+\bibitem[Bag {\it et~al}., 2001]{Bag01} Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.
+
+\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
+Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
+in nigral dopaminergic neurnol cell: correlation with the known
+pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
+Res.}, \textbf{110}(Suppl. 1), 76--84.
+
+\bibitem[Lehmann, 1986]{Leh86}
+Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.
+
+\bibitem[Crenshaw and Jones, 2003]{Cre03}
+Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
+cancer management: one tumor, one chip. \textit{Bioinformatics},
+doi:10.1093/bioinformatics/btn000.
+
+\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
+Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
+(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
+???--???.
+
+\bibitem[Bardet, 1920]{Bar20}
+Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
+polydactylie et retinite pigmentaire (contribution a l'etude des
+formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
+institution, Paris, France.
+
+\end{thebibliography}
 \end{document}

### 2013-08-15 document.tex
supp figures added
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 relies mostly on Brain (cite) and the web application builds on the top of the Play! framework (cite). 
-Classification tasks use the ELK reasoner (cite). The computer hosting the web application has 8Go of memory 
+Classification tasks use the ELK reasoner (cite). The computer hosting the web application has 8 Gb of memory 
 with 4 processors, this architecture allows fast parallel reasoning, thanks to ELK’s design. More functionalities 
@@
 The pro categories are equivalent to the opposite pattern. Figure 5 illustrates the equivalent definitions for the 
-FTC class Anti-blood coagulation agent (FTC_A0007596).
+FTC class Anti-blood coagulation agent (FTC\_A0007596).
 
@@
 
-Figure 6: Example of evaluation point. Classes from the ATC are manually 
+\begin{figure}[!tpb]%figureS2
+\centerline{\includegraphics{figS2.png}}
+\caption{Example of evaluation point. Classes from the ATC are manually 
 mapped to classes from the FTC. The drugs present in each side of the equivalences 
 are compared, and some metrics are derived from it (false positive/negative and true positive). 
-A summary of the evaluation is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.
+A summary of the evaluation is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.}\label{fig:S02}
+\end{figure}
 
@@
 \begin{itemize}
-\item \textbf{molecular function} \\
+\item \textbf{Molecular Function} \\
 Identifier: http://purl.obolibrary.org/obo/GO\_0003674 \\
@@
 \begin{itemize}
-\item \textbf{biological process} \\
+\item \textbf{Biological Process} \\
 Identifier: http://purl.obolibrary.org/obo/GO\_0008150 \\
@@
 Definition: As defined by schema.org. A chemical or biologic substance, used as a medical therapy, that has a 
-physiological effect on an organism.
+physiological effect on an organism. \\
 Comment: In the context of the FTC, DrugBank chemicals are considered for their role as therapeutic agent 
@@
 
-therapeutic agent
-Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_C1
-Label: therapeutic agent
+\begin{itemize}
+\item \textbf{Therapeutic Agent} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_C1 \\
+Label: therapeutic agent \\
 Definition: Role of a drug capable of producing a therapeutic effect.
+\end{itemize}
 
-DrugBank compound
-Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_C2
-Label: DrugBank compound
+\begin{itemize}
+\item \textbf{DrugBank Compound} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_C2 \\
+Label: DrugBank compound \\
 Definition: Drug coming from DrugBank.
+\end{itemize}
 
@@
 OWL object properties present inside the FTC knowledge base.
-part-of
-Identifier: http://purl.obolibrary.org/obo/BFO_0000050
-Characteristic: Transitive
-Label: part-of
-Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#partof
-has-part
-Identifier: http://purl.obolibrary.org/obo/BFO_0000051
-Characteristic: Transitive
-Label: has-part
-Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology-ext.relations.shtml#haspart
-regulates
-Identifier: http://purl.obolibrary.org/obo/RO_0002211
-Chained property: regulates o part-of -> regulates
-Label: regulates
-Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
-negatively-regulates
-Identifier: http://purl.obolibrary.org/obo/RO_0002212
-SubPropertyOf: regulates
-Label: negatively-regulates
-Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
-positively-regulates
-Identifier: http://purl.obolibrary.org/obo/RO_0002213
-SubPropertyOf: regulates
-Label: positively-regulates
-Definition: As defined and used in the Gene Ontology. More information at http://www.geneontology.org/GO.ontology.relations.shtml#regulates
-involved-in
-Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R1
-Label: involved-in
-Domain: protein
-Range: biological process
-Definition: Entails the participation of a protein in a biological process
-has-function
-Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R2
-Label: has-function
-Domain: protein
-Range: molecular function
-Definition: Describes the molecular function a protein can realize.
-perturbs
-Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R3
-Label: perturbs
-Domain: drug
-Range: protein
+
+\begin{itemize}
+\item \textbf{Part Of} \\
+Identifier: http://purl.obolibrary.org/obo/BFO\_0000050 \\
+Characteristic: Transitive \\
+Label: part-of \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Has Part} \\
+Identifier: http://purl.obolibrary.org/obo/BFO\_0000051 \\
+Characteristic: Transitive \\
+Label: has-part \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002211 \\
+Chained property: regulates o part-of -> regulates \\
+Label: regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Negatively Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002212 \\
+SubPropertyOf: regulates \\
+Label: negatively-regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Positively Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002213 \\
+SubPropertyOf: regulates \\
+Label: positively-regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Involved In} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R1 \\
+Label: involved-in \\
+Domain: protein \\
+Range: biological process \\
+Definition: Entails the participation of a protein in a biological process. \\
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Has Function} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R2 \\
+Label: has-function \\
+Domain: protein \\
+Range: molecular function \\
+Definition: Describes the molecular function a protein can realize. \\
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R3 \\
+Label: perturbs \\
+Domain: drug \\
+Range: protein \\
 Definition: Specific biochemical interaction through which a drug substance 
@@
 targets to which the drug binds, such as an enzyme or receptor.
-negatively-perturbs
-Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R4
-Label: negatively-perturbs
-SubPropertyOf: perturbs
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Negatively Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R4 \\
+Label: negatively-perturbs \\
+SubPropertyOf: perturbs \\
 Definition: Specific biochemical interaction through which a drug substance will decrease the activity of a protein. 
 The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
-positively-perturbs
-Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC_R5
-Label: positively-perturbs
-SubPropertyOf: perturbs
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Positively Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R5 \\
+Label: positively-perturbs \\
+SubPropertyOf: perturbs \\
 Definition: Specific biochemical interaction through which a drug substance will increase the activity of a protein. 
 The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
+\end{itemize}
+
+\section*{Pairwise comparison of mode of action similarities}
+
+Supplementary Figure S1: Pairwise comparison of mode of action similarities. 
+The compounds are grouped by ATC code (2 levels - legend not shown) on panel A. 
+A hierarchical clustering step (based on manhattan distance) was applied on panel B. 
+The dendrogram reflects the original structure of the FTC. 
+Panel C shows the compounds sorted by DrugBank identifiers only (≈ random). 
+The colours displayed on panel B and C correspond to one level ATC categories and are indexed in the legend panel.
+
+\begin{figure*}[!tpb]%figureS3
+\centerline{\includegraphics{figS3.png}}
+\caption{Caption to be written.}\label{fig:S03}
+\end{figure*}
 

### 2013-08-15 document.tex
supp figures all included
Signed-off-by: loopasam <samuel.croset@gmail.com>
references added
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-08-15 document.tex
@@
 
-\title[short Title]{The Functional Therapeutic Chemical Classification System}
+\title[The Functional Therapeutic Chemical Classification System]{The Functional Therapeutic Chemical Classification System}
 \author[Croset \textit{et~al}]{Samuel Croset\,$^{1, \footnote{to whom correspondence should be addressed}}$, John Overington\,$^{1}$ and Dietrich Rebholz-Schuhmann\,$^{1}$
@@
 
-Drug repurposing is the application of known drugs and compounds to new indications (i.e., new diseases) (cite wikipedia and better source). 
+Drug repurposing is the use of known active compounds for new therapeutic indications \citep{Sanseau01072011}. 
 When administered in a living organism, a compound can indeed play various roles and affect different biological processes. 
 Accurately identifying these different functions helps to predict the potential side-effects a drug could have and can also lead to 
-interesting repurposing opportunities (http://www.ncbi.nlm.nih.gov/pubmed/23340113). For instance, sildenafil was initially developed 
-against angina and has been repurposed towards erectile dysfunction during the clinical trials (cite). Approved compounds are 
+interesting repurposing opportunities \citep{Medina-Franco2013}. For instance, sildenafil was initially developed 
+against angina and has been repurposed towards erectile dysfunction during the clinical trials \citep{Ashburn2004}. Approved compounds are 
 attractive because they have been extensively studied and have by definition already successfully passed clinical trials, where 
 most drugs fail because of safety or efficacy issues.
-There is an increasing number of approaches to predict repurposing opportunities using computational methods (cite). Most 
-methods operate on the profiles of physicochemical descriptors derived from molecular structures (cite). Other methods characterise 
-the drugs on more abstract levels, such as the gene expression signature (cite) or via the reported side-effects (cite). These approaches 
+There is an increasing number of approaches to predict repurposing opportunities using computational methods \citep{Sanseau01072011}. Most 
+methods operate on the profiles of physicochemical descriptors derived from molecular structures \citep{Haupt2011}. Other methods characterise 
+the drugs on more abstract levels, such as the gene expression signature \citep{Iorio2010} or 
+via the reported side-effects \citep{Campillos2008}. 
+These approaches 
 have in common to look for similarities within existing drugs and forward similar compounds as repurposing hypotheses.
-A feature of particular interest to describe drugs is the mode of action (MoA). The MoA describes a functional or anatomical change, 
-at the cellular level, resulting from the exposure of a living organism to a substance (cite wikipedia or better source). For instance 
+A feature of particular interest to describe drugs is the mode of action (MoA). According to Wikipedia, 
+the MoA describes a functional or anatomical change, 
+at the cellular level, resulting from the exposure of a living organism to a substance. For instance 
 terms such as transcriptional regulation agent or anticoagulant define MoAs and characterise the roles of a certain type of drugs. The 
@@
 them straightforwardly with values and numbers like one can do for a 3D molecular structure or for a gene expression profile. 
-Nonetheless, the meaning of a concept can be formalised with controlled vocabularies and ontologies (cite); originating from description logics, such frameworks help to define the mathematical meaning of symbols and strings of characters.
+Nonetheless, the meaning of a concept can be formalised with controlled vocabularies and ontologies \citep{Gruber1995}; 
+originating from description logics, such frameworks help to define the mathematical meaning of symbols and strings of characters. 
 In an ontology or knowledge base, concepts are organised and linked following the logical type of relation they have between them. 
-In the Gene Ontology (GO) for example (cite), biological processes and molecular functions are manually curated in this resource and the meaning of terms is specified by the relations linking two GO terms.
-MoA definitions are present in other classifications such as the Medical Subject Headings (cite) or the Chemical Entities of 
-Biological Interest (cite) for example. The Anatomical Therapeutic Chemical Classification System (ATC) also describes to some 
+In the Gene Ontology (GO) for example \citep{Ashburner2000}, biological processes and molecular functions are manually curated in this resource 
+and the meaning of terms is specified by the relations linking two GO terms.
+MoA definitions are present in other classifications such as the Medical Subject Headings \citep{Nelson2004} or the Chemical Entities of 
+Biological Interest \citep{Hastings2012} for example. The Anatomical Therapeutic 
+Chemical Classification System (ATC) \citep{world2000anatomical} also describes to some 
 extent the action of drugs at the anatomical level. All these resources are valuable for the community as a source of carefully 
@@
 Classification System (FTC), presented here in this manuscript. The FTC is automatically built by leveraging the content 
-of various biomedical databases using description logics and automated reasoning. Over 20’000 new MoA categories are defined 
+of various biomedical databases using description logics and automated reasoning. Over 20'000 new MoA categories are defined 
 in the resource and further populated with approved drugs using the Web Ontology Language (OWL) in combination with a reasoner. 
@@
 Finally, we illustrate also how the FTC can be used as a pharmacological toolbox to analyse drug repurposing for Alzheimer’s disease.
-
- Text Text Text Text. Figure~\ref{fig:01} shows that the above method  Text 
-   Text Text.  \citep{Bag01} wants to know about text follows
 
@@
 FTC (see Availability and Implementation). A reasoner, a program capable of understanding such axioms, performed this task. 
-The process to build the FTC is summarised in Figure 0.
+The process to build the FTC is summarised in Figure~\ref{fig:01}.
 
@@
 the classification time was always superior to a few minutes (data not shown). The FTC has a taxonomic structure as illustrated 
-on Figure 1, which arises when the reasoner classifies the knowledge base. Categories can have multiple parents and multiple children. 
+on Figure~\ref{fig:02}, which arises when the reasoner classifies the knowledge base. Categories can have multiple parents and multiple children. 
 The reader can browse and access the content of the FTC online at {{https://www.ebi.ac.uk/chembl/ftc}}.
@@
 The subgroups B, C and D are inhibitors of the cholinergic system and some of them, such as galantamine (DB00674), are already 
-investigated to treat Alzheimer’s disease and other related dementias (pubchem). This class of agent is in line with the 
-cholinergic hypothesis (cite PMID:10071091), stating that Alzheimer’s disease could be caused by dysfunctions in the processing of acetylcholine.
-The subgroup E is exclusively composed of barbiturates (central nervous system depressants).
-The presence of this pharmacological class of compounds as an Alzheimer treatment is more surprising, as very little literature reports on it.
-Further investigations reveal that the neuronal acetylcholine receptor subunit alpha-7 (P36544), a common off-target of barbiturates, binds beta-amyloids with high affinity.
-As beta-amyloids are themselves strongly involved in the pathology (cite http://europepmc.org/abstract/MED/10681545), barbiturates could affect the state of the condition, similarly to cholinergic inhibitors.
+investigated to treat Alzheimer’s disease and other related dementias. This class of agent is in line with the 
+cholinergic hypothesis \citep{Francis1999}, stating that Alzheimer’s disease could be caused by dysfunctions in the processing of 
+acetylcholine. The subgroup E is exclusively composed of barbiturates (central nervous system depressants). 
+The presence of this pharmacological class of compounds as an Alzheimer treatment is more surprising, as very little 
+literature reports on it. Further investigations reveal that the neuronal acetylcholine receptor subunit alpha-7, a 
+common off-target of barbiturates, binds beta-amyloids with high affinity \citep{Wang2000}. As beta-amyloids are themselves strongly involved 
+in the pathology, barbiturates could affect the state of the condition \citep{Wang2000}, similarly to cholinergic inhibitors.
 The group F contains four compounds; nicotine and varenicline have been grouped because of similar pharmacology (group G). 
-Nicotine has been shown to improve some of the symptoms of Alzheimer’s disease (PMID:1410164), it is therefore expected to 
+Nicotine has been shown to improve some of the symptoms of Alzheimer’s disease \citep{Jones1992}, it is therefore expected to 
 find this molecule in the predictions. Varenicline possess a pharmacology related to that of nicotine, which would explain 
@@
 The groups H and I contain respectively one molecule each. These compounds have been classified as agents perturbing some of 
-the physiological function of the Tau protein, key actor in Alzheimer's disease (cite). Vorinostat (group H) is currently 
-indicated for the treatment of cutaneous manifestations in patients with T-cell lymphoma (drugbank), yet a study has 
+the physiological function of the Tau protein, key actor in Alzheimer's disease \citep{Grundke-Iqbal1986}. Vorinostat (group H) is currently 
+indicated for the treatment of cutaneous manifestations in patients with T-cell lymphoma, yet a study has 
 shown in vivo (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
-deficit (cite http://www.nature.com/npp/journal/v35/n4/full/npp2009197a.html). The presence of lithium (group I) was 
+deficit \citep{Kilgore2010}. The presence of lithium (group I) was 
 unexpected and suspected of being a false positive at first glance. However a recent study demonstrated a long-term 
-protective effect for the ion in regards to Alzheimer’s disease (PMID:22500970 and 21525519). The last group (J) contains 
+protective effect for the ion in regards to Alzheimer’s disease \citep{Young2011}. The last group (J) contains 
 ezetimibe and hesperetin. These two compounds are primarily used as cholesterol lowering agents (statins). As cholesterol 
 metabolism in the brain appears to be related to dementia, statins are believed to prevent or improve the symptoms of the 
-patients. Although early studies (PMID:17877925) have failed to clearly show a beneficial effect, the investigation is still open.
+patients. Although early studies \citep{Wolozin2004} have failed to clearly show a beneficial effect, the investigation is still open.
 From the examples briefly presented above, reported and confirmed by the literature, the FTC appears to be suitable to 
@@
 
-%\bibliographystyle{natbib}
-%\bibliographystyle{achemnat}
-%\bibliographystyle{plainnat}
-%\bibliographystyle{abbrv}
-%\bibliographystyle{bioinformatics}
-%
-%\bibliographystyle{plain}
-%
-%\bibliography{Document}
+\bibliographystyle{natbib}
+\bibliographystyle{achemnat}
+\bibliographystyle{plainnat}
+\bibliographystyle{abbrv}
+\bibliographystyle{bioinformatics}
 
-\begin{thebibliography}{}
-\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03} Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.
+\bibliographystyle{plain}
 
-\bibitem[Bag {\it et~al}., 2001]{Bag01} Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.
+\bibliography{document}
 
-\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
-Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
-in nigral dopaminergic neurnol cell: correlation with the known
-pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
-Res.}, \textbf{110}(Suppl. 1), 76--84.
-
-\bibitem[Lehmann, 1986]{Leh86}
-Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.
-
-\bibitem[Crenshaw and Jones, 2003]{Cre03}
-Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
-cancer management: one tumor, one chip. \textit{Bioinformatics},
-doi:10.1093/bioinformatics/btn000.
-
-\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
-Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
-(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
-???--???.
-
-\bibitem[Bardet, 1920]{Bar20}
-Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
-polydactylie et retinite pigmentaire (contribution a l'etude des
-formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
-institution, Paris, France.
-
-\end{thebibliography}
+%\begin{thebibliography}{}
+%\bibitem[Bardet, 1920]{Bar20}
+%Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
+%polydactylie et retinite pigmentaire (contribution a l'etude des
+%\end{thebibliography}
 \end{document}

### 2013-08-18 document.tex
references added
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-08-15 document.tex
@@
 available on the wiki at {{https://github.com/loopasam/ftc/wiki/Description-Logics}}. The FTC implementation 
-relies mostly on Brain (cite) and the web application builds on the top of the Play! framework (cite). 
-Classification tasks use the ELK reasoner (cite). The computer hosting the web application has 8 Gb of memory 
+relies mostly on Brain \citep{Croset2013} and the web application builds on the top of the Play! framework. 
+Classification tasks use the ELK reasoner \citep{Kazakov2011}. The computer hosting the web application has 8 Gb of memory 
 with 4 processors, this architecture allows fast parallel reasoning, thanks to ELK’s design. More functionalities 
@@
 The mode of action categories present in the FTC are defined based on the terms coming from the 
-Gene Ontology (GO). Both the molecular function and biological process branches are used for this purpose, yet handled differently.
+Gene Ontology (GO) \citep{Ashburner2000}. Both the molecular function and biological process branches are used for 
+this purpose, yet handled differently.
 
@@
 The DrugBank database is a unique bioinformatics and cheminformatics resource 
-that combines detailed drug data with comprehensive drug target information (cite). The approved 
+that combines detailed drug data with comprehensive drug target information \citep{Knox2011}. The approved 
 drugs (small molecules and biotherapeutics) acting on proteins are extracted from the database and 
 imported in the FTC knowledge base. In order to be selected, a compound must firstly be approved and 
-secondly have a pharmacological action on at least one human protein target present in Uniprot (cite). 
-The protein targets all have at least one manually asserted GO annotation for a biological process or a 
+secondly have a pharmacological action on at least one human protein target present in Uniprot \citep{TheUniprotConsortium2013}. 
+The protein targets all have at least one manually asserted GO annotation \citep{Dimmer2012} for a biological process or a 
 molecular function. DrugBank links compounds to targets via actions. The DrugBank actions are somehow 
@@
 too and subclasses of the core class protein. Each DrugBank compound is then connected to 
-its target via the following axiom pattern: drug SubClassOf perturbs some protein. 
-E.g. Ximelagatran SubClassOf negatively-perturbs some Prothrombin.
+its target via the following axiom pattern: \emph{drug SubClassOf perturbs some protein}. 
+E.g. \emph{Ximelagatran SubClassOf negatively-perturbs some Prothrombin}.
 
@@
 The GO annotation program aims to provide high-quality GO annotations to 
-proteins in UniProt (cite). In the context of the FTC, such annotations are used 
+proteins in UniProt. In the context of the FTC, such annotations are used 
 to create axioms linking protein targets to molecular functions and biological processes. 
@@
 
-\section{Knowledge base classification}Supplementary Material
+\section{Knowledge base classification}
 The knowledge base is fully built at this step and contains core classes, mode 
@@
 considered as gold standard. The assessment of the FTC is done against another similar 
-classification, the Anatomical Therapeutic Chemical Classification System (ATC) (cite). 
+classification, the Anatomical Therapeutic Chemical Classification System (ATC) \citep{world2000anatomical}). 
 The ATC has been developed to serve as a tool for drug utilization research in order to 
@@
 \section{Semantic similarity}
-The semantic similarity measure performed over the FTC is a derivative of the Jaccard index (cite). 
+The semantic similarity measure performed over the FTC is a derivative of the Jaccard index \citep{jaccard1912distribution} \citep{rogers1960computer}. 
 It is probably best understood as an example: If we consider two classes A and B, the semantic similarity 
@@
 
-%\bibliographystyle{natbib}
-%\bibliographystyle{achemnat}
-%\bibliographystyle{plainnat}
-%\bibliographystyle{abbrv}
-%\bibliographystyle{bioinformatics}
-%
-%\bibliographystyle{plain}
-%
-%\bibliography{Document}
+\begin{figure*}[!tpb]%figureS4
+\centerline{\includegraphics{figS4.png}}
+\caption{Caption to be written.}\label{fig:S04}
+\end{figure*}
 
-\begin{thebibliography}{}
-\bibitem[Bofelli {\it et~al}., 2000]{Boffelli03} Bofelli,F., Name2, Name3 (2003) Article title, {\it Journal Name}, {\bf 199}, 133-154.
+\begin{figure*}[!tpb]%figureS5
+\centerline{\includegraphics{figS5.png}}
+\caption{Caption to be written.}\label{fig:S05}
+\end{figure*}
 
-\bibitem[Bag {\it et~al}., 2001]{Bag01} Bag,M., Name2, Name3 (2001) Article title, {\it Journal Name}, {\bf 99}, 33-54.
+\bibliographystyle{natbib}
+\bibliographystyle{achemnat}
+\bibliographystyle{plainnat}
+\bibliographystyle{abbrv}
+\bibliographystyle{bioinformatics}
 
-\bibitem[Yoo \textit{et~al}., 2003]{Yoo03}
-Yoo,M.S. \textit{et~al}. (2003) Oxidative stress regulated genes
-in nigral dopaminergic neurnol cell: correlation with the known
-pathology in Parkinson's disease. \textit{Brain Res. Mol. Brain
-Res.}, \textbf{110}(Suppl. 1), 76--84.
+\bibliographystyle{plain}
 
-\bibitem[Lehmann, 1986]{Leh86}
-Lehmann,E.L. (1986) Chapter title. \textit{Book Title}. Vol.~1, 2nd edn. Springer-Verlag, New York.
+\bibliography{document}
 
-\bibitem[Crenshaw and Jones, 2003]{Cre03}
-Crenshaw, B.,III, and Jones, W.B.,Jr (2003) The future of clinical
-cancer management: one tumor, one chip. \textit{Bioinformatics},
-doi:10.1093/bioinformatics/btn000.
-
-\bibitem[Auhtor \textit{et~al}. (2000)]{Aut00}
-Auhtor,A.B. \textit{et~al}. (2000) Chapter title. In Smith, A.C.
-(ed.), \textit{Book Title}, 2nd edn. Publisher, Location, Vol. 1, pp.
-???--???.
-
-\bibitem[Bardet, 1920]{Bar20}
-Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
-polydactylie et retinite pigmentaire (contribution a l'etude des
-formes cliniques de l'obesite hypophysaire). PhD Thesis, name of
-institution, Paris, France.
-
-\end{thebibliography}
 \end{document}

### 2013-08-18 document.tex
references added
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-08-15 document.tex
@@
 The knowledge base behind the FTC is built by integrating information coming from various sources. 
-The GO terms serve as template to create the FTC categories describing the MoAs; DrugBank provides the 
-known links between drugs and their protein targets. Uniprot maps targets to their respective GO annotations. 
+The GO terms serve as template to create the FTC categories describing the MoAs; DrugBank \citep{Knox2011} provides the 
+known links between drugs and their protein targets. Uniprot \citep{TheUniprotConsortium2013} maps targets to their 
+respective GO annotations \citep{Dimmer2012}. 
 Drugs are further assigned into mode of action categories based of the OWL constructs and axioms defined in the 
@@
 It takes around 5 seconds on a modern 
-laptop (4 cores processor with 4 Gb of memory) to classify the knowledge base using the ELK reasoner (cite). 
-Other OWL reasoners have been tried (e.g, Hermit (cite), Pellet (cite), etc...) but none of them were suitable for this purpose; 
+laptop (4 cores processor with 4 Gb of memory) to classify the knowledge base using the ELK reasoner \citep{Kazakov2011}. 
+Other OWL reasoners have been tried (e.g, Hermit, Pellet, etc.) but none of them were suitable for this purpose; 
 the classification time was always superior to a few minutes (data not shown). The FTC has a taxonomic structure as illustrated 
@@
 The code behind the creation of the resource is entirely open and available 
-at {{https://github.com/loopasam/ftc}}. The web application built on the top of the FTC can be 
-find at {{https://www.ebi.ac.uk/chembl/ftc}} and the documentation can be accessed at {{https://github.com/loopasam/ftc/wiki}}. The full
+at {{https://github.com/loopasam/ftc}}. The web application built on the top of the FTC using the Brain library \citep{Croset2013}, and can be 
+find at {{https://www.ebi.ac.uk/chembl/ftc}}. The documentation can be accessed at {{https://github.com/loopasam/ftc/wiki}}. The full
 description of the methodology used to generate the classification is available as supplementary material on Bioinformatics website.

### 2013-08-19 document.tex
final editing on the way
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-08-18 document.tex
@@
 Classification tasks use the ELK reasoner \citep{Kazakov2011}. The computer hosting the web application has 8 Gb of memory 
-with 4 processors, this architecture allows fast parallel reasoning, thanks to ELK’s design. More functionalities 
-will be added to the web application following user requirements (lean implementation).
+with 4 processors, this architecture allows fast parallel reasoning, thanks to ELK's design. More functionalities 
+will be added to the web application following user requirements (so called \emph{lean implementation}).
 
@@
 All the biological processes featured by the GO are looked-up one by one. All the 
-time a process is linked to another process (X) via a positive or negative regulation link, two FTC 
-classes are created: Anti-X agent and Pro-X agent. For instance the GO term positive regulation of blood 
-coagulation is linked to the term blood coagulation via a positively regulates relation, therefore two FTC 
-categories Anti-blood coagulation agent and Pro-blood coagulation agent are created. The identifiers of the 
+time a process is linked to another process (\emph{X}) via a \emph{positive} or \emph{negative regulation} link, two FTC 
+classes are created: \emph{Anti-X agent} and \emph{Pro-X agent}. For instance the GO term \emph{positive regulation of blood 
+coagulation} is linked to the term \emph{blood coagulation} via a \emph{positively regulates} relation, therefore two FTC 
+categories \emph{Anti-blood coagulation agent} and \emph{Pro-blood coagulation agent} are created. The identifiers of the 
 new FTC classes are also derived from the GO term used to create the class pattern. The GO numeric identifier 
-is re-used and the letter A or P is appended before to emphasize the anti or pro pattern. From the example presented 
-previously, the FTC class Anti-blood coagulation has FTC\_A0007596 as identifier, because the GO term blood coagulation is 
-referenced by GO:0007596. Following the same logic, FTC\_P0007596 is the identifier of the class Pro-blood coagulation. The 
+is re-used and the letter \emph{A} or \emph{P} is appended before to emphasize the \emph{anti} or \emph{pro} pattern. 
+From the example presented 
+previously, the FTC class \emph{Anti-blood coagulation} has FTC\_A0007596 as identifier, because the GO term \emph{blood coagulation} is 
+referenced by GO:0007596. Following the same logic, FTC\_P0007596 is the identifier of the class \emph{Pro-blood coagulation agent}. The 
 design choice for identifiers and labels allows the FTC to fully rely on the high quality work provided by the GO curation 
@@
 The mode and mechanism of actions related to molecular functions are created in the following manner: All the time 
-a molecular function (Y) is encountered then two FTC categories are created, as for the processes: Anti-Y agent and 
-Pro-Y agent. The identifiers are assigned the same way as described before. For instance, out of the GO term androgen 
-receptor activity (GO:0004882) two FTC classes are derived: Pro-androgen receptor activity agent (FTC\_P0004882) and 
-Anti-androgen receptor activity agent (FTC\_A0004882).
+a molecular function (\emph{Y}) is encountered then two FTC categories are created, as for the processes: \emph{Anti-Y agent} and 
+\emph{Pro-Y agent}. The identifiers are assigned the same way as described before. For instance, out of the GO term \emph{androgen 
+receptor activity} (GO:0004882) two FTC classes are derived: \emph{Pro-androgen receptor activity agent} (FTC\_P0004882) and 
+\emph{nti-androgen receptor activity agent} (FTC\_A0004882).
 
@@
 FTC classes are generated as presented in the previous section. Up to this point, these categories 
-are only individual tokens with a human readable label as well as an identifier. The next step is going to 
+are only tokens with a human readable label as well as an identifier. The next step is going to 
 assign equivalent definitions to each FTC class. A reasoner can understand such definitions and will automatically 
 classify the knowledge base accordingly; drugs will then be assigned to FTC categories and the taxonomic structure 
-arises after the reasoning step. Equivalent definitions are written as OWL class expressions using the entities of 
-the knowledge base. The core classes and properties are presented in the supplementary material, this section assumes 
-that the reader is already familiar with the FTC specifications (summarised at {{https://github.com/loopasam/ftc/wiki/Knowledge-Base}}). 
+arises after this reasoning step. Equivalent definitions are written as OWL class expressions using the entities of 
+the knowledge base (summarised at {{https://github.com/loopasam/ftc/wiki/Knowledge-Base}} and in section ).
+
+//TODO refer to section
+
 There are two types of equivalences: The first one captures perturbation of regulatory 

### 2013-08-19 document.tex
supplementary polished and good
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 
-\subsection{Categories related to biological processes}
+\subsection{Categories related to biological processes}\label{catprocess}
 All the biological processes featured by the GO are looked-up one by one. All the 
@@
 
-\subsection{Categories related to molecular functions}
+\subsection{Categories related to molecular functions}\label{catfunc}
 The mode and mechanism of actions related to molecular functions are created in the following manner: All the time 
@@
 receptor activity} (GO:0004882) two FTC classes are derived: \emph{Pro-androgen receptor activity agent} (FTC\_P0004882) and 
-\emph{nti-androgen receptor activity agent} (FTC\_A0004882).
+\emph{Anti-androgen receptor activity agent} (FTC\_A0004882).
 
@@
 arises after this reasoning step. Equivalent definitions are written as OWL class expressions using the entities of 
-the knowledge base (summarised at {{https://github.com/loopasam/ftc/wiki/Knowledge-Base}} and in section ).
-
-//TODO refer to section
-
+the knowledge base (summarised at {{https://github.com/loopasam/ftc/wiki/Knowledge-Base}} and in section \ref{specs}).
 There are two types of equivalences: The first one captures perturbation of regulatory 
-biological processes (so called regulatory patterns) and the second one handles the perturbed functions (functional patterns).
+biological processes (so called \emph{regulatory patterns}) and the second one handles the perturbed functions (\emph{functional patterns}).
 
@@
 Some of the FTC categories are created from the biological processes present in the 
-GO (cf previous sections); these categories have two arbitrary equivalent definitions, representing the 
+GO (cf section \ref{catprocess}); these categories have two arbitrary equivalent definitions, representing the 
 two possible ways a compound might impact the biological process. Anti-biological process agent FTC categories 
 contain the drugs that negatively perturb a target involved in the positive regulation of the biological process. 
-The anti categories also feature the compounds that positively perturb a negative regulator of the same process. 
-The pro categories are equivalent to the opposite pattern. Figure 5 illustrates the equivalent definitions for the 
-FTC class Anti-blood coagulation agent (FTC\_A0007596).
+The \emph{anti} categories also feature the compounds that positively perturb a negative regulator of the same process. 
+The \emph{pro} categories are equivalent to the opposite pattern. Supplementary Figure \ref{fig:S01} illustrates the equivalent definitions for the 
+FTC class \emph{Anti-blood coagulation agent} (FTC\_A0007596).
 
@@
 \centerline{\includegraphics{figS1.png}}
-\caption{Example of equivalent definitions (gray boxes) for the FTC category Anti-blood coagulation 
-agent (FTC\_A0007596). After the data integration step, a reasoner will look the definitions and identify 
-which drugs comply with the FTC definitions. Definitions are expressed using the Web Ontology Language (OWL), 
+\caption{Example of equivalent definitions (gray boxes) for the FTC category \emph{Anti-blood coagulation 
+agent} (FTC\_A0007596). After the data integration step, a reasoner will look the definitions and identify 
+which drugs satisfy the FTC definitions. Definitions are expressed using the Web Ontology Language (OWL), 
 serialised here using the Manchester syntax.}\label{fig:S01}
@@
 \subsection{Functional pattern}
-The FTC categories generated from the GO molecular functions are also equivalent to a logical 
-definition. Anti FTC categories dealing with molecular activities are asserted as equals to the drugs 
-that negatively perturb the function. Pro categories are equivalent to the drugs that positively perturb the 
+The FTC categories generated from the GO molecular functions (cf section \ref{catfunc}) are also equivalent to a logical 
+definition. \emph{Anti} FTC categories dealing with molecular activities are asserted as equals to the drugs 
+that negatively perturb the function. \emph{Pro} categories are equivalent to the drugs that positively perturb the 
 function of interest. A summary of the patterns definitions is available on the 
@@
 further populated with some information coming from various public databases. Only manually curated 
-information is considered. This section illustrates how the core entities interact with the different data types 
-in order to create the axioms.
+information is considered.
 
@@
 The protein targets all have at least one manually asserted GO annotation \citep{Dimmer2012} for a biological process or a 
-molecular function. DrugBank links compounds to targets via actions. The DrugBank actions are somehow 
-structured and consistent: Concepts such as inhibitor or agonist are reused throughout the database for 
+molecular function. DrugBank links compounds to targets via \emph{actions}. The DrugBank actions are somehow 
+structured and consistent: Concepts such as \emph{inhibitor} or \emph{agonist} are reused throughout the database for 
 example, yet they are not strictly formalised as a controlled vocabulary. These actions are manually 
-standardised to the core properties of the FTC according to their meaning: For instance the action 
-antagonist is mapped to the FTC negatively-perturbs.
+standardised to the core properties of the FTC according to their biochemical meaning: For instance the action 
+\emph{antagonist} is mapped to the FTC \emph{negatively-perturbs} property.
 Compounds coming from DrugBank are represented as OWL classes and asserted 
-as subclasses of DrugBank compound (FTC\_C2). Protein targets are described as classes 
-too and subclasses of the core class protein. Each DrugBank compound is then connected to 
-its target via the following axiom pattern: \emph{drug SubClassOf perturbs some protein}. 
+as subclasses of the class \emph{DrugBank compound} (FTC\_C2). Protein targets are described as OWL classes 
+too and subclasses of the core class \emph{Protein}. Each DrugBank compound is then connected to 
+its target via the following axiom pattern: \emph{[drug] SubClassOf perturbs some [protein]}. 
 E.g. \emph{Ximelagatran SubClassOf negatively-perturbs some Prothrombin}.
@@
 The GO annotation program aims to provide high-quality GO annotations to 
-proteins in UniProt. In the context of the FTC, such annotations are used 
+proteins in UniProt \citep{Dimmer2012}. In the context of the FTC, such annotations are used 
 to create axioms linking protein targets to molecular functions and biological processes. 
 Each protein annotated with a function creates an axiom such 
-as protein SubClassOf has-function some molecular function. 
+as \emph{[protein] SubClassOf has-function some [molecular function]}. 
 Each protein annotated to a biological process creates an axiom such 
-as protein SubClassOf involved-in some biological process. 
-E.g. Prothrombin SubClassOf involved-in some positive regulation of blood coagulation. 
-Each protein can be involved in multiple processes and capable of realizing multiple functions.
+as \emph{[protein] SubClassOf involved-in some [biological process]}. 
+E.g. \emph{Prothrombin SubClassOf involved-in some positive regulation of blood coagulation}. 
+Each protein can be involved in multiple processes and capable of realizing multiple functions;
+the polypharmacology is captured at this level.
 
@@
 the different data together and to explicitly express the logical links between resources. The FTC 
-knowledge base follows an OWL2 EL profile, which enable the use of fast and parallelised reasoners such 
-as ELK (cite). During the classification process, the reasoner checks whether the mode of action equivalent 
+knowledge base follows an OWL2 EL profile \citep{Motik2009}, which enable the use of fast and parallelised reasoners such 
+as ELK. During the classification process, the reasoner checks whether the mode of action equivalent 
 definitions are satisfied or not and assigns drugs inside the corresponding FTC categories. The tree structure 
@@
 more classes from the ATC. The idea is to look at the set of drugs contained in both side of the 
-equivalence and estimate the overlap. This is illustrated in the Figure 6. Evaluation points are defined 
+equivalence and estimate the overlap, as illustrated in the Supplementary Figure \ref{fig:S02}. Evaluation points are defined 
 by hand and not themselves evaluated. The full list of evaluation points as well as a summary of the results 
@@
 \section{Mode of action similarity against indication}
-A statistical analysis was performed over the data presented on Figure 3. 
+A statistical analysis was performed over the data presented on Figure 4. 
 When two compounds are randomly taken, they have on average a higher mode of action similarity when 
-they are assigned to the same ATC category (one ATC level). In order to estimate whether this observation was 
-due to chance only, we formulated the following null hypothesis (H0): For a pair of drug A and B, it does not 
-matter to which ATC category they belong to, the average similarity is always the same. 
-The alternative hypothesis (H1) was: For a pair of drug A and B, if A and B have the same ATC code, 
-we expect on average to obtain a higher similarity value than if A and B have different ATC codes. 
+they are assigned to the same ATC category (one ATC level).
+In order to estimate whether this observation was 
+due to chance only, we formulated the following null hypothesis (H0): \emph{For a pair of drug A and B, it does not 
+matter to which ATC category they belong to, their similarity is always average.}
+The alternative hypothesis (H1) was: \emph{For a pair of drug A and B, if A and B have the same ATC code, 
+we expect on average to obtain a higher similarity value than if A and B have different ATC codes.}
+
 A permutation test was then performed for each ATC category. For example, we started with the ATC 
-category A (first row on Figure 3), looked at the similarity values when pairs of compounds both belong 
+category A (first row on Figure 4), looked at the similarity values when pairs of compounds both belong 
 to the category A (top right corner square) and compared it to the similarity values when the pair of compounds 
@@
 performed in order to see whether this observation was due to chance only. We were able to reject the null 
-hypothesis for a significance level of 0.05 all the times. As the test is performed multiple times, 
-we applied a Bonferroni correction for multiple hypothesis testing, which changed the significance level accordingly. 
+hypothesis for a significance level of 0.05 all the times. 
 The choice for a permutation test was driven by the fact that MoA similarity values do not follow any type 
@@
 
-\section{Knowledge base specification}
+\section{Knowledge base specification}\label{specs}
 This section presents the scaffold of the knowledge base underlying the FTC. 
@@
 Label: drug \\
-Definition: As defined by schema.org. A chemical or biologic substance, used as a medical therapy, that has a 
+Definition: As defined by schema.org: A chemical or biologic substance, used as a medical therapy, that has a 
 physiological effect on an organism. \\
@@
 Identifier: http://purl.obolibrary.org/obo/RO\_0002211 \\
-Chained property: regulates o part-of -> regulates \\
+Chained property: regulates o part-of -\textgreater regulates \\
 Label: regulates \\
@@
 
-\section*{Pairwise comparison of mode of action similarities}
-
-Supplementary Figure S1: Pairwise comparison of mode of action similarities. 
-The compounds are grouped by ATC code (2 levels - legend not shown) on panel A. 
-A hierarchical clustering step (based on manhattan distance) was applied on panel B. 
-The dendrogram reflects the original structure of the FTC. 
-Panel C shows the compounds sorted by DrugBank identifiers only (≈ random). 
-The colours displayed on panel B and C correspond to one level ATC categories and are indexed in the legend panel.
+\begin{figure*}[!tpb]%figureS3
+\centerline{\includegraphics{figS3.png}}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but with two ATC levels considered. It is possible to observe some clustering around the diagonal line (no
+statistical analysis performed).}\label{fig:S03}
+\end{figure*}
 
 \begin{figure*}[!tpb]%figureS3
-\centerline{\includegraphics{figS3.png}}
-\caption{Caption to be written.}\label{fig:S03}
-\end{figure*}
-
-\begin{figure*}[!tpb]%figureS4
 \centerline{\includegraphics{figS4.png}}
-\caption{Caption to be written.}\label{fig:S04}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but with a hierarchical clustering applied on the top of it, based on the manhattan distance.
+The dendrogram reflects the original structure of the FTC: Functionaly similar compounds being grouped toghether.
+This figure can be used for further analysis over functional clusters of drugs.}\label{fig:S04}
 \end{figure*}
@@
 \centerline{\includegraphics{figS5.png}}
-\caption{Caption to be written.}\label{fig:S05}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but the compounds have been sorted horizontally and vertically only based on their identifiers, which
+we considered as equivalent to a pseudo-random sorting. In this case
+no visual pattern is observable, the similarity values appear homogeneously distributed (no statistical analysis performed).}\label{fig:S05}
 \end{figure*}

### 2013-08-19 document.tex
First sections have been final proofed
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-08-18 document.tex
@@
 into consideration gene expression signatures or chemical features for instance. 
-We present here a novel method based on knowledge integration using semantic technologies. 
+We present here a novel method based on knowledge integration using semantic technologies, 
+to capture the functional role of approuved chemical compounds.
+\section{Results:}
 In order to computationally generate repurposing hypotheses, we used the Web Ontology Language (OWL) to 
-mathematically define the meaning of over 20'000 terms formally describing various modes and mechanisms of action.
-
-\section{Results:}
-Based on an integration of public data, we have automatically assigned over a thousand of approved drugs into 
+mathematically define the meaning of over 20'000 terms formally describing various modes and mechanisms of action. Based on an 
+integration of public data, we have automatically assigned over a thousand of approved drugs into 
 these mode of action categories. The resulting new resource is called the Functional Therapeutic Chemical Classification System (FTC) 
@@
 Accurately identifying these different functions helps to predict the potential side-effects a drug could have and can also lead to 
-interesting repurposing opportunities \citep{Medina-Franco2013}. For instance, sildenafil was initially developed 
+interesting repurposing opportunities \citep{Medina-Franco2013}. For instance, \emph{sildenafil} was initially developed 
 against angina and has been repurposed towards erectile dysfunction during the clinical trials \citep{Ashburn2004}. Approved compounds are 
@@
 have in common to look for similarities within existing drugs and forward similar compounds as repurposing hypotheses.
-A feature of particular interest to describe drugs is the mode of action (MoA). According to Wikipedia, 
-the MoA describes a functional or anatomical change, 
-at the cellular level, resulting from the exposure of a living organism to a substance. For instance 
-terms such as transcriptional regulation agent or anticoagulant define MoAs and characterise the roles of a certain type of drugs. The 
+A feature of particular interest to describe drugs is the Mode of Action (MoA). According to Wikipedia, 
+the MoA describes \emph{a functional or anatomical change, 
+at the cellular level, resulting from the exposure of a living organism to a substance}. For instance 
+terms such as \emph{transcriptional regulation agent} or \emph{anticoagulant} define MoAs and characterise the 
+roles of a certain type of drugs. The 
 MoA abstracts over the relations between molecular functions, protein targets and drug activities; it is the central concept linking a 
@@
 In an ontology or knowledge base, concepts are organised and linked following the logical type of relation they have between them. 
-In the Gene Ontology (GO) for example \citep{Ashburner2000}, biological processes and molecular functions are manually curated in this resource 
-and the meaning of terms is specified by the relations linking two GO terms.
+In the Gene Ontology (GO) for example \citep{Ashburner2000}, biological processes and molecular functions terms are manually curated  
+and their meaning is specified by the relation types linking two GO terms.
 MoA definitions are present in other classifications such as the Medical Subject Headings \citep{Nelson2004} or the Chemical Entities of 
@@
 extent the action of drugs at the anatomical level. All these resources are valuable for the community as a source of carefully 
-manually curated information. Moreover, the categories described in these classification systems are sometimes used to annotate 
-drugs: For instance the compound sildenafil has been manually annotated as vasodilator agent (CHEBI:35620 or MeSH:D27.505.954.411.918).
+manually curated information; moreover, the categories described in these classification systems are sometimes used to annotate 
+drugs: For instance the compound \emph{sildenafil} has been manually annotated as \emph{vasodilator agent} (CHEBI:35620 or MeSH:D27.505.954.411.918).
 The classifications mentioned previously are not specially designed for drug repurposing; they purposefully report only the 
@@
 in the resource and further populated with approved drugs using the Web Ontology Language (OWL) in combination with a reasoner. 
-Most of the drugs are present in multiple FTC categories, reflecting the various roles a compound can play inside a biological system. 
+Most of the drugs are present in multiple FTC categories, reflecting the various roles a compound can play inside a biological system which 
+can serve as starting point for repurposing. 
 The resource was evaluated against the ATC, traditional classification scheme introduced before. We present as well some preliminary 
@@
 The GO terms serve as template to create the FTC categories describing the MoAs; DrugBank \citep{Knox2011} provides the 
-known links between drugs and their protein targets. Uniprot \citep{TheUniprotConsortium2013} maps targets to their 
+known links between drugs and their protein targets and Uniprot \citep{TheUniprotConsortium2013} maps targets to their 
 respective GO annotations \citep{Dimmer2012}. 
-Drugs are further assigned into mode of action categories based of the OWL constructs and axioms defined in the 
-FTC (see Availability and Implementation). A reasoner, a program capable of understanding such axioms, performed this task. 
+Drugs are further assigned into MoA categories based of the OWL constructs and axioms defined in the 
+FTC (see \emph{Availability and Implementation}). A reasoner, a program capable of understanding such axioms, performed this task. 
 The process to build the FTC is summarised in Figure~\ref{fig:01}.
-
-\begin{figure*}[!tpb]%figure1
-\centerline{\includegraphics{fig1.png}}
-\caption{Parent categories to the FTC class Pro-fibrinolysis agent (FTC\_P0042730). The classification is a direct 
-acyclic graph where categories are describing increasingly specific concepts. Arrows entails subclass relationships 
-between the terms (is a relation).}\label{fig:01}
-\end{figure*}
-
 It takes around 5 seconds on a modern 
@@
 The reader can browse and access the content of the FTC online at {{https://www.ebi.ac.uk/chembl/ftc}}.
-
-\begin{figure}[!tpb]%figure2
-\centerline{\includegraphics{fig2.png}}
-\caption{Parent categories to the FTC class Pro-fibrinolysis agent (FTC\_P0042730). The classification is a direct 
-acyclic graph where categories are describing increasingly specific concepts. Arrows entails subclass relationships 
-between the terms (is a relation).}\label{fig:02}
-\end{figure}
-
 In total there are 1280 approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets. 
@@
 when direct and 
-indirect drugs are considered. FTC categories not containing drugs (e.g, FTC\_A0001771 - Anti-immunological synapse 
-formation agent) represent modes and mechanisms of action for which no approved compounds exist yet or that have not been identified 
+indirect drugs are considered. FTC categories not containing drugs (e.g, FTC\_A0001771 - \emph{Anti-immunological synapse 
+formation agent}) represent modes and mechanisms of action for which no approved compounds exist yet or that have not been identified 
 as such in the FTC.
+
+\begin{figure*}[!tpb]%figure1
+\centerline{\includegraphics{fig1.png}}
+\caption{Summary of the methodology used to construct the resource. The FTC categories are first created (A) from the existing GO terms
+describing molecular functions and biological processes. Then, each category has an equivalent definition assign to it (B).
+The content of various databases in incorporated (C) and linked using the FTC specific properties. Finally a reasoner runs over the knowledge
+base and assigns drugs to MoA categories based on whether or not a definition can be satisfied.}\label{fig:01}
+\end{figure*}
+
+\begin{figure}[!tpb]%figure2
+\centerline{\includegraphics{fig2.png}}
+\caption{Parent categories to the FTC class \emph{Pro-fibrinolysis agent} (FTC\_P0042730). The classification is a direct 
+acyclic graph where categories are describing increasingly specific concepts. Arrows entails subclass relationships 
+between the terms (\emph{is a} relation).}\label{fig:02}
+\end{figure}
 
@@
 
-\section{Availability and Implementation}
+\section{Availability and Implementation}\label{impl}
 The code behind the creation of the resource is entirely open and available 

### 2013-08-19 document.tex
caption figure 1
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 \centerline{\includegraphics{fig1.png}}
-\caption{Summary of the methodology used to construct the resource. The FTC categories are first created (A) from the existing GO terms
-describing molecular functions and biological processes. Then, each category has an equivalent definition assign to it (B).
-The content of various databases in incorporated (C) and linked using the FTC specific properties. Finally a reasoner runs over the knowledge
+\caption{Summary of the methodology used to construct the resource. (A) The FTC categories are first created from the existing GO terms
+describing molecular functions and biological processes. (B) Then, each category has an equivalent definition assign to it.
+(C) The content of various databases in incorporated and linked using the FTC specific properties. (D) Finally a reasoner runs over the knowledge
 base and assigns drugs to MoA categories based on whether or not a definition can be satisfied.}\label{fig:01}

### 2013-08-20 document.tex
evaluation proofed
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 \caption{Summary of the methodology used to construct the resource. (A) The FTC categories are first created from the existing GO terms
-describing molecular functions and biological processes. (B) Then, each category has an equivalent definition assign to it.
+describing molecular functions and biological processes. (B) Then, each category has a logical equivalent definition assign to it.
 (C) The content of various databases in incorporated and linked using the FTC specific properties. (D) Finally a reasoner runs over the knowledge
-base and assigns drugs to MoA categories based on whether or not a definition can be satisfied.}\label{fig:01}
+base and assigns drugs to MoA categories based on whether or not a definition can be satisfied. The taxonomic structure of the
+FTC appears also in the reasoning step, from the entailements of the equivalent defintions.}\label{fig:01}
 \end{figure*}
@@
 has been assessed. In practice, these two resources have different goals; therefore the evaluation should be 
-interpreted accordingly. The full methodology behind the evaluation is described in the section “Evaluation methodology”. 
-Briefly, 68 classes present in the FTC have an equivalent meaning to a combination of classes present in the ATC. 
-This semantic equivalence was manually asserted and called evaluation point (see “Evaluation points” from Availability and Implementation). 
-The set of drugs present in both classes for each evaluation points were then compared and some metrics are derived 
-from this comparison. For example, the FTC category Anti-hydrogen:potassium-exchanging ATPase activity agent (FTC\_A0008900) has 
-been manually asserted as equivalent to the ATC category Proton pump inhibitors (A02BC). A summary of this 
+interpreted accordingly. The full methodology behind the evaluation is 
+described in the section 6 of the supplementary material (\emph{Evaluation methodology}). 
+Briefly, 68 categories present in the FTC have an equivalent meaning to a combination of categories present in the ATC. 
+This semantic equivalence was manually asserted 
+and called \emph{evaluation point} (see supplementary material section 6.1 for details). 
+The set of drugs present in both classes for each evaluation points were then compared and some metrics derived. 
+For example, the FTC category \emph{Anti-hydrogen:potassium-exchanging ATPase activity agent} (FTC\_A0008900) has 
+been manually asserted as equivalent to the ATC category \emph{Proton pump inhibitors} (A02BC). A summary of this 
 evaluation point is furthermore available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/FTC\_A0008900}}.
 Out of the 1280 DrugBank compounds present in the FTC, 1134 are also present in the ATC, therefore only those were 
-considered. The evaluation points cover a total of 471 DrugBank compounds. Out of these, 277 compounds are true positives, 
-meaning that given an evaluation point, they are present in both FTC and ATC categories. The proton pump inhibitor 
-evaluation point is such an example where all the drugs (omeprazole, esomeprazole, pantoprazole, lansoprazole, rabeprazole) are 
-present both in the FTC category and in the corresponding ATC class. The total number of compounds presents in an ATC class but 
+considered. The \emph{evaluation points} cover a total of 471 DrugBank compounds or around 41\% of common drugs to both classifications.
+Out of these, 277 compounds are true positives, 
+meaning that given an evaluation point, they are present in both FTC and ATC categories. The \emph{proton pump inhibitor} 
+evaluation point is such an example where all the drugs (\emph{omeprazole, esomeprazole, pantoprazole, lansoprazole, rabeprazole}) are 
+present both in the FTC category and in the corresponding ATC class(es). The total number of compounds present in an ATC class but 
 not in the corresponding FTC category (false negatives) is 36. Finally, 279 compounds are present in a FTC class but not in the 
-corresponding ATC class; such agents are the false positives. From these values, it is possible to derive a 
+corresponding ATC class(es); such agents are the false positives. From these values, it is possible to derive a 
 recall of 88\%; this percentage indicates that the automatic build of the FTC covers a good portion of the content 
 already present in the ATC. The precision of 50\% shows that the FTC contains for a given MoA many more drugs than the 
-equivalent ATC categories.
+equivalent ATC categories. This result was expected and comes the original idea behind the FTC: Representing in a systematic fashion
+the implicit and explicit MoAs of drugs, in particular the ones not already indexed by current classification scheme.
 

### 2013-08-20 document.tex
polish figure 3
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 categories are related to biological processes and 19'064 to the molecular functions. A summary of the metrics behind the latest 
-build is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}. Out of all FTC classes, 1432 
+build is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.
+Out of all FTC classes, 1432 
 categories (\textgreater 6\%) directly contain at least one approved drug. This number increases up to 2532 (\textgreater 11\%) 
@@
 \subsection{Exploration}
-The FTC was designed to assist drug repurposing analyses by explicitly representing the poly-pharmacology of 
+The FTC was designed to assist drug repurposing analyses by explicitly representing the polypharmacology of 
 approved drugs. In this section we exemplify how the resource can be used to perform different types of analysis.
@@
 exist to re-orient a drug into doing something new. The therapeutic agents described in the FTC can have several MoAs, 
-representing the intrinsic polypharmacology of the approved compounds. Figure 2 illustrates the poly-pharmacology 
-spectrum by showing the distribution of number of MoAs per compound. This number ranges from one to more 
-than eighty for certain drugs, when indirect parents are considered. Not all the MoAs are biologically relevant, 
-some FTC categories are particularly abstract (e.g, Anti-biological process agent) yet they represent discrete 
+representing the intrinsic polypharmacology of the approved compounds. Figure \ref{fig:03} illustrates the polypharmacology 
+spectrum by showing the distribution of number of MoAs per compound. When only direct categories are considered, compounds belong on average
+to 13.5 MoA categories. This number increases to 61.2 when parent categories are taken into consideration (super classes).
+Not all the MoAs are disease relevant, 
+some FTC categories are particularly abstract (e.g, \emph{Anti-biological process agent}) yet they represent discrete 
 categories to which the drug belongs with an explicit and clear meaning. These discrete MoAs are a good starting 
-point to understand what a compound can do when administered in a human system. The human readable definitions 
+point to understand what a compound can do when administered in a human system, as it will be shown with Alzheimer's disease.
+The human readable definitions 
 associated to each FTC category furthermore help the user to choose the adequate MoA for the desired outcome.
+According to these results, compound's polypharmacology is 
+well represented in the FTC, as shown by the numerous MoAs each approuved drug can exhibit.
 
@@
 \centerline{\includegraphics{fig3.png}}
-\caption{Distribution of the direct and indirect number of MoAs per drug. 
-Means are indicated with a solid line. When considering indirect categories, the distribution range is wider.}\label{fig:03}
+\caption{Distribution of the direct (blue) and indirect (red) number of MoAs per drug.
+Means are indicated with a solid line. On average each compound has 13.5 MoAs when only direct classes are considered. The number rises
+to 61.2 when indirect MoAs are included. Indirect MoAs are the ancestor classes in the taxonomy as shown in Figure \ref{fig:02}.
+The ditribution range is wider (number) when indirect MoAs are considered, emphasizing the fact that some drugs
+are well characterised in databases and could be used for a variety of specific biological tasks. Finally some compounds have been assigned
+to a small number of FTC categories; in such cases little is known or reported about their pharmacology and repurposing opportunities
+might be limited.}\label{fig:03}
 \end{figure*}

### 2013-08-20 document.tex
some more polishing
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 integration of public data, we have automatically assigned over a thousand of approved drugs into 
-these mode of action categories. The resulting new resource is called the Functional Therapeutic Chemical Classification System (FTC) 
+these MoA categories. The resulting new resource is called the Functional Therapeutic Chemical Classification System (FTC) 
 and was further evaluated against the content of the traditional Anatomical Therapeutic Chemical Classification System (ATC). 
@@
 to 61.2 when indirect MoAs are included. Indirect MoAs are the ancestor classes in the taxonomy as shown in Figure \ref{fig:02}.
-The ditribution range is wider (number) when indirect MoAs are considered, emphasizing the fact that some drugs
+The distribution range is wider when indirect MoAs are considered (range=299; min=7; max=306) versus direct MoAs only (range=79; min=3; max=82).
+These results emphasises the fact that some drugs
 are well characterised in databases and could be used for a variety of specific biological tasks. Finally some compounds have been assigned
@@
 
-\subsubsection{Drugs with similar mode of actions have similar indications}
-The list of MoAs attributed to a drug can be exploited as a descriptor for the therapeutic agent. 
+\subsubsection{Drugs with similar MoAs have similar indications}
+The list of MoAs attributed to a drug can be exploited as a descriptor for the therapeutic agent:
 The tree structure of the FTC can be used to derive some similarity metrics over the MoAs. The underlying 
-heuristic is to assume that the closer two entities are in a taxonomy or ontology, the more similar they are. 
-We used a straightforward approach derived from the Jaccard index (see method for more details) in order to compare 
+heuristic is to assume that the closer two entities are in the taxonomy, the more similar they are. 
+We used a straightforward approach derived from the Jaccard index (see supplementary section 7) in order to compare 
 approved drugs based on the similarity of their MoAs. For instance, the similarity between two compounds present in 
-the same FTC category is 1 (maximum). The similarity between an anti-blood coagulant and pro-blood coagulant is 0.29, 
-reflecting the fact that such compounds are dissimilar in regards to the outcome of their biological effect. As the MoA is 
+the same FTC category is 1 (maximum). The similarity between an \emph{anti-blood coagulant} and \emph{pro-blood coagulant} is 0.29, 
+reflecting the fact that such compounds are not similar in regards to the outcome of their biological effect. As the MoA is 
 intuitively expected to be the central concept leading to the indication of the drug, we expected that on average, drugs with 
-similar MoAs would be indicated towards similar therapeutic areas. The heat map presented in Figure 3 shows a pairwise 
+similar MoAs would be indicated towards similar therapeutic areas. The heat map presented in Figure \ref{fig:04} shows a pairwise 
 comparison of all the drugs of the FTC based on their relative MoA similarity. The compounds are further grouped by 
 therapeutic indications as defined by the ATC.
-
-\begin{figure*}[!tpb]%figure4
-\centerline{\includegraphics{fig4.png}}
-\caption{Pairwise comparison of mode of action similarities. Only the first level of 
-the ATC is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
-The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black). Drugs are grouped according 
-to their first ATC level (colours on the side). For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
-which appears as “B” (dark orange) on the plot. The average similarity of drugs present in the same therapeutic category 
-is significantly higher on average when separately compared to all other indications.}\label{fig:04}
-\end{figure*}
-
-The heat map displayed in Figure 3 reveals some square patches around the central diagonal. 
-The overall similarity appears higher when compounds from the same ATC group are considered. 
-A significance analysis (cf details covered in sections “Semantic similarity” 
-and “Mode of action similarity against indication” from Availability and Implementation) 
+The heatmap (Figure \ref{fig:04}) reveals some square patches around the central diagonal;
+the overall similarity appears higher when compounds from the same ATC group are considered. 
+A significance analysis (details covered in supplementary section 8)
 revealed that the average MoA similarity of compounds belonging to the same ATC category is 
@@
 different therapeutic areas and are used for different purposes according to the ATC. We are currently further analysing 
-such cases in a systematic fashion. Supplementary Figure S1A present similar association behavior when two levels of the ATC are 
-considered (no statistical significance performed). Supplementary Figure S1B re-uses the same data as Figure 1 (one ATC level) 
+such cases in a systematic fashion. Supplementary Figure 3 present similar association behavior when two levels of the ATC are 
+considered (no statistical significance performed). Supplementary Figure 4 re-uses the same data as Figure \ref{fig:04} (one ATC level) 
 but with a clustering function apply to it (hierarchical clustering - manhattan distance) in order to reveal 
-functional clusters of drugs. Finally, Figure S1C shows the distribution when compounds are sorted based on 
+functional clusters of drugs. Finally, Supplementary Figure 5 shows the distribution when compounds are sorted based on 
 their identifiers; no patterns are identifiable in this case. Taken together, these results emphasize that the MoAs as 
-defined in the FTC are indeed on average associated with the therapeutic indication of a drug.
+defined in the FTC are indeed on average associated with the therapeutic indication of a drug. This result supports the validity of the
+resource and its potential to computationally address indication discovery.
 
-\subsection{Drug repurposing hypotheses for Alzheimer’s disease}
+\begin{figure*}[!tpb]%figure4
+\centerline{\includegraphics{fig4.png}}
+\caption{Pairwise comparison of MoAs similarities. Only the first level of 
+the ATC is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
+The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black). Drugs are grouped according 
+to their first ATC level (colours on the side). For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
+which appears as ``B'' (dark orange) on the plot.
+Some compounds belong to multiple ATC categories (Mulitple - black) and some others do not have
+an ATC code (NoCategory - white)
+The average similarity of drugs present in the same therapeutic category 
+is significantly higher on average when separately compared to all other indications.}\label{fig:04}
+\end{figure*}
+
+\subsection{Drug repurposing hypotheses for Alzheimer's disease}
 In this section we provide examples how actual drug repurposing hypotheses can be derived from the FTC. 
 The approach presented here makes use of the FTC categories as analogs to compartments of a toolbox helping to 
-find drugs to treat Alzheimer’s disease. Five FTC categories containing drugs are directly related to the biological 
-processes of the neurodegenerative condition: Anti-amyloid precursor protein biosynthetic process agent (FTC\_A0042983), 
-Anti-tau protein kinase activity agent (FTC\_A0050321), Anti-tau protein binding agent (FTC\_A0048156), Anti-beta amyloid 
-binding agent (FTC\_A0001540) and Pro-beta amyloid binding agent (FTC\_P0001540). We have then considered the drugs present 
-inside each of these classes as potential candidates. Figure 4 shows these drugs, which have been further manually grouped 
-based on the overall similarity of their actions (letters on Figure 4).
+find drugs to treat Alzheimer's disease. Five FTC categories containing drugs are directly related to the biological 
+processes of the neurodegenerative condition: \emph{Anti-amyloid precursor protein biosynthetic process agent} (FTC\_A0042983), 
+\emph{Anti-tau protein kinase activity agent} (FTC\_A0050321), \emph{Anti-tau protein binding agent} (FTC\_A0048156), 
+\emph{Anti-beta amyloid 
+binding agent} (FTC\_A0001540) and finally \emph{Pro-beta amyloid binding agent} (FTC\_P0001540). We have then considered the drugs present 
+inside each of these classes as potential candidates. Figure \ref{fig:05} shows these drugs, which have been further manually grouped 
+based on the overall similarity of their actions (letters on Figure \ref{fig:05}).
 
@@
 \centerline{\includegraphics{fig5.png}}
-\caption{FTC categories (letters A, F, H, I and J) describing some of the modes and mechanisms of 
-action that could impact Alzheimer’s disease. Drugs classified in these FTC categories are listed and further 
-manually grouped based on their mode of action similarities (subgroups B, C, D, E and G).}\label{fig:05}
+\caption{FTC categories describing some of the modes and mechanisms of 
+action that could impact Alzheimer's disease. The categories have been manually picked on the basis that they could directly affect
+the dementia. Drugs classified in these FTC categories further 
+manually grouped based on their MoAs similarities (subgroups B, C, D, E and G).}\label{fig:05}
 \end{figure}
 
-The subgroups B, C and D are inhibitors of the cholinergic system and some of them, such as galantamine (DB00674), are already 
-investigated to treat Alzheimer’s disease and other related dementias. This class of agent is in line with the 
-cholinergic hypothesis \citep{Francis1999}, stating that Alzheimer’s disease could be caused by dysfunctions in the processing of 
+The subgroups B, C and D are inhibitors of the cholinergic system and some of them, such as \emph{galantamine} (DB00674), are already 
+investigated to treat Alzheimer's disease and other related dementias. This class of agent is in line with the 
+cholinergic hypothesis \citep{Francis1999}, stating that Alzheimer's disease could be caused by dysfunctions in the processing of the
 acetylcholine. The subgroup E is exclusively composed of barbiturates (central nervous system depressants). 
-The presence of this pharmacological class of compounds as an Alzheimer treatment is more surprising, as very little 
+The presence of this pharmacological class of compounds as an Alzheimer's disease treatment is more surprising, as very little 
 literature reports on it. Further investigations reveal that the neuronal acetylcholine receptor subunit alpha-7, a 
 common off-target of barbiturates, binds beta-amyloids with high affinity \citep{Wang2000}. As beta-amyloids are themselves strongly involved 
-in the pathology, barbiturates could affect the state of the condition \citep{Wang2000}, similarly to cholinergic inhibitors.
-The group F contains four compounds; nicotine and varenicline have been grouped because of similar pharmacology (group G). 
-Nicotine has been shown to improve some of the symptoms of Alzheimer’s disease \citep{Jones1992}, it is therefore expected to 
-find this molecule in the predictions. Varenicline possess a pharmacology related to that of nicotine, which would explain 
-the presence of the drug in this category. The two remaining drugs of group F are respectively pralidoxime and dipivefrin; 
-little information is available regarding their potential action against the disease, yet these compounds seem linked to the 
+in the pathology, barbiturates could affect the state of the condition, similarly to cholinergic inhibitors \citep{Wang2000}.
+The group F contains four compounds; \emph{nicotine} and \emph{varenicline} have been further grouped together because of 
+similar pharmacology (group G).
+\emph{Nicotine} has been shown to improve some of the symptoms of Alzheimer's disease \citep{Jones1992}, it is therefore expected to 
+find this molecule in the predictions. \emph{Varenicline} possess a pharmacology related to that of \emph{nicotine}, which would explain 
+the presence of the drug in this category. The two remaining drugs of group F are respectively \emph{pralidoxime} and \emph{dipivefrin}; 
+little information is available regarding their potential action against the condition, yet these compounds seem linked to the 
 cholinergic hypothesis and could be considered for experimental testing.
 The groups H and I contain respectively one molecule each. These compounds have been classified as agents perturbing some of 
-the physiological function of the Tau protein, key actor in Alzheimer's disease \citep{Grundke-Iqbal1986}. Vorinostat (group H) is currently 
+the physiological function of the Tau protein, key actor in Alzheimer's 
+disease \citep{Grundke-Iqbal1986}. \emph{Vorinostat} (group H) is currently 
 indicated for the treatment of cutaneous manifestations in patients with T-cell lymphoma, yet a study has 
 shown in vivo (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
-deficit \citep{Kilgore2010}. The presence of lithium (group I) was 
+deficit \citep{Kilgore2010}. The presence of \emph{lithium} (group I) was 
 unexpected and suspected of being a false positive at first glance. However a recent study demonstrated a long-term 
-protective effect for the ion in regards to Alzheimer’s disease \citep{Young2011}. The last group (J) contains 
-ezetimibe and hesperetin. These two compounds are primarily used as cholesterol lowering agents (statins). As cholesterol 
+protective effect for the ion in regards to Alzheimer's disease \citep{Young2011}. The last group (J) contains 
+\emph{ezetimibe} and \emph{hesperetin}. These two compounds are primarily used as cholesterol lowering agents (statins). As cholesterol 
 metabolism in the brain appears to be related to dementia, statins are believed to prevent or improve the symptoms of the 
@@
 identify real repurposing hypotheses tailored to a disease. The MoAs help to retrieve the compounds which might impact 
-the treatment of a given set of symptoms.
+the treatment of a given set of symptoms or dysfunctionning biological processes.
 
@@
 This manuscript introduces the FTC, a free, open and extensible resource, which can assist drug repurposing 
-initiatives or enhance computational studies making use of the concept of mode of action. The construction of 
+initiatives or enhance computational studies making use of the concept of \emph{mode of action}. The construction of 
 the classification relies on description logics as a means to define the mode and mechanism of action of approved 

### 2013-08-20 document.tex
figure 5 changed as well as companion text
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 inside each of these classes as potential candidates. Figure \ref{fig:05} shows these drugs, which have been further manually grouped 
-based on the overall similarity of their actions (letters on Figure \ref{fig:05}).
+based on the overall similarity of their actions (numbers on Figure \ref{fig:05}).
 
@@
 \caption{FTC categories describing some of the modes and mechanisms of 
-action that could impact Alzheimer's disease. The categories have been manually picked on the basis that they could directly affect
+action that could impact Alzheimer's disease (letters on figure). The categories have been manually picked on the basis that they could directly affect
 the dementia. Drugs classified in these FTC categories further 
-manually grouped based on their MoAs similarities (subgroups B, C, D, E and G).}\label{fig:05}
+manually grouped based on their MoAs similarities (numbers on figure).}\label{fig:05}
 \end{figure}
 
-The subgroups B, C and D are inhibitors of the cholinergic system and some of them, such as \emph{galantamine} (DB00674), are already 
+The subgroups (1), (2) and (3) are inhibitors of the cholinergic system and some of them, such as \emph{galantamine} (DB00674), are already 
 investigated to treat Alzheimer's disease and other related dementias. This class of agent is in line with the 
 cholinergic hypothesis \citep{Francis1999}, stating that Alzheimer's disease could be caused by dysfunctions in the processing of the
-acetylcholine. The subgroup E is exclusively composed of barbiturates (central nervous system depressants). 
+acetylcholine. The subgroup (4) is exclusively composed of barbiturates (central nervous system depressants). 
 The presence of this pharmacological class of compounds as an Alzheimer's disease treatment is more surprising, as very little 
@@
 in the pathology, barbiturates could affect the state of the condition, similarly to cholinergic inhibitors \citep{Wang2000}.
-The group F contains four compounds; \emph{nicotine} and \emph{varenicline} have been further grouped together because of 
-similar pharmacology (group G).
+The group (B) contains four compounds; \emph{nicotine} and \emph{varenicline} have been further grouped together because of 
+similar pharmacology (group 5).
 \emph{Nicotine} has been shown to improve some of the symptoms of Alzheimer's disease \citep{Jones1992}, it is therefore expected to 
@@
 cholinergic hypothesis and could be considered for experimental testing.
-The groups H and I contain respectively one molecule each. These compounds have been classified as agents perturbing some of 
+The groups (C) and (D) contain respectively one molecule each. These compounds have been classified as agents perturbing some of 
 the physiological function of the Tau protein, key actor in Alzheimer's 
-disease \citep{Grundke-Iqbal1986}. \emph{Vorinostat} (group H) is currently 
+disease \citep{Grundke-Iqbal1986}. \emph{Vorinostat} (group C) is currently 
 indicated for the treatment of cutaneous manifestations in patients with T-cell lymphoma, yet a study has 
 shown in vivo (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
-deficit \citep{Kilgore2010}. The presence of \emph{lithium} (group I) was 
+deficit \citep{Kilgore2010}. The presence of \emph{lithium} (group D) was 
 unexpected and suspected of being a false positive at first glance. However a recent study demonstrated a long-term 
-protective effect for the ion in regards to Alzheimer's disease \citep{Young2011}. The last group (J) contains 
+protective effect for the ion in regards to Alzheimer's disease \citep{Young2011}. The last group (E) contains 
 \emph{ezetimibe} and \emph{hesperetin}. These two compounds are primarily used as cholesterol lowering agents (statins). As cholesterol 
@@
 From the examples briefly presented above, reported and confirmed by the literature, the FTC appears to be suitable to 
-identify real repurposing hypotheses tailored to a disease. The MoAs help to retrieve the compounds which might impact 
-the treatment of a given set of symptoms or dysfunctionning biological processes.
+identify real repurposing hypotheses tailored to a disease. Correctly 
+identifying MoAs of interest helps to retrieve the compounds which might impact 
+the treatment of a condition.
 

### 2013-08-21 document.tex
all proofs completed
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 
-\subsection{Exploration}
+\subsection{Exploration}\label{exploration}
 The FTC was designed to assist drug repurposing analyses by explicitly representing the polypharmacology of 
@@
 are accurately defined, in order to classify compounds based on the semantics of their relations. The properties 
-linking drugs to their respective protein targets (positive and negative perturbations) are however simplistic. At the 
+linking drugs to their respective protein targets (\emph{positive} and \emph{negative perturbations}) are however simplistic. At the 
 time being, no consideration is given regarding the binding strength between the drug and the proteins, yet it is a key 
@@
 of the protein could mediate different functions. Only one of such activity types can sometimes be inhibited 
-by a drug (cite felix paper), yet we are assuming in the FTC that as long as a drug affects a protein, 
+by a drug \citep{kruger}, yet we are assuming in the FTC that as long as a drug affects a protein, 
 it can therefore alter all its known functions.
 These limitations come from the semantics behind the axioms structuring the classification, themselves based on the 
-information available from the databases (e.g. GOA, DrugBank, etc...). Despite entailing not entirely accurately the 
+information available from the databases. Despite entailing not entirely accurately the 
 biochemical reality, the axioms help to generate a larger number of hypotheses, the primary goal of the FTC. The dosage 
-issue is partially addressed by the regulator pattern (see section in Availability and Implementation): It should be 
-easier to experimentally adjust the concentration of the compounds classified as pro or anti biological process agents 
+issue is partially addressed by the \emph{regulator pattern} (see section 3.1 of supplementary material): It should be 
+easier to experimentally adjust the concentration of the compounds classified as \emph{pro} or \emph{anti} biological process agents 
 in order to modulate a physiological effect.
@@
 the original data providers. Erroneous or missing information will lead to misclassification by the reasoner. 
-Some expected outcomes are also missing from the predictions; sildenafil for instance was expected to be classified 
-as pro-penile erection agent (FTC\_A0043084), yet the lack of appropriate GO annotation prevents it. After discussion 
+Some expected outcomes are also missing from the predictions; \emph{sildenafil} for instance was expected to be classified 
+as \emph{pro-penile erection agent} (FTC\_A0043084), yet the lack of appropriate GO annotation prevents it. After discussion 
 with the GOA curation team, a manual annotation can only be asserted based on published experimental results. No document 
-was found to support the involvement of the cGMP-specific 3',5'-cyclic phosphodiesterase (O76074 - sildenafil’s main target) 
-in the negative regulation of penile erection (GO:0060407), therefore no annotation can be made. Further work could be done 
+was found to support the involvement of the cGMP-specific 3',5'-cyclic phosphodiesterase (\emph{sildenafil}'s main target) 
+in the \emph{negative regulation of penile erection} (GO:0060407), therefore no annotation can be made. Further work could be done 
 in this direction, by trying to automatically infer more annotations or by using the electronically generated ones, in order 
@@
 \subsection{Interpreting the evaluation}
-The high recall value (88\%) supports the idea behind the automated build of the FTC: The data from different 
-repositories funded and curated in parallel, can be integrated to automatically create a new resource. This new resource (FTC) 
+Out of the evaluation, the high recall value (88\%) supports the idea behind the automated build of the FTC: The data from different 
+repositories funded and curated in parallel, can be integrated to automatically create a new resource. This new classification (FTC) 
 contains most of the known information present in an external gold standard (ATC) and relies on description logics to 
-leverage the native information. Here we have assessed the content of the FTC against the ATC, knowing that these two 
+leverage the native information. In the context of this work we have compared the content of 
+the FTC against the ATC, knowing that these two 
 taxonomies have diverging goals. During the evaluation, equivalences have been manually asserted between classes, 
-which are assumed to have fairly similar meaning and containing similar sets of compounds. These manual assertions 
+which are assumed to have fairly similar meaning and containing similar sets of compounds.
+These manual assertions 
 are however a weakness, as they are themselves not evaluated (free parameter). The presence or absence of a link was 
-determined only by one curator and any mistake can influence consequently the recall and precision values. In this regards, 
+determined only by one curator and any mistake can influence consequently the recall and precision values.
+The precision of 50\% tells that the FTC tends to over-assigns compounds to MoA categories. The low precision value is acceptable 
+in our case, as one of the underlying motivation of the FTC is to broadly represent polypharmacology, specially the one not present in
+gold standards such as the ATC, referencing only legal usage.
+In this regards, 
 the evaluation should be considered more as a safety control rather than a formal assessment of a predictive method.
@@
 reprofiling opportunity. These considerations do not interfere with the repurposing predictions generated 
-based on semantic similarities or discrete categories as presented in the Exploration section. Finally, 
+based on semantic similarities or discrete categories as presented in the section \ref{exploration}. Finally, 
 note that the ATC/FTC equivalences are open and editable online, any modification will be automatically 
@@
 The code behind the creation of the resource is entirely open and available 
-at {{https://github.com/loopasam/ftc}}. The web application built on the top of the FTC using the Brain library \citep{Croset2013}, and can be 
+at {{https://github.com/loopasam/ftc}}. The web application built on the top of the FTC using the Brain library \citep{Croset2013} and can be 
 find at {{https://www.ebi.ac.uk/chembl/ftc}}. The documentation can be accessed at {{https://github.com/loopasam/ftc/wiki}}. The full
@@
 \section*{Acknowledgement}
-Text Text Text Text Text Text  Text Text.  \citealp{Boffelli03} might want to know about  text text text text
+We would like to thank Rita Santos, Sarah Carl and Felix Kruger for the fruitful input on this manuscript.
 
-\paragraph{Funding\textcolon} Text Text Text Text Text Text  Text Text.
+\paragraph{Funding\textcolon} EMBL member states. S.C. is a member of Darwin College, University of Cambridge.
 

### 2013-08-21 app/build/FullOwlExporter.java
updated to drug from schema.org rather than chebi
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@ -80,7 +80,7 @@ public class FullOwlExporter extends OwlExporter {
 									//Add the class for the drug if not present
 									this.getBrain().addClass(drugBankFullClassName);
 									this.getBrain().label(drugBankClassName, "'" + drug.getName() + "'");
-									this.getBrain().subClassOf(drugBankClassName, "CHEBI_23888");
+									this.getBrain().subClassOf(drugBankClassName, "Drug");
 									this.getBrain().subClassOf(drugBankClassName, "FTC_C2");
 								}
 
@@ -137,7 +137,7 @@ public class FullOwlExporter extends OwlExporter {
 					"<a href='http://amigo.geneontology.org/cgi-bin/amigo/term_details?term=" + idForUrl + "'>" + goTerm.getName() + "</a>.");
 			this.getBrain().subClassOf(antiClassId, "FTC_C1");
 
-			String antiExpression = "CHEBI_23888 and FTC_R4 some (Protein and FTC_R2 some (GO_0003674 and " +
+			String antiExpression = "Drug and FTC_R4 some (Protein and FTC_R2 some (GO_0003674 and " +
 					goTermId + "))";
 
 			this.getBrain().equivalentClasses(antiClassId, antiExpression);
@@ -151,7 +151,7 @@ public class FullOwlExporter extends OwlExporter {
 					"rate or extent of the " +
 					"<a href='http://amigo.geneontology.org/cgi-bin/amigo/term_details?term=" + idForUrl + "'>" + goTerm.getName() + "</a>.");			
 			this.getBrain().subClassOf(proClassId, "FTC_C1");
-			String proExpression = "CHEBI_23888 and FTC_R5 some (Protein and FTC_R2 some (GO_0003674 and " +
+			String proExpression = "Drug and FTC_R5 some (Protein and FTC_R2 some (GO_0003674 and " +
 					goTermId + "))";
 
 			this.getBrain().equivalentClasses(proClassId, proExpression);
@@ -191,7 +191,7 @@ public class FullOwlExporter extends OwlExporter {
 					//('negative regulation of blood coagulation') we get the descendant classes and not the subclasses
 					//Some classes got better classified this way, (ex: pro-fibrinolysis becomes a subclass
 					//of anti-blood coagulation when using the full class name.)
-					String antiExpression = "CHEBI_23888 and FTC_R4 some (Protein and FTC_R1 some (GO_0008150 and " +
[…]

### 2013-08-22 document.tex
final polishing - ready for submission bioinformatics
Signed-off-by: loopasam <samuel.croset@gmail.com>

@@
 
+\begin{figure*}[!tpb]%figure1
+\centerline{\includegraphics{fig1.png}}
+\caption{Summary of the methodology used to construct the resource. (A) The FTC categories are first created from the existing GO terms
+describing molecular functions and biological processes. (B) Then, each category has a logical equivalent definition assign to it.
+(C) The content of various databases in incorporated and linked using the FTC specific properties. (D) Finally a reasoner runs over the knowledge
+base and assigns drugs to MoA categories based on whether or not a definition can be satisfied. The taxonomic structure of the
+FTC appears also in the reasoning step, from the entailements of the equivalent defintions.}\label{fig:01}
+\end{figure*}
+
 Drug repurposing is the use of known active compounds for new therapeutic indications \citep{Sanseau01072011}. 
-When administered in a living organism, a compound can indeed play various roles and affect different biological processes. 
-Accurately identifying these different functions helps to predict the potential side-effects a drug could have and can also lead to 
+When administered in a living organism, a compound can indeed play various roles and affect different biological processes; 
+accurately identifying these different functions helps to predict the potential side-effects a drug could have and can also lead to 
 interesting repurposing opportunities \citep{Medina-Franco2013}. For instance, \emph{sildenafil} was initially developed 
@@
 originating from description logics, such frameworks help to define the mathematical meaning of symbols and strings of characters. 
-In an ontology or knowledge base, concepts are organised and linked following the logical type of relation they have between them. 
+In an ontology or knowledge base, concepts are organised and linked following the logical type of relation they have among them. 
 In the Gene Ontology (GO) for example \citep{Ashburner2000}, biological processes and molecular functions terms are manually curated  
-and their meaning is specified by the relation types linking two GO terms.
+and their meaning specified by the relation types linking two GO terms.
 MoA definitions are present in other classifications such as the Medical Subject Headings \citep{Nelson2004} or the Chemical Entities of 
@@
 extent the action of drugs at the anatomical level. All these resources are valuable for the community as a source of carefully 
-manually curated information; moreover, the categories described in these classification systems are sometimes used to annotate 
+manually curated information. Moreover, the categories described in these classification systems are sometimes used to annotate 
 drugs: For instance the compound \emph{sildenafil} has been manually annotated as \emph{vasodilator agent} (CHEBI:35620 or MeSH:D27.505.954.411.918).
@@
 
-\begin{figure*}[!tpb]%figure1
-\centerline{\includegraphics{fig1.png}}
-\caption{Summary of the methodology used to construct the resource. (A) The FTC categories are first created from the existing GO terms
-describing molecular functions and biological processes. (B) Then, each category has a logical equivalent definition assign to it.
-(C) The content of various databases in incorporated and linked using the FTC specific properties. (D) Finally a reasoner runs over the knowledge
-base and assigns drugs to MoA categories based on whether or not a definition can be satisfied. The taxonomic structure of the
-FTC appears also in the reasoning step, from the entailements of the equivalent defintions.}\label{fig:01}
-\end{figure*}
-
 \begin{figure}[!tpb]%figure2
@@
 \caption{Parent categories to the FTC class \emph{Pro-fibrinolysis agent} (FTC\_P0042730). The classification is a direct 
-acyclic graph where categories are describing increasingly specific concepts. Arrows entails subclass relationships 
+acyclic graph where categories are describing increasingly specific concepts. Arrows entail subclass relationships 
 between the terms (\emph{is a} relation).}\label{fig:02}
@@
 has been assessed. In practice, these two resources have different goals; therefore the evaluation should be 
-interpreted accordingly. The full methodology behind the evaluation is 
+interpreted accordingly (cf section \ref{interpreting}). The full methodology behind the evaluation is 
 described in the section 6 of the supplementary material (\emph{Evaluation methodology}). 
@@
 considered. The \emph{evaluation points} cover a total of 471 DrugBank compounds or around 41\% of common drugs to both classifications.
-Out of these, 277 compounds are true positives, 
+Out of these, 275 compounds are true positives, 
 meaning that given an evaluation point, they are present in both FTC and ATC categories. The \emph{proton pump inhibitor} 
@@
 present both in the FTC category and in the corresponding ATC class(es). The total number of compounds present in an ATC class but 
-not in the corresponding FTC category (false negatives) is 36. Finally, 279 compounds are present in a FTC class but not in the 
+not in the corresponding FTC category (false negatives) is 35. Finally, 280 compounds are present in a FTC class but not in the 
 corresponding ATC class(es); such agents are the false positives. From these values, it is possible to derive a 
-recall of 88\%; this percentage indicates that the automatic build of the FTC covers a good portion of the content 
+recall of 89\%; this percentage indicates that the automatic build of the FTC covers a good portion of the content 
 already present in the ATC. The precision of 50\% shows that the FTC contains for a given MoA many more drugs than the 
@@
 \subsubsection{Polypharmacology spectrum}
-With more information on a drug’s molecular targets and their physiological roles, the more opportunities 
+The more information on a drug’s molecular targets and their physiological roles, the more opportunities 
 exist to re-orient a drug into doing something new. The therapeutic agents described in the FTC can have several MoAs, 
@@
 associated to each FTC category furthermore help the user to choose the adequate MoA for the desired outcome.
-According to these results, compound's polypharmacology is 
+Compound's polypharmacology is 
 well represented in the FTC, as shown by the numerous MoAs each approuved drug can exhibit.
@@
 therapeutic indications as defined by the ATC.
-The heatmap (Figure \ref{fig:04}) reveals some square patches around the central diagonal;
+The heatmap reveals some square patches around the central diagonal;
 the overall similarity appears higher when compounds from the same ATC group are considered. 
@@
 \centerline{\includegraphics{fig4.png}}
-\caption{Pairwise comparison of MoAs similarities. Only the first level of 
-the ATC is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
-The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black). Drugs are grouped according 
-to their first ATC level (colours on the side). For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
-which appears as ``B'' (dark orange) on the plot.
-Some compounds belong to multiple ATC categories (Mulitple - black) and some others do not have
-an ATC code (NoCategory - white)
+\caption{Pairwise comparison of MoAs similarities. Therapeutic indications are represented by ATC categories which are the colors on the side.
+For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
+which appears as \emph{B} (dark orange) on the plot.
+Only the first ATC level 
+is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
+The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black).
+Some compounds belong to multiple ATC categories (\emph{Mulitple}) and some others do not have
+an ATC code (\emph{NoCategory}).
 The average similarity of drugs present in the same therapeutic category 
@@
 The FTC is a novel classification for approved drugs, which can be used as a starting point to 
-generate drug repurposing hypotheses. We have guided the reader through the various ways the resource 
+generate drug repurposing hypotheses. We have guided the reader through some of the ways the resource 
 can be explored. No experimental validations have been presented in this manuscript, yet references to 
@@
 
-\subsection{Interpreting the evaluation}
-Out of the evaluation, the high recall value (88\%) supports the idea behind the automated build of the FTC: The data from different 
+\subsection{Interpreting the evaluation}\label{interpreting}
+Out of the evaluation, the high recall value (89\%) supports the idea behind the automated build of the FTC: The data from different 
 repositories funded and curated in parallel, can be integrated to automatically create a new resource. This new classification (FTC) 
@@
 This manuscript introduces the FTC, a free, open and extensible resource, which can assist drug repurposing 
-initiatives or enhance computational studies making use of the concept of \emph{mode of action}. The construction of 
+initiatives or enhance computational studies making use of the concept of \emph{mode of action}.
+The resource annotates the biological functions of drugs, in a similar way the GO serves to annotate gene products.
+The construction of 
 the classification relies on description logics as a means to define the mode and mechanism of action of approved 
@@
 The code behind the creation of the resource is entirely open and available 
-at {{https://github.com/loopasam/ftc}}. The web application built on the top of the FTC using the Brain library \citep{Croset2013} and can be 
+at {{https://github.com/loopasam/ftc}}. The web application and the FTC are built using the Brain library \citep{Croset2013} and can be 
 find at {{https://www.ebi.ac.uk/chembl/ftc}}. The documentation can be accessed at {{https://github.com/loopasam/ftc/wiki}}. The full

### 2013-08-22 document.tex
final polishing done
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-08-20 document.tex
@@
 
-\title[The Functional Therapeutic Chemical Classification System]{The Functional Therapeutic Chemical Classification System}
-\author[Croset \textit{et~al}]{Samuel Croset\,$^{1, \footnote{to whom correspondence should be addressed}}$, John Overington\,$^{1}$ and Dietrich Rebholz-Schuhmann\,$^{1}$
-}
-\address{$^{1}$European Molecular Biology Laboratory
-European Bioinformatics Institute (EMBL-EBI), Wellcome Trust Genome Campus, Hinxton, Cambridge CB10 1SD,
-United Kingdom\\}
+\title[Supplementary information - The Functional Therapeutic Chemical Classification System]{Supplementary information - 
+The Functional Therapeutic Chemical Classification System}
 
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
+\section*{Availability and Implementation}
 
-\editor{Associate Editor: XXXXXXX}
+This section describes the building mechanism behind the FTC. 
+First, a list of categories describing the mode and mechanism of action of drugs is defined. 
+Then in a second step the newly created categories are automatically populated with approved compounds. 
+Finally, the FTC is evaluated and repurposing hypotheses can be generated.
 
-\maketitle
+\section{Source code}
+The code behind the creation of the resource is entirely open and available 
+at {{https://github.com/loopasam/ftc}}. The web application built on the top of the 
+FTC can be find at {{https://www.ebi.ac.uk/chembl/ftc}} and the documentation can be 
+accessed at {{https://github.com/loopasam/ftc/wiki}}. The reader should be familiar with 
+description logics and the Web Ontology Language (OWL) to fully understand the construction of the 
+knowledge base. An introduction to description logics from the perspective of the biomedical scientist is 
+available on the wiki at {{https://github.com/loopasam/ftc/wiki/Description-Logics}}. The FTC implementation 
+relies mostly on Brain \citep{Croset2013} and the web application builds on the top of the Play! framework. 
+Classification tasks use the ELK reasoner \citep{Kazakov2011}. The computer hosting the web application has 8 Gb of memory 
+with 4 processors, this architecture allows fast parallel reasoning, thanks to ELK's design. More functionalities 
+will be added to the web application following user requirements (so called \emph{lean implementation}).
 
-\begin{abstract}
+\section{FTC categories creation}
+The mode of action categories present in the FTC are defined based on the terms coming from the 
+Gene Ontology (GO) \citep{Ashburner2000}. Both the molecular function and biological process branches are used for 
+this purpose, yet handled slightly differently.
 
-\section{Motivation:}
-Drug repurposing is the discovery of new indications for compounds already approved and used in a clinical setting. 
-Recently, some computational approaches have been developed to unveil new opportunities in a systematic fashion, by taking 
-into consideration gene expression signatures or chemical features for instance. 
-We present here a novel method based on knowledge integration using semantic technologies, 
-to capture the functional role of approuved chemical compounds.
-\section{Results:}
-In order to computationally generate repurposing hypotheses, we used the Web Ontology Language (OWL) to 
-mathematically define the meaning of over 20'000 terms formally describing various modes and mechanisms of action. Based on an 
-integration of public data, we have automatically assigned over a thousand of approved drugs into 
-these MoA categories. The resulting new resource is called the Functional Therapeutic Chemical Classification System (FTC) 
-and was further evaluated against the content of the traditional Anatomical Therapeutic Chemical Classification System (ATC). 
-We illustrate how the new classification can be used to generate drug repurposing hypotheses, using Alzheimer’s disease as a use-case.
+\subsection{Categories related to biological processes}\label{catprocess}
+All the biological processes featured by the GO are looked-up one by one. All the 
+time a process is linked to another process (\emph{X}) via a \emph{positive} or \emph{negative regulation} link, two FTC 
+classes are created: \emph{Anti-X agent} and \emph{Pro-X agent}. For instance the GO term \emph{positive regulation of blood 
+coagulation} is linked to the term \emph{blood coagulation} via a \emph{positively regulates} relation, therefore two FTC 
+categories \emph{Anti-blood coagulation agent} and \emph{Pro-blood coagulation agent} are created. The identifiers of the 
+new FTC classes are also derived from the GO term used to create the class pattern. The GO numeric identifier 
+is re-used and the letter \emph{A} or \emph{P} is appended before to emphasize the \emph{anti} or \emph{pro} pattern. 
+From the example presented 
+previously, the FTC class \emph{Anti-blood coagulation} has FTC\_A0007596 as identifier, because the GO term \emph{blood coagulation} is 
+referenced by GO:0007596. Following the same logic, FTC\_P0007596 is the identifier of the class \emph{Pro-blood coagulation agent}. The 
+design choice for identifiers and labels allows the FTC to fully rely on the high quality work provided by the GO curation 
+team and scale over it.
 
-\section{Availability:} A web application built on the top of the resource is freely available at {{https://www.ebi.ac.uk/chembl/ftc}}. The 
-source code of the project is available at {{https://github.com/loopasam/ftc}}.
+\subsection{Categories related to molecular functions}\label{catfunc}
+The mode and mechanism of actions related to molecular functions are created in the following manner: All the time 
+a molecular function (\emph{Y}) is encountered then two FTC categories are created, as for the processes: \emph{Anti-Y agent} and 
+\emph{Pro-Y agent}. The identifiers are assigned the same way as described before. For instance, out of the GO term \emph{androgen 
+receptor activity} (GO:0004882) two FTC classes are derived: \emph{Pro-androgen receptor activity agent} (FTC\_P0004882) and 
+\emph{Anti-androgen receptor activity agent} (FTC\_A0004882).
 
-\section{Supplementary Information:}
-Supplementary data are available at Bioinformatics online.
+\section{Equivalent definitions}
+FTC classes are generated as presented in the previous section. Up to this point, these categories 
+are only tokens with a human readable label as well as an identifier. The next step is going to 
+assign equivalent definitions to each FTC class. A reasoner can understand such definitions and will automatically 
+classify the knowledge base accordingly; drugs will then be assigned to FTC categories and the taxonomic structure 
+arises after this reasoning step. Equivalent definitions are written as OWL class expressions using the entities of 
+the knowledge base (summarised at {{https://github.com/loopasam/ftc/wiki/Knowledge-Base}} and in section \ref{specs}).
+There are two types of equivalences: The first one captures perturbation of regulatory 
+biological processes (so called \emph{regulatory patterns}) and the second one handles the perturbed functions (\emph{functional patterns}).
 
-\section{Contact:} \href{croset@ebi.ac.uk}{croset@ebi.ac.uk}
-\end{abstract}
+\subsection{Regulatory pattern}
+Some of the FTC categories are created from the biological processes present in the 
+GO (cf section \ref{catprocess}); these categories have two arbitrary equivalent definitions, representing the 
+two possible ways a compound might impact the biological process. Anti-biological process agent FTC categories 
+contain the drugs that negatively perturb a target involved in the positive regulation of the biological process. 
+The \emph{anti} categories also feature the compounds that positively perturb a negative regulator of the same process. 
+The \emph{pro} categories are equivalent to the opposite pattern. Supplementary Figure \ref{fig:S01} illustrates the equivalent definitions for the 
+FTC class \emph{Anti-blood coagulation agent} (FTC\_A0007596).
 
-\section{Motivation}
+\begin{figure}[!tpb]%figureS1
+\centerline{\includegraphics{figS1.png}}
+\caption{Example of equivalent definitions (gray boxes) for the FTC category \emph{Anti-blood coagulation 
+agent} (FTC\_A0007596). After the data integration step, a reasoner will look the definitions and identify 
+which drugs satisfy the FTC definitions. Definitions are expressed using the Web Ontology Language (OWL), 
+serialised here using the Manchester syntax.}\label{fig:S01}
+\end{figure}
 
-Drug repurposing is the use of known active compounds for new therapeutic indications \citep{Sanseau01072011}. 
-When administered in a living organism, a compound can indeed play various roles and affect different biological processes. 
-Accurately identifying these different functions helps to predict the potential side-effects a drug could have and can also lead to 
-interesting repurposing opportunities \citep{Medina-Franco2013}. For instance, \emph{sildenafil} was initially developed 
-against angina and has been repurposed towards erectile dysfunction during the clinical trials \citep{Ashburn2004}. Approved compounds are 
-attractive because they have been extensively studied and have by definition already successfully passed clinical trials, where 
-most drugs fail because of safety or efficacy issues.
-There is an increasing number of approaches to predict repurposing opportunities using computational methods \citep{Sanseau01072011}. Most 
-methods operate on the profiles of physicochemical descriptors derived from molecular structures \citep{Haupt2011}. Other methods characterise 
-the drugs on more abstract levels, such as the gene expression signature \citep{Iorio2010} or 
-via the reported side-effects \citep{Campillos2008}. 
-These approaches 
-have in common to look for similarities within existing drugs and forward similar compounds as repurposing hypotheses.
-A feature of particular interest to describe drugs is the Mode of Action (MoA). According to Wikipedia, 
-the MoA describes \emph{a functional or anatomical change, 
-at the cellular level, resulting from the exposure of a living organism to a substance}. For instance 
-terms such as \emph{transcriptional regulation agent} or \emph{anticoagulant} define MoAs and characterise the 
-roles of a certain type of drugs. The 
-MoA abstracts over the relations between molecular functions, protein targets and drug activities; it is the central concept linking a 
-chemical structure to a set of biological effects. Intuitively, the indication of a drug logically depends on its MoA.
-Despite its widespread use in drug discovery, the MoA has not been used yet as a descriptor for repurposing analyses. One reason for this 
-might be the challenge of formally defining MoAs. Indeed, MoAs are terms or categories, it is not possible to represent 
-them straightforwardly with values and numbers like one can do for a 3D molecular structure or for a gene expression profile. 
-Nonetheless, the meaning of a concept can be formalised with controlled vocabularies and ontologies \citep{Gruber1995}; 
-originating from description logics, such frameworks help to define the mathematical meaning of symbols and strings of characters. 
-In an ontology or knowledge base, concepts are organised and linked following the logical type of relation they have between them. 
-In the Gene Ontology (GO) for example \citep{Ashburner2000}, biological processes and molecular functions terms are manually curated  
-and their meaning is specified by the relation types linking two GO terms.
-MoA definitions are present in other classifications such as the Medical Subject Headings \citep{Nelson2004} or the Chemical Entities of 
-Biological Interest \citep{Hastings2012} for example. The Anatomical Therapeutic 
-Chemical Classification System (ATC) \citep{world2000anatomical} also describes to some 
-extent the action of drugs at the anatomical level. All these resources are valuable for the community as a source of carefully 
-manually curated information; moreover, the categories described in these classification systems are sometimes used to annotate 
-drugs: For instance the compound \emph{sildenafil} has been manually annotated as \emph{vasodilator agent} (CHEBI:35620 or MeSH:D27.505.954.411.918).
-The classifications mentioned previously are not specially designed for drug repurposing; they purposefully report only the 
-well-known and major MoAs of chemical compounds. The pharmacological spectrum of each drug is not necessarily well covered, yet 
-it would be the best way to predict new indications. In our context, an ideal knowledge base would feature the known MoAs of a 
-drug as well as some predicted ones to be tested in experiments. The MoA categories should derive and scale over primary molecular 
-evidences exposed in biomedical databases in an automated way.
-To address the lack of systematic MoA annotations, we have implemented the Functional Therapeutic Chemical 
-Classification System (FTC), presented here in this manuscript. The FTC is automatically built by leveraging the content 
-of various biomedical databases using description logics and automated reasoning. Over 20'000 new MoA categories are defined 
-in the resource and further populated with approved drugs using the Web Ontology Language (OWL) in combination with a reasoner. 
-Most of the drugs are present in multiple FTC categories, reflecting the various roles a compound can play inside a biological system which 
-can serve as starting point for repurposing. 
-The resource was evaluated against the ATC, traditional classification scheme introduced before. We present as well some preliminary 
-analyses over the data, by looking at the relation between the MoA and the indication of a compound using semantic similarity. 
-Finally, we illustrate also how the FTC can be used as a pharmacological toolbox to analyse drug repurposing for Alzheimer’s disease.
+\subsection{Functional pattern}
+The FTC categories generated from the GO molecular functions (cf section \ref{catfunc}) are also equivalent to a logical 
+definition. \emph{Anti} FTC categories dealing with molecular activities are asserted as equals to the drugs 
+that negatively perturb the function. \emph{Pro} categories are equivalent to the drugs that positively perturb the 
+function of interest. A summary of the patterns definitions is available on the 
+online wiki at {{https://github.com/loopasam/ftc/wiki/Mode-of-Action}}.
 
-\section{Results}
+\section{Data integration}
+At this stage, the knowledge base contains the created FTC classes associated with their 
+logical definitions, as well as the GO and the core FTC entities. The knowledge base is then 
+further populated with some information coming from various public databases. Only manually curated 
+information is considered.
 
-The knowledge base behind the FTC is built by integrating information coming from various sources. 
-The GO terms serve as template to create the FTC categories describing the MoAs; DrugBank \citep{Knox2011} provides the 
-known links between drugs and their protein targets and Uniprot \citep{TheUniprotConsortium2013} maps targets to their 
-respective GO annotations \citep{Dimmer2012}. 
-Drugs are further assigned into MoA categories based of the OWL constructs and axioms defined in the 
-FTC (see \emph{Availability and Implementation}). A reasoner, a program capable of understanding such axioms, performed this task. 
-The process to build the FTC is summarised in Figure~\ref{fig:01}.
-It takes around 5 seconds on a modern 
-laptop (4 cores processor with 4 Gb of memory) to classify the knowledge base using the ELK reasoner \citep{Kazakov2011}. 
-Other OWL reasoners have been tried (e.g, Hermit, Pellet, etc.) but none of them were suitable for this purpose; 
-the classification time was always superior to a few minutes (data not shown). The FTC has a taxonomic structure as illustrated 
-on Figure~\ref{fig:02}, which arises when the reasoner classifies the knowledge base. Categories can have multiple parents and multiple children. 
-The reader can browse and access the content of the FTC online at {{https://www.ebi.ac.uk/chembl/ftc}}.
-In total there are 1280 approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets. 
-The FTC introduces 23'353 new categories describing the mode and mechanism of action of therapeutic compounds. 4289 of these 
-categories are related to biological processes and 19'064 to the molecular functions. A summary of the metrics behind the latest 
-build is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.
-Out of all FTC classes, 1432 
-categories (\textgreater 6\%) directly contain at least one approved drug. This number increases up to 2532 (\textgreater 11\%) 
-when direct and 
-indirect drugs are considered. FTC categories not containing drugs (e.g, FTC\_A0001771 - \emph{Anti-immunological synapse 
-formation agent}) represent modes and mechanisms of action for which no approved compounds exist yet or that have not been identified 
-as such in the FTC.
+\subsection{DrugBank}
+The DrugBank database is a unique bioinformatics and cheminformatics resource 
+that combines detailed drug data with comprehensive drug target information \citep{Knox2011}. The approved 
+drugs (small molecules and biotherapeutics) acting on proteins are extracted from the database and 
+imported in the FTC knowledge base. In order to be selected, a compound must firstly be approved and 
+secondly have a pharmacological action on at least one human protein target present in Uniprot \citep{TheUniprotConsortium2013}. 
+The protein targets all have at least one manually asserted GO annotation \citep{Dimmer2012} for a biological process or a 
+molecular function. DrugBank links compounds to targets via \emph{actions}. The DrugBank actions are somehow 
+structured and consistent: Concepts such as \emph{inhibitor} or \emph{agonist} are reused throughout the database for 
+example, yet they are not strictly formalised as a controlled vocabulary. These actions are manually 
+standardised to the core properties of the FTC according to their biochemical meaning: For instance the action 
+\emph{antagonist} is mapped to the FTC \emph{negatively-perturbs} property.
+Compounds coming from DrugBank are represented as OWL classes and asserted 
+as subclasses of the class \emph{DrugBank compound} (FTC\_C2). Protein targets are described as OWL classes 
+too and subclasses of the core class \emph{Protein}. Each DrugBank compound is then connected to 
+its target via the following axiom pattern: \emph{[drug] SubClassOf perturbs some [protein]}. 
+E.g. \emph{Ximelagatran SubClassOf negatively-perturbs some Prothrombin}.
 
-\begin{figure*}[!tpb]%figure1
-\centerline{\includegraphics{fig1.png}}
-\caption{Summary of the methodology used to construct the resource. (A) The FTC categories are first created from the existing GO terms
-describing molecular functions and biological processes. (B) Then, each category has a logical equivalent definition assign to it.
-(C) The content of various databases in incorporated and linked using the FTC specific properties. (D) Finally a reasoner runs over the knowledge
-base and assigns drugs to MoA categories based on whether or not a definition can be satisfied. The taxonomic structure of the
-FTC appears also in the reasoning step, from the entailements of the equivalent defintions.}\label{fig:01}
+\subsection{Gene Ontology annotations}
+The GO annotation program aims to provide high-quality GO annotations to 
+proteins in UniProt \citep{Dimmer2012}. In the context of the FTC, such annotations are used 
+to create axioms linking protein targets to molecular functions and biological processes. 
+Each protein annotated with a function creates an axiom such 
+as \emph{[protein] SubClassOf has-function some [molecular function]}. 
+Each protein annotated to a biological process creates an axiom such 
+as \emph{[protein] SubClassOf involved-in some [biological process]}. 
+E.g. \emph{Prothrombin SubClassOf involved-in some positive regulation of blood coagulation}. 
+Each protein can be involved in multiple processes and capable of realizing multiple functions;
+the polypharmacology is captured at this level.
+
+\section{Knowledge base classification}
+The knowledge base is fully built at this step and contains core classes, mode 
+of actions descriptions alongside the actions of approved DrugBank compounds on 
+protein targets in Uniprot. The proteins are linked to their molecular functions and involvement in 
+biological processes via the GO annotations. The logical specifications of the FTC are there to glue 
+the different data together and to explicitly express the logical links between resources. The FTC 
+knowledge base follows an OWL2 EL profile \citep{Motik2009}, which enable the use of fast and parallelised reasoners such 
+as ELK. During the classification process, the reasoner checks whether the mode of action equivalent 
+definitions are satisfied or not and assigns drugs inside the corresponding FTC categories. The tree structure 
+of FTC appears also at this step from the logical definitions.
+
+\section{Evaluation methodology}
+As the classification of therapeutic agents is done in an automated way, it 
+is important to evaluate the results generated against a known resource which will be 
+considered as gold standard. The assessment of the FTC is done against another similar 
+classification, the Anatomical Therapeutic Chemical Classification System (ATC) \citep{world2000anatomical}). 
+The ATC has been developed to serve as a tool for drug utilization research in order to 
+improve quality of drug use. In this resource, the information is manually curated, and drugs 
+are assigned to categories based of their legally approved indications. The goal of the ATC differs 
+from the one of the FTC, yet the two resources are sharing some very similar concepts, which can be 
+used for the evaluation. Categories of both classifications contain approved drugs with a DrugBank 
+identifier, meaning that some of the drugs indexed in the FTC are also present in the ATC. From that, 
+it is possible to define some evaluation points, which will help to assess the automated classification process.
+
+\subsection{Evaluation Points}
+An evaluation point is defined as an equivalence between a class from the FTC with one or 
+more classes from the ATC. The idea is to look at the set of drugs contained in both side of the 
+equivalence and estimate the overlap, as illustrated in the Supplementary Figure \ref{fig:S02}. Evaluation points are defined 
+by hand and not themselves evaluated. The full list of evaluation points as well as a summary of the results 
+are available online at https://www.ebi.ac.uk/chembl/ftc/evaluation/. Each evaluation point has a series of 
+true/false positive and false negative drugs associated with it.
+
+\begin{figure}[!tpb]%figureS2
+\centerline{\includegraphics{figS2.png}}
+\caption{Example of evaluation point. Classes from the ATC are manually 
+mapped to classes from the FTC. The drugs present in each side of the equivalences 
+are compared, and some metrics are derived from it (false positive/negative and true positive). 
+A summary of the evaluation is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.}\label{fig:S02}
+\end{figure}
+
+\subsection{True Positives}
+Drugs that are present in both the FTC and the equivalent ATC class(es) are called true positives. 
+These compounds reflect that the automated classification was capable of retrieving correctly the 
+information present in the gold standard (ATC).
+
+\subsection{False Negatives}
+These drugs are present in the ATC class(es) but not in the corresponding FTC class. 
+The automated classification failed to retrieve these compounds. The smaller the number of 
+false negatives is, the better the FTC is at recalling drugs. A small number of false negatives 
+means that if a drug is present in the ATC (gold standard), then it is likely that the drug will 
+also be correctly categorised in the FTC.
+
+\subsection{False Positives}
+The false positives are the drugs present in the FTC category of the evaluation point but not in the corresponding ATC classes. 
+A high number of false positives means that the FTC is over-assigning compounds to classes.
+The false positives relates to the accuracy of the classification. In the context of this work, 
+some false positives could also be considered as drug repurposing opportunities.
+
+\subsection{Precision}
+Precision is the probability that a randomly selected drug from the FTC is present in the equivalent ATC classes. 
+The value is standardised as a percentage and corresponds to the formula: True Positive / (True Positive + False Positive).
+
+\subsection{Recall}
+Recall is the probability that a randomly selected drug from the ATC has been assigned to the correct 
+corresponding class in the FTC. The value is standardised as a percentage and corresponds 
+to the formula: True Positive / (True Positive + False Negative).
+
+\section{Semantic similarity}
+The semantic similarity measure performed over the FTC is a derivative of the 
+Jaccard index \citep{jaccard1912distribution} \citep{rogers1960computer} (see \cite{pesquita2009semantic} for review). 
+It is probably best understood as an example: If we consider two classes A and B, the semantic similarity 
+between these classes corresponds to the number of OWL superclasses (direct and indirect, obtained with a reasoner) that 
+are shared by A and B (intersection) divided by the number of superclasses of A or B (union). The index ranges 
+from 0 (totally different) to 1 (identical).
+
+\section{Mode of action similarity against indication}
+A statistical analysis was performed over the data presented on Figure 4. 
+When two compounds are randomly taken, they have on average a higher mode of action similarity when 
+they are assigned to the same ATC category (one ATC level).
+In order to estimate whether this observation was 
+due to chance only, we formulated the following null hypothesis (H0): \emph{For a pair of drug A and B, it does not 
+matter to which ATC category they belong to, their similarity is always average.}
+The alternative hypothesis (H1) was: \emph{For a pair of drug A and B, if A and B have the same ATC code, 
+we expect on average to obtain a higher similarity value than if A and B have different ATC codes.}
+
+A permutation test was then performed for each ATC category. For example, we started with the ATC 
+category A (first row on Figure 4), looked at the similarity values when pairs of compounds both belong 
+to the category A (top right corner square) and compared it to the similarity values when the pair of compounds 
+belong to different categories (A and B, A and C and so forth). For each comparison we obtained two distributions of 
+values (not shown). On average the similarity values are always higher when the two compounds 
+belong to the same category (A/A versus A/B for instance). A permutation test (n = 20'000) was 
+performed in order to see whether this observation was due to chance only. We were able to reject the null 
+hypothesis for a significance level of 0.05 all the times. 
+The choice for a permutation test was driven by the fact that MoA similarity values do not follow any type 
+of standard distribution (data not shown). For the significance levels set, the null hypothesis was rejected for 
+all first level ATC categories.
+
+\section{Knowledge base specification}\label{specs}
+This section presents the scaffold of the knowledge base underlying the FTC. 
+The logic structuring the FTC comes essentially from a set of core OWL properties (rich RBox). 
+Some of these properties originate from the GO. When necessary some new ones have also been introduced. 
+In order to understand how these properties interact, first will be presented the fundamental classes present 
+at the top of the FTC classification.
+
+\subsection{Core classes}
+The high level concepts covered by the FTC are represented as OWL classes and enumerated below. 
+Some of these core classes are coming from external ontologies, in which case the original URI is preserved.
+
+\begin{itemize}
+\item \textbf{Molecular Function} \\
+Identifier: http://purl.obolibrary.org/obo/GO\_0003674 \\
+Label: molecular function \\
+Definition: As defined by the Gene Ontology. Elemental activities, such as catalysis or binding, describing 
+the actions of a gene product at the molecular level. A given gene product may exhibit one or more molecular functions.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Biological Process} \\
+Identifier: http://purl.obolibrary.org/obo/GO\_0008150 \\
+Label: biological process \\
+Definition: As defined by the Gene Ontology: Any process specifically pertinent to the functioning of 
+integrated living units: cells, tissues, organs, and organisms. A process is a collection of molecular 
+events with a defined beginning and end.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Protein} \\
+Identifier: http://purl.uniprot.org/core/Protein \\
+Label: protein \\
+Definition: As defined by Uniprot. Description of a protein.
+Comment: Gene products present inside the FTC are all human proteins. Uniprot URIs are used.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Drug} \\
+Identifier: http://schema.org/Drug \\
+Label: drug \\
+Definition: As defined by schema.org: A chemical or biologic substance, used as a medical therapy, that has a 
+physiological effect on an organism. \\
+Comment: In the context of the FTC, DrugBank chemicals are considered for their role as therapeutic agent 
+rather than for their chemical structure. Mode of action classes are representing agent types and are subclasses of this concept.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Therapeutic Agent} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_C1 \\
+Label: therapeutic agent \\
+Definition: Role of a drug capable of producing a therapeutic effect.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{DrugBank Compound} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_C2 \\
+Label: DrugBank compound \\
+Definition: Drug coming from DrugBank.
+\end{itemize}
+
+\subsection{Core properties}
+The expressivity of the FTC comes mostly for the properties (RBox). Below is a list of the 
+OWL object properties present inside the FTC knowledge base.
+
+\begin{itemize}
+\item \textbf{Part Of} \\
+Identifier: http://purl.obolibrary.org/obo/BFO\_0000050 \\
+Characteristic: Transitive \\
+Label: part-of \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Has Part} \\
+Identifier: http://purl.obolibrary.org/obo/BFO\_0000051 \\
+Characteristic: Transitive \\
+Label: has-part \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002211 \\
+Chained property: regulates o part-of -\textgreater regulates \\
+Label: regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Negatively Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002212 \\
+SubPropertyOf: regulates \\
+Label: negatively-regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Positively Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002213 \\
+SubPropertyOf: regulates \\
+Label: positively-regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Involved In} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R1 \\
+Label: involved-in \\
+Domain: protein \\
+Range: biological process \\
+Definition: Entails the participation of a protein in a biological process. \\
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Has Function} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R2 \\
+Label: has-function \\
+Domain: protein \\
+Range: molecular function \\
+Definition: Describes the molecular function a protein can realize. \\
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R3 \\
+Label: perturbs \\
+Domain: drug \\
+Range: protein \\
+Definition: Specific biochemical interaction through which a drug substance 
+will affect the activity of a protein (mechanism of action). The property refers to the specific molecular 
+targets to which the drug binds, such as an enzyme or receptor.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Negatively Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R4 \\
+Label: negatively-perturbs \\
+SubPropertyOf: perturbs \\
+Definition: Specific biochemical interaction through which a drug substance will decrease the activity of a protein. 
+The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Positively Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R5 \\
+Label: positively-perturbs \\
+SubPropertyOf: perturbs \\
+Definition: Specific biochemical interaction through which a drug substance will increase the activity of a protein. 
+The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
+\end{itemize}
+
+\begin{figure*}[!tpb]%figureS3
+\centerline{\includegraphics{figS3.png}}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but with two ATC levels considered. It is possible to observe some clustering around the diagonal line (no
+statistical analysis performed).}\label{fig:S03}
 \end{figure*}
 
-\begin{figure}[!tpb]%figure2
-\centerline{\includegraphics{fig2.png}}
-\caption{Parent categories to the FTC class \emph{Pro-fibrinolysis agent} (FTC\_P0042730). The classification is a direct 
-acyclic graph where categories are describing increasingly specific concepts. Arrows entails subclass relationships 
-between the terms (\emph{is a} relation).}\label{fig:02}
-\end{figure}
-
-\subsection{Evaluation}
-The content of the FTC has been evaluated against the information present in the ATC. The motivation behind 
-this step is to have an estimation of how the automatically-built classification (FTC) compares to a manually curated 
-one (ATC). In the context of this work, we have considered the ATC as a gold standard against which the FTC 
-has been assessed. In practice, these two resources have different goals; therefore the evaluation should be 
-interpreted accordingly. The full methodology behind the evaluation is 
-described in the section 6 of the supplementary material (\emph{Evaluation methodology}). 
-Briefly, 68 categories present in the FTC have an equivalent meaning to a combination of categories present in the ATC. 
-This semantic equivalence was manually asserted 
-and called \emph{evaluation point} (see supplementary material section 6.1 for details). 
-The set of drugs present in both classes for each evaluation points were then compared and some metrics derived. 
-For example, the FTC category \emph{Anti-hydrogen:potassium-exchanging ATPase activity agent} (FTC\_A0008900) has 
-been manually asserted as equivalent to the ATC category \emph{Proton pump inhibitors} (A02BC). A summary of this 
-evaluation point is furthermore available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/FTC\_A0008900}}.
-Out of the 1280 DrugBank compounds present in the FTC, 1134 are also present in the ATC, therefore only those were 
-considered. The \emph{evaluation points} cover a total of 471 DrugBank compounds or around 41\% of common drugs to both classifications.
-Out of these, 277 compounds are true positives, 
-meaning that given an evaluation point, they are present in both FTC and ATC categories. The \emph{proton pump inhibitor} 
-evaluation point is such an example where all the drugs (\emph{omeprazole, esomeprazole, pantoprazole, lansoprazole, rabeprazole}) are 
-present both in the FTC category and in the corresponding ATC class(es). The total number of compounds present in an ATC class but 
-not in the corresponding FTC category (false negatives) is 36. Finally, 279 compounds are present in a FTC class but not in the 
-corresponding ATC class(es); such agents are the false positives. From these values, it is possible to derive a 
-recall of 88\%; this percentage indicates that the automatic build of the FTC covers a good portion of the content 
-already present in the ATC. The precision of 50\% shows that the FTC contains for a given MoA many more drugs than the 
-equivalent ATC categories. This result was expected and comes the original idea behind the FTC: Representing in a systematic fashion
-the implicit and explicit MoAs of drugs, in particular the ones not already indexed by current classification scheme.
-
-\subsection{Exploration}
-The FTC was designed to assist drug repurposing analyses by explicitly representing the polypharmacology of 
-approved drugs. In this section we exemplify how the resource can be used to perform different types of analysis.
-
-\subsubsection{Polypharmacology spectrum}
-With more information on a drug’s molecular targets and their physiological roles, the more opportunities 
-exist to re-orient a drug into doing something new. The therapeutic agents described in the FTC can have several MoAs, 
-representing the intrinsic polypharmacology of the approved compounds. Figure \ref{fig:03} illustrates the polypharmacology 
-spectrum by showing the distribution of number of MoAs per compound. When only direct categories are considered, compounds belong on average
-to 13.5 MoA categories. This number increases to 61.2 when parent categories are taken into consideration (super classes).
-Not all the MoAs are disease relevant, 
-some FTC categories are particularly abstract (e.g, \emph{Anti-biological process agent}) yet they represent discrete 
-categories to which the drug belongs with an explicit and clear meaning. These discrete MoAs are a good starting 
-point to understand what a compound can do when administered in a human system, as it will be shown with Alzheimer's disease.
-The human readable definitions 
-associated to each FTC category furthermore help the user to choose the adequate MoA for the desired outcome.
-According to these results, compound's polypharmacology is 
-well represented in the FTC, as shown by the numerous MoAs each approuved drug can exhibit.
-
-\begin{figure*}[!tpb]%figure3
-\centerline{\includegraphics{fig3.png}}
-\caption{Distribution of the direct (blue) and indirect (red) number of MoAs per drug.
-Means are indicated with a solid line. On average each compound has 13.5 MoAs when only direct classes are considered. The number rises
-to 61.2 when indirect MoAs are included. Indirect MoAs are the ancestor classes in the taxonomy as shown in Figure \ref{fig:02}.
-The distribution range is wider when indirect MoAs are considered (range=299; min=7; max=306) versus direct MoAs only (range=79; min=3; max=82).
-These results emphasises the fact that some drugs
-are well characterised in databases and could be used for a variety of specific biological tasks. Finally some compounds have been assigned
-to a small number of FTC categories; in such cases little is known or reported about their pharmacology and repurposing opportunities
-might be limited.}\label{fig:03}
+\begin{figure*}[!tpb]%figureS3
+\centerline{\includegraphics{figS4.png}}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but with a hierarchical clustering applied on the top of it, based on the manhattan distance.
+The dendrogram reflects the original structure of the FTC: Functionaly similar compounds being grouped together.
+This figure can be used for further analysis over functional clusters of drugs.}\label{fig:S04}
 \end{figure*}
 
-\subsubsection{Drugs with similar MoAs have similar indications}
-The list of MoAs attributed to a drug can be exploited as a descriptor for the therapeutic agent:
-The tree structure of the FTC can be used to derive some similarity metrics over the MoAs. The underlying 
-heuristic is to assume that the closer two entities are in the taxonomy, the more similar they are. 
-We used a straightforward approach derived from the Jaccard index (see supplementary section 7) in order to compare 
-approved drugs based on the similarity of their MoAs. For instance, the similarity between two compounds present in 
-the same FTC category is 1 (maximum). The similarity between an \emph{anti-blood coagulant} and \emph{pro-blood coagulant} is 0.29, 
-reflecting the fact that such compounds are not similar in regards to the outcome of their biological effect. As the MoA is 
-intuitively expected to be the central concept leading to the indication of the drug, we expected that on average, drugs with 
-similar MoAs would be indicated towards similar therapeutic areas. The heat map presented in Figure \ref{fig:04} shows a pairwise 
-comparison of all the drugs of the FTC based on their relative MoA similarity. The compounds are further grouped by 
-therapeutic indications as defined by the ATC.
-The heatmap (Figure \ref{fig:04}) reveals some square patches around the central diagonal;
-the overall similarity appears higher when compounds from the same ATC group are considered. 
-A significance analysis (details covered in supplementary section 8)
-revealed that the average MoA similarity of compounds belonging to the same ATC category is 
-significantly higher than when compounds belonging to different categories are compared. Indeed, for 
-each category, the p-value was inferior to 0.05 based on 20'000 random permutations over the similarity values. 
-This result supports the idea that drugs with similar MoAs have similar indications. Note that the mean of the 
-similarity values was considered for the statistical analysis; some outliers are also present in the map, which 
-can be interpreted as repurposing hypotheses. These outliers have indeed similar MoAs, yet they belong to totally 
-different therapeutic areas and are used for different purposes according to the ATC. We are currently further analysing 
-such cases in a systematic fashion. Supplementary Figure 3 present similar association behavior when two levels of the ATC are 
-considered (no statistical significance performed). Supplementary Figure 4 re-uses the same data as Figure \ref{fig:04} (one ATC level) 
-but with a clustering function apply to it (hierarchical clustering - manhattan distance) in order to reveal 
-functional clusters of drugs. Finally, Supplementary Figure 5 shows the distribution when compounds are sorted based on 
-their identifiers; no patterns are identifiable in this case. Taken together, these results emphasize that the MoAs as 
-defined in the FTC are indeed on average associated with the therapeutic indication of a drug. This result supports the validity of the
-resource and its potential to computationally address indication discovery.
-
-\begin{figure*}[!tpb]%figure4
-\centerline{\includegraphics{fig4.png}}
-\caption{Pairwise comparison of MoAs similarities. Only the first level of 
-the ATC is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
-The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black). Drugs are grouped according 
-to their first ATC level (colours on the side). For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
-which appears as ``B'' (dark orange) on the plot.
-Some compounds belong to multiple ATC categories (Mulitple - black) and some others do not have
-an ATC code (NoCategory - white)
-The average similarity of drugs present in the same therapeutic category 
-is significantly higher on average when separately compared to all other indications.}\label{fig:04}
+\begin{figure*}[!tpb]%figureS5
+\centerline{\includegraphics{figS5.png}}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but the compounds have been sorted horizontally and vertically only based on their identifiers, which
+we considered as equivalent to a pseudo-random sorting. In this case
+no visual pattern is observable, the similarity values appear homogeneously distributed (no statistical analysis performed).}\label{fig:S05}
 \end{figure*}
-
-\subsection{Drug repurposing hypotheses for Alzheimer's disease}
-In this section we provide examples how actual drug repurposing hypotheses can be derived from the FTC. 
-The approach presented here makes use of the FTC categories as analogs to compartments of a toolbox helping to 
-find drugs to treat Alzheimer's disease. Five FTC categories containing drugs are directly related to the biological 
-processes of the neurodegenerative condition: \emph{Anti-amyloid precursor protein biosynthetic process agent} (FTC\_A0042983), 
-\emph{Anti-tau protein kinase activity agent} (FTC\_A0050321), \emph{Anti-tau protein binding agent} (FTC\_A0048156), 
-\emph{Anti-beta amyloid 
-binding agent} (FTC\_A0001540) and finally \emph{Pro-beta amyloid binding agent} (FTC\_P0001540). We have then considered the drugs present 
-inside each of these classes as potential candidates. Figure \ref{fig:05} shows these drugs, which have been further manually grouped 
-based on the overall similarity of their actions (numbers on Figure \ref{fig:05}).
-
-\begin{figure}[!tpb]%figure5
-\centerline{\includegraphics{fig5.png}}
-\caption{FTC categories describing some of the modes and mechanisms of 
-action that could impact Alzheimer's disease (letters on figure). The categories have been manually picked on the basis that they could directly affect
-the dementia. Drugs classified in these FTC categories further 
-manually grouped based on their MoAs similarities (numbers on figure).}\label{fig:05}
-\end{figure}
-
-The subgroups (1), (2) and (3) are inhibitors of the cholinergic system and some of them, such as \emph{galantamine} (DB00674), are already 
-investigated to treat Alzheimer's disease and other related dementias. This class of agent is in line with the 
-cholinergic hypothesis \citep{Francis1999}, stating that Alzheimer's disease could be caused by dysfunctions in the processing of the
-acetylcholine. The subgroup (4) is exclusively composed of barbiturates (central nervous system depressants). 
-The presence of this pharmacological class of compounds as an Alzheimer's disease treatment is more surprising, as very little 
-literature reports on it. Further investigations reveal that the neuronal acetylcholine receptor subunit alpha-7, a 
-common off-target of barbiturates, binds beta-amyloids with high affinity \citep{Wang2000}. As beta-amyloids are themselves strongly involved 
-in the pathology, barbiturates could affect the state of the condition, similarly to cholinergic inhibitors \citep{Wang2000}.
-The group (B) contains four compounds; \emph{nicotine} and \emph{varenicline} have been further grouped together because of 
-similar pharmacology (group 5).
-\emph{Nicotine} has been shown to improve some of the symptoms of Alzheimer's disease \citep{Jones1992}, it is therefore expected to 
-find this molecule in the predictions. \emph{Varenicline} possess a pharmacology related to that of \emph{nicotine}, which would explain 
-the presence of the drug in this category. The two remaining drugs of group F are respectively \emph{pralidoxime} and \emph{dipivefrin}; 
-little information is available regarding their potential action against the condition, yet these compounds seem linked to the 
-cholinergic hypothesis and could be considered for experimental testing.
-The groups (C) and (D) contain respectively one molecule each. These compounds have been classified as agents perturbing some of 
-the physiological function of the Tau protein, key actor in Alzheimer's 
-disease \citep{Grundke-Iqbal1986}. \emph{Vorinostat} (group C) is currently 
-indicated for the treatment of cutaneous manifestations in patients with T-cell lymphoma, yet a study has 
-shown in vivo (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
-deficit \citep{Kilgore2010}. The presence of \emph{lithium} (group D) was 
-unexpected and suspected of being a false positive at first glance. However a recent study demonstrated a long-term 
-protective effect for the ion in regards to Alzheimer's disease \citep{Young2011}. The last group (E) contains 
-\emph{ezetimibe} and \emph{hesperetin}. These two compounds are primarily used as cholesterol lowering agents (statins). As cholesterol 
-metabolism in the brain appears to be related to dementia, statins are believed to prevent or improve the symptoms of the 
-patients. Although early studies \citep{Wolozin2004} have failed to clearly show a beneficial effect, the investigation is still open.
-From the examples briefly presented above, reported and confirmed by the literature, the FTC appears to be suitable to 
-identify real repurposing hypotheses tailored to a disease. Correctly 
-identifying MoAs of interest helps to retrieve the compounds which might impact 
-the treatment of a condition.
-
-\section{Discussion}
-
-The FTC is a novel classification for approved drugs, which can be used as a starting point to 
-generate drug repurposing hypotheses. We have guided the reader through the various ways the resource 
-can be explored. No experimental validations have been presented in this manuscript, yet references to 
-supporting literature was provided in order to legitimate the claims. The main goal of this document is to 
-introduce the resource to the community; deeper analyses are being performed over the classification at the time of writing.
-
-\subsection{Biological assumptions}
-An asset of the FTC is its ability to handle efficiently categorical data: Classes and relationships 
-are accurately defined, in order to classify compounds based on the semantics of their relations. The properties 
-linking drugs to their respective protein targets (positive and negative perturbations) are however simplistic. At the 
-time being, no consideration is given regarding the binding strength between the drug and the proteins, yet it is a key 
-factor to derive potent and specific activities in the human body. This is also the case for other types of numerical data, 
-such as the dosage; the FTC can predict a role for a drug, yet it cannot provide any information about the concentration or 
-the administration route necessary to obtain the potential effects. The current relations between targets and their involvement 
-in biological processes are also not a fully accurate representation of the biological phenomenon. In a cell, specific domains 
-of the protein could mediate different functions. Only one of such activity types can sometimes be inhibited 
-by a drug (cite felix paper), yet we are assuming in the FTC that as long as a drug affects a protein, 
-it can therefore alter all its known functions.
-These limitations come from the semantics behind the axioms structuring the classification, themselves based on the 
-information available from the databases (e.g. GOA, DrugBank, etc...). Despite entailing not entirely accurately the 
-biochemical reality, the axioms help to generate a larger number of hypotheses, the primary goal of the FTC. The dosage 
-issue is partially addressed by the regulator pattern (see section in Availability and Implementation): It should be 
-easier to experimentally adjust the concentration of the compounds classified as pro or anti biological process agents 
-in order to modulate a physiological effect.
-The predictions generated by the FTC depend on the resolution of the curated information released by 
-the original data providers. Erroneous or missing information will lead to misclassification by the reasoner. 
-Some expected outcomes are also missing from the predictions; sildenafil for instance was expected to be classified 
-as pro-penile erection agent (FTC\_A0043084), yet the lack of appropriate GO annotation prevents it. After discussion 
-with the GOA curation team, a manual annotation can only be asserted based on published experimental results. No document 
-was found to support the involvement of the cGMP-specific 3',5'-cyclic phosphodiesterase (O76074 - sildenafil’s main target) 
-in the negative regulation of penile erection (GO:0060407), therefore no annotation can be made. Further work could be done 
-in this direction, by trying to automatically infer more annotations or by using the electronically generated ones, in order 
-to generate broader yet potentially less plausible repurposing hypotheses.
-
-\subsection{Interpreting the evaluation}
-The high recall value (88\%) supports the idea behind the automated build of the FTC: The data from different 
-repositories funded and curated in parallel, can be integrated to automatically create a new resource. This new resource (FTC) 
-contains most of the known information present in an external gold standard (ATC) and relies on description logics to 
-leverage the native information. Here we have assessed the content of the FTC against the ATC, knowing that these two 
-taxonomies have diverging goals. During the evaluation, equivalences have been manually asserted between classes, 
-which are assumed to have fairly similar meaning and containing similar sets of compounds. These manual assertions 
-are however a weakness, as they are themselves not evaluated (free parameter). The presence or absence of a link was 
-determined only by one curator and any mistake can influence consequently the recall and precision values. In this regards, 
-the evaluation should be considered more as a safety control rather than a formal assessment of a predictive method.
-The false positives derived from the evaluation can also be considered as drug repurposing hypotheses: These drugs 
-can indeed be interpreted as suitable for the ATC category, yet not indexed as such. However, these predictions 
-should be interpreted with caution, as it is currently impossible to distinguish a false positive from a 
-reprofiling opportunity. These considerations do not interfere with the repurposing predictions generated 
-based on semantic similarities or discrete categories as presented in the Exploration section. Finally, 
-note that the ATC/FTC equivalences are open and editable online, any modification will be automatically 
-incorporated in the next release of the resource. It is also possible to evaluate the FTC against a different 
-taxonomy, like the Medical Subject Headings for example, which can be subject to future work.
-
-\section{Conclusion}
-This manuscript introduces the FTC, a free, open and extensible resource, which can assist drug repurposing 
-initiatives or enhance computational studies making use of the concept of \emph{mode of action}. The construction of 
-the classification relies on description logics as a means to define the mode and mechanism of action of approved 
-drugs. We shown the validity of the approach by comparing the content of the FTC to another well implemented taxonomy, 
-the ATC. We further illustrate the tight relationship between the MoA and the indication of a drug and shown how the 
-resource can serve as a starting point to formulate hypotheses for Alzheimer's disease. The work leverages the 
-semantics of distinct databases, working in parallel on different thematics. The platform will be further used 
-to generate predictions in a systematic fashion, which can then be experimentally tested in the laboratory for validation.
-
-\section{Availability and Implementation}\label{impl}
-The code behind the creation of the resource is entirely open and available 
-at {{https://github.com/loopasam/ftc}}. The web application built on the top of the FTC using the Brain library \citep{Croset2013}, and can be 
-find at {{https://www.ebi.ac.uk/chembl/ftc}}. The documentation can be accessed at {{https://github.com/loopasam/ftc/wiki}}. The full
-description of the methodology used to generate the classification is available as supplementary material on Bioinformatics website.
-
-\section*{Acknowledgement}
-Text Text Text Text Text Text  Text Text.  \citealp{Boffelli03} might want to know about  text text text text
-
-\paragraph{Funding\textcolon} Text Text Text Text Text Text  Text Text.
 
@@
 
-%\begin{thebibliography}{}
-%\bibitem[Bardet, 1920]{Bar20}
-%Bardet, G. (1920) Sur un syndrome d'obesite infantile avec
-%polydactylie et retinite pigmentaire (contribution a l'etude des
-%\end{thebibliography}
 \end{document}

### 2013-08-22 document.tex
added some random file for reference template
Signed-off-by: loopasam <samuel.croset@gmail.com>

--- 2013-08-22 document.tex
@@
 \title[The Functional Therapeutic Chemical Classification System]{The Functional Therapeutic Chemical Classification System}
-\author[Croset \textit{et~al}]{Samuel Croset\,$^{1, \footnote{to whom correspondence should be addressed}}$, John Overington\,$^{1}$ and Dietrich Rebholz-Schuhmann\,$^{1}$
+\author[Croset \textit{et~al}]{Samuel Croset\,$^{1, \footnote{to whom correspondence should be addressed}}$, John P. Overington\,$^{1}$ and Dietrich Rebholz-Schuhmann\,$^{1}$
 }
-\address{$^{1}$European Molecular Biology Laboratory
+\address{$^{1}$European Molecular Biology Laboratory, 
 European Bioinformatics Institute (EMBL-EBI), Wellcome Trust Genome Campus, Hinxton, Cambridge CB10 1SD,
@@
 We present here a novel method based on knowledge integration using semantic technologies, 
-to capture the functional role of approuved chemical compounds.
+to capture the functional role of approved chemical compounds.
 \section{Results:}
@@
 Compound's polypharmacology is 
-well represented in the FTC, as shown by the numerous MoAs each approuved drug can exhibit.
+well represented in the FTC, as shown by the numerous MoAs each approved drug can exhibit.
 

### 2013-10-08 reply-to-editors.txt
some modifs after review

Reviewer: 1
Comments to the Author
This manuscript describes a classification system for automatic creation MoA annotations and generation of drug repurposing hypotheses. The authors have developed an online database and released its source code to disseminate this data. The article is clearly laid out and well explained. The approach is novel and should be useful both to the bioinformatic community and a wider scientific audience, but several issues certainly require further clarification.

Major revisions
1) The authors describes FTC is built by integrating information coming from various sources and using ELK algorithm to classify the knowledge. However, I cannot find any clear and detail introduction in manuscript or supplement. This should be explained in detail, being one of the most distinctive features of the paper.

2) Result section 2, 2.1 is mostly methods benchmarking. I suggest those section should be renamed as “methodology” section. In addition, the detail classification description in supplementary (i.e. section 3, 3.1 and 3.2) should be wrote at manuscript. It would make it clearer how the methods are developed and tested.

3) In Figure 1, part C (Incorporation of databases), authors show ‘Prothrombin’ (P00734) is involved in ‘positive regulation of blood coagulation’ (GO_0030194). I tried to check this information from UniProt, but I cannot find any description there (http://www.uniprot.org/uniprot/P00734). How is this relationship module structured? This should be explained in detail.

4) How much of the different results between ATC and FTC? It should be listed and accessed via web.

Minor points:
1) Authors used ATC corpus to evaluate proposed FTC algorithm. The detailed evaluation results were shown on project’s web. I suggest the results should be listed at manuscript, not only at web.
2) 4 Gb of memory --> 4 GB of memory

Reviewer: 2
Comments to the Author
The authors present a taxonomy, the Functional Therapeutic Chemical Classification System, and a knowledge base to assist drug repurposing analyses. The mode of action of a drug, represented by Gene Ontology derived terms, forms the basis for the drug similarity. The authors evaluate their taxonomy by comparing it against the Anatomical Therapeutic Chemical Classification System, and show the usefulness of their system by applying it in a use case to find new drugs for Alzheimers disease.

The knowledge base and taxonomy is very well-documented, but requires deep knowledge about both the system and description logic in general in order to use. Also, presenting a single example use case with only literature evidence to back up the results is not very convincing for a research paper.

Major
Please show that the system can reproduce one known case of drug repurposing.

The authors should describe differences with their system to related efforts such as work by Hoehndorf, Dumontier, and Gkoutos. Bioinformatics 28.16 (2012): 2169-2175, and the Open PHACTS discovery platform http://www.openphacts.org/. A reference to a recent review paper such as the one by Andronis et al in Brief Bioinform. 2011 Jul;12(4):357-68 would be helpful as well.

In addition, the following specific revisions are needed:
p. 2, line 39: “In total there are 1280 approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets.” I looked up the current Drugbank figures: the database contains 6811 drug entries including 1528 FDA-approved small molecule drugs, 150 FDA-approved biotech (protein/peptide) drugs, 87 nutraceuticals and 5080 experimental drugs. It is therefore not clear to me where the number 1280 comes from. Later the authors talk about “approved” drugs. Do they refer to FDA-approved?

p. 4 line 47: “Note that the mean of the similarity values was considered for the statistical analysis; some outliers are also present in the map, which can be interpreted as repurposing hypotheses.” Is this the only interpretation?

p. 4 line 51: “Five FTC categories containing drugs are directly related to the biological processes of the neurodegenerative condition:..” It seems like a user needs to know a lot of details about the system with > 23000 categories in order to use it. How can the user be helped here?

Minor

p. 2 line 26: “Drugs are further assigned into MoA categories based of the OWL constructs and axioms defined in the FTC (see Availability and Implementation).” Provide an example here.

p. 5 line 52: “The presence of lithium (group D) was unexpected and suspected of being a false positive at first glance.” Please explain why this result was unexpected.

Supplemental material:
“The web application built on the top of the FTC can be find …” should be “FTC can be found”

“The FTC implementation relies mostly on Brain [Croset et al., 2013] and the web application builds on the top of the Play! framework”. Please provide reference for the Play! Framework.

### 2013-10-08 document.tex
some modifs after review

@@
 It takes around 5 seconds on a modern 
-laptop (4 cores processor with 4 Gb of memory) to classify the knowledge base using the ELK reasoner \citep{Kazakov2011}. 
+laptop (4 cores processor with 4 GB of memory) to classify the knowledge base using the ELK reasoner \citep{Kazakov2011}. 
 Other OWL reasoners have been tried (e.g, Hermit, Pellet, etc.) but none of them were suitable for this purpose; 

### 2013-10-08 app/controllers/Security.java
new design on the way

@@ -50,15 +50,18 @@ public class Security extends Controller {
 		}
 
 		//The user is supposed to let access to it's information
-		String urlGoogleOAuth = "https://accounts.google.com/o/oauth2/auth?" +
-				"scope=https%3A%2F%2Fwww.googleapis.com%2Fauth%2Fuserinfo.email+" +
-				"&state=%2Fprofile&redirect_uri=http%3A%2F%2F" +
-				"localhost:9000%2Fchembl%2Fftc%2Fauth" +
-				"&response_type=code" +
-				"&client_id=1009954177381.apps.googleusercontent.com" +
-				"&approval_prompt=auto";
+//		String urlGoogleOAuth = "https://accounts.google.com/o/oauth2/auth?" +
+//				"scope=https%3A%2F%2Fwww.googleapis.com%2Fauth%2Fuserinfo.email+" +
+//				"&state=%2Fprofile&redirect_uri=http%3A%2F%2F" +
+//				"localhost:9000%2Fchembl%2Fftc%2Fauth" +
+//				"&response_type=code" +
+//				"&client_id=1009954177381.apps.googleusercontent.com" +
+//				"&approval_prompt=auto";
 		//The response is redirected on the auth() method.
-		redirect(urlGoogleOAuth);
+//		redirect(urlGoogleOAuth);
+		
+		//TODO just for debug, uncomment for prod and delete line
+		auth("");
 	}
 
 
@@ -66,40 +69,42 @@ public class Security extends Controller {
 	public static void auth(String code) {
 
 		//Depending on what the user entered, could be an error
-		if(code == null){
-			redirect("Application.index");
-		}
+//		if(code == null){
+//			redirect("Application.index");
+//		}
 
 		//call for access token
-		HttpResponse res = WS.url("https://accounts.google.com/o/oauth2/token").
-				setParameter("code", code).setParameter("client_id", "1009954177381.apps.googleusercontent.com").
-				setParameter("client_secret", "ClzAiwp6nuTvRqbQpkMq4GS7").setParameter("redirect_uri", "http://localhost:9000/chembl/ftc/auth").
-				setParameter("grant_type", "authorization_code").post();
+//		HttpResponse res = WS.url("https://accounts.google.com/o/oauth2/token").
+//				setParameter("code", code).setParameter("client_id", "1009954177381.apps.googleusercontent.com").
[…]

### 2013-10-09 document.tex
some modifs done

@@
 (C) The content of various databases in incorporated and linked using the FTC specific properties. (D) Finally a reasoner runs over the knowledge
-base and assigns drugs to MoA categories based on whether or not a definition can be satisfied. The taxonomic structure of the
+base and assigns drugs to MoA categories based on whether or not a definition can be satisfied. For example here, the drug
+\emph{ximelagatran} will be assigned as member of the category \emph{Anti-blood coagulation agent} because of the logical
+links \emph{ximelagatran negatively-perturbs prothrombin} and \emph{prothrombin involved-in positive regulation of blood coagulation}.
+The taxonomic structure of the
 FTC appears also in the reasoning step, from the entailements of the equivalent defintions.}\label{fig:01}
@@
 most drugs fail because of safety or efficacy issues.
-There is an increasing number of approaches to predict repurposing opportunities using computational methods \citep{Sanseau01072011}. Most 
+There is an increasing number of approaches to predict repurposing opportunities using computational methods (see \cite{dudley2011exploiting} or \cite{andronis2011literature} for recent reviews).
+Most
 methods operate on the profiles of physicochemical descriptors derived from molecular structures \citep{Haupt2011}. Other methods characterise 
@@
 Drugs are further assigned into MoA categories based of the OWL constructs and axioms defined in the 
-FTC (see \emph{Availability and Implementation}). A reasoner, a program capable of understanding such axioms, performed this task. 
-The process to build the FTC is summarised in Figure~\ref{fig:01}.
+FTC. A reasoner, a program capable of understanding such axioms, performed this task (see \emph{Availability and Implementation} for details). 
+The process to build the FTC is summarised in Figure~\ref{fig:01} alongside an example.
 It takes around 5 seconds on a modern 
@@
 The reader can browse and access the content of the FTC online at {{https://www.ebi.ac.uk/chembl/ftc}}.
-In total there are 1280 approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets. 
+In total there are 1280 FDA-approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets.
+Only drugs acting on at least one human protein target were selected.
 The FTC introduces 23'353 new categories describing the mode and mechanism of action of therapeutic compounds. 4289 of these 
@@
 similarity values was considered for the statistical analysis; some outliers are also present in the map, which 
-can be interpreted as repurposing hypotheses. These outliers have indeed similar MoAs, yet they belong to totally 
+can be interpreted as repositioning hypotheses. These outliers have indeed similar MoAs, yet they belong to totally 
 different therapeutic areas and are used for different purposes according to the ATC. We are currently further analysing 
-such cases in a systematic fashion. Supplementary Figure 3 present similar association behavior when two levels of the ATC are 
+such cases in a systematic fashion. Hypotheses have to be manually examined and interpreted, as ATC categories are
+only covering some of the legal usage of the drugs. We expect to find off-label indications in 
+the predictions for instance, as well as false positives.
+
+Supplementary Figure 3 present similar association behaviour when two levels of the ATC are 
 considered (no statistical significance performed). Supplementary Figure 4 re-uses the same data as Figure \ref{fig:04} (one ATC level) 
@@
 shown in vivo (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
-deficit \citep{Kilgore2010}. The presence of \emph{lithium} (group D) was 
-unexpected and suspected of being a false positive at first glance. However a recent study demonstrated a long-term 
+deficit \citep{Kilgore2010}. 
+The presence of the \emph{lithium} (group D) was 
+confirmed by a recent study demonstrating a long-term 
 protective effect for the ion in regards to Alzheimer's disease \citep{Young2011}. The last group (E) contains 

### 2013-10-09 document.tex
stop reviews

@@
 categories to which the drug belongs with an explicit and clear meaning. These discrete MoAs are a good starting 
-point to understand what a compound can do when administered in a human system, as it will be shown with Alzheimer's disease.
-The human readable definitions 
-associated to each FTC category furthermore help the user to choose the adequate MoA for the desired outcome.
+point to understand what a compound can do when administered in a human system.
 Compound's polypharmacology is 
 well represented in the FTC, as shown by the numerous MoAs each approved drug can exhibit.
+We decided to further look at a well-known repositioning example, 
+in order to see whether the FTC was suitable to identify the new
+uses of an old drug. We picked the \emph{thalidomide} for 
+this exercise ({{https://www.ebi.ac.uk/chembl/ftc/agent/DB01041}}). The molecule was 
+first indicated to treat morning sickness
+in pregnant women, but got quickly withdrawn from the market after the dramatic identification 
+of teratogenic effect in newborns, caused by the drug. The accepted molecular mechanism behind the side-effect is  
+an impairment of the angiogenic process responsible for the development of members, affecting in particular the limbs \citep{therapontos2009thalidomide}.
+The \emph{thalidomide} was accurately classified as \emph{Anti-cell migration involved in sprouting angiogenesis agent} in the FTC, capturing the known toxicity of the drug.
+The \emph{thalidomide} is currently investigated for a multitude of new usages,
+in particular for anti-cancer and immunomodulatory activities among others \citep{teo2005thalidomide}.
+These new indications are well represented in the FTC too, for 
+example by the categories \emph{Anti-vascular endothelial growth factor production agent} or 
+\emph{Anti-cell division agent} for neoplastic activities, 
+or by the classes \emph{Anti-cytokine secretion agent} and \emph{Anti-I-kappaB kinase/NF-kappaB cascade agent} for its effect
+on the immune system. These observations demonstrate that the FTC can successfully capture the 
+molecular reasons behind the repositioning of an old compound. 
+Moreover the classification can also provide valuable insight regarding
+potential toxicity too.
 
@@
 introduce the resource to the community; deeper analyses are being performed over the classification at the time of writing.
+This classification leverages the information present in various databases and ontologies, similarly to the 
+Open PHACTS initiative \citep{williams2012open} and to the work done by \cite{hoehndorf2012identifying}. The FTC
+mostly differentiates itself from these projects by providing a whole new set of novel categories, 
+dedicated to tackle a very specific problem: drug repositioning.
 

### 2013-10-09 app/build/Builder.java
some more boostrapping on the way

@@ -15,6 +15,8 @@ import play.test.Fixtures;
 import uk.ac.ebi.brain.error.BrainException;
 
 public class Builder {
+	
+	//TODO create a zip file method for download here.
 
 	public void createTmpStructure() {
 		Logger.info("Creating the temporary directory structure...");

### 2013-10-09 app/controllers/Application.java

@@ -210,6 +210,8 @@ public class Application extends Controller {
 	//Performs the query
 	public static void query(String query) throws NonExistingEntityException{
 
+		Logger.info(query);
+		
 		try {
 			//Checks whether the query is parsable
 			brain.parseLabelClassExpression(query);

### 2013-10-09 app/models/OwlResult.java

@@ -8,6 +8,8 @@ import javax.persistence.Entity;
 import javax.persistence.Lob;
 import javax.persistence.MapKeyColumn;
 
+import org.hibernate.annotations.Type;
+
 import play.db.jpa.Model;
 
 /**
@@ -19,6 +21,7 @@ import play.db.jpa.Model;
 public class OwlResult extends Model {
 
 	@Lob
+	@Type(type = "org.hibernate.type.TextType")
 	public String query;
 	
 	public int numberOfTimes;

### 2013-10-10 document.tex
core article revised

@@
 Open PHACTS initiative \citep{williams2012open} and to the work done by \cite{hoehndorf2012identifying}. The FTC
-mostly differentiates itself from these projects by providing a whole new set of novel categories, 
-dedicated to tackle a very specific problem: drug repositioning.
+mostly differentiates itself from these projects by providing a whole set of new categories on the top of the integrated
+information, dedicated to tackle a very specific problem: drug repositioning.
 

### 2013-10-10 document.tex
supps revisions on the way

--- 2013-10-09 document.tex
@@
 
-\title[The Functional Therapeutic Chemical Classification System]{The Functional Therapeutic Chemical Classification System}
-\author[Croset \textit{et~al}]{Samuel Croset\,$^{1, \footnote{to whom correspondence should be addressed}}$, John P. Overington\,$^{1}$ and Dietrich Rebholz-Schuhmann\,$^{1}$
-}
-\address{$^{1}$European Molecular Biology Laboratory, 
-European Bioinformatics Institute (EMBL-EBI), Wellcome Trust Genome Campus, Hinxton, Cambridge CB10 1SD,
-United Kingdom\\}
+\title[Supplementary information - The Functional Therapeutic Chemical Classification System]{Supplementary information - 
+The Functional Therapeutic Chemical Classification System}
 
-\history{Received on XXXXX; revised on XXXXX; accepted on XXXXX}
+\section*{Availability and Implementation}
 
-\editor{Associate Editor: XXXXXXX}
+This section describes the building mechanism behind the FTC.
+First, a list of categories describing the mode and mechanism of action of drugs is defined. 
+Then in a second step the newly created categories are automatically populated with approved compounds. 
+Finally, the FTC is evaluated and repositioning hypotheses can be generated.
 
-\maketitle
+\section{Source code}
+The code behind the creation of the resource is entirely open and available 
+at {{https://github.com/loopasam/ftc}}. The web application built on the top of the 
+FTC can be found at {{https://www.ebi.ac.uk/chembl/ftc}} and the documentation can be 
+accessed at {{https://github.com/loopasam/ftc/wiki}}. The reader should be familiar with 
+description logics and the Web Ontology Language (OWL) to fully understand the construction of the 
+knowledge base. An introduction to description logics from the perspective of the biomedical scientist is 
+available on the wiki at {{https://github.com/loopasam/ftc/wiki/Description-Logics}}. We also refer the readers non-familiar
+with description logics and reasoning algorithms to excellent introductory text by \cite{krotzsch2012description}.
+The FTC implementation 
+relies mostly on Brain \citep{Croset2013} and the web application builds on the top of the Play! framework \citep{playframework}. 
+Classification tasks use the ELK reasoner \citep{Kazakov2011}. The computer hosting the web application has 8 Gb of memory 
+with 4 processors, this architecture allows fast parallel reasoning, thanks to ELK's design. More functionalities 
+will be added to the web application following user requirements (so called \emph{lean implementation}).
 
-\begin{abstract}
+\section{FTC categories creation}
+The mode of action categories present in the FTC are defined based on the terms coming from the 
+Gene Ontology (GO) \citep{Ashburner2000}. Both the molecular function and biological process branches are used for 
+this purpose, yet handled slightly differently.
 
-\section{Motivation:}
-Drug repurposing is the discovery of new indications for compounds already approved and used in a clinical setting. 
-Recently, some computational approaches have been developed to unveil new opportunities in a systematic fashion, by taking 
-into consideration gene expression signatures or chemical features for instance. 
-We present here a novel method based on knowledge integration using semantic technologies, 
-to capture the functional role of approved chemical compounds.
-\section{Results:}
-In order to computationally generate repurposing hypotheses, we used the Web Ontology Language (OWL) to 
-mathematically define the meaning of over 20'000 terms formally describing various modes and mechanisms of action. Based on an 
-integration of public data, we have automatically assigned over a thousand of approved drugs into 
-these MoA categories. The resulting new resource is called the Functional Therapeutic Chemical Classification System (FTC) 
-and was further evaluated against the content of the traditional Anatomical Therapeutic Chemical Classification System (ATC). 
-We illustrate how the new classification can be used to generate drug repurposing hypotheses, using Alzheimer’s disease as a use-case.
+\subsection{Categories related to biological processes}\label{catprocess}
+All the biological processes featured by the GO are looked-up one by one. All the 
+time a process is linked to another process (\emph{X}) via a \emph{positive} or \emph{negative regulation} link, two FTC 
+classes are created: \emph{Anti-X agent} and \emph{Pro-X agent}. For instance the GO term \emph{positive regulation of blood 
+coagulation} is linked to the term \emph{blood coagulation} via a \emph{positively regulates} relation, therefore two FTC 
+categories \emph{Anti-blood coagulation agent} and \emph{Pro-blood coagulation agent} are created. The identifiers of the 
+new FTC classes are also derived from the GO term used to create the class pattern. The GO numeric identifier 
+is re-used and the letter \emph{A} or \emph{P} is appended before to emphasize the \emph{anti} or \emph{pro} pattern. 
+From the example presented 
+previously, the FTC class \emph{Anti-blood coagulation} has FTC\_A0007596 as identifier, because the GO term \emph{blood coagulation} is 
+referenced by GO:0007596. Following the same logic, FTC\_P0007596 is the identifier of the class \emph{Pro-blood coagulation agent}. The 
+design choice for identifiers and labels allows the FTC to fully rely on the high quality work provided by the GO curation 
+team and scale over it.
 
-\section{Availability:} A web application built on the top of the resource is freely available at {{https://www.ebi.ac.uk/chembl/ftc}}. The 
-source code of the project is available at {{https://github.com/loopasam/ftc}}.
+\subsection{Categories related to molecular functions}\label{catfunc}
+The mode and mechanism of actions related to molecular functions are created in the following manner: All the time 
+a molecular function (\emph{Y}) is encountered then two FTC categories are created, as for the processes: \emph{Anti-Y agent} and 
+\emph{Pro-Y agent}. The identifiers are assigned the same way as described before. For instance, out of the GO term \emph{androgen 
+receptor activity} (GO:0004882) two FTC classes are derived: \emph{Pro-androgen receptor activity agent} (FTC\_P0004882) and 
+\emph{Anti-androgen receptor activity agent} (FTC\_A0004882).
 
-\section{Supplementary Information:}
-Supplementary data are available at Bioinformatics online.
+\section{Equivalent definitions}
+FTC classes are generated as presented in the previous section. Up to this point, these categories 
+are only tokens with a human readable label as well as an identifier. The next step is going to 
+assign equivalent definitions to each FTC class. A reasoner can understand such definitions and will automatically 
+classify the knowledge base accordingly; drugs will then be assigned to FTC categories and the taxonomic structure 
+arises after this reasoning step. Equivalent definitions are written as OWL class expressions using the entities of 
+the knowledge base (summarised at {{https://github.com/loopasam/ftc/wiki/Knowledge-Base}} and in section \ref{specs}).
+There are two types of equivalences: The first one captures perturbation of regulatory 
+biological processes (so called \emph{regulatory patterns}) and the second one handles the perturbed functions (\emph{functional patterns}).
 
-\section{Contact:} \href{croset@ebi.ac.uk}{croset@ebi.ac.uk}
-\end{abstract}
+\subsection{Regulatory pattern}
+Some of the FTC categories are created from the biological processes present in the 
+GO (cf section \ref{catprocess}); these categories have two arbitrary equivalent definitions, representing the 
+two possible ways a compound might impact the biological process. Anti-biological process agent FTC categories 
+contain the drugs that negatively perturb a target involved in the positive regulation of the biological process. 
+The \emph{anti} categories also feature the compounds that positively perturb a negative regulator of the same process. 
+The \emph{pro} categories are equivalent to the opposite pattern. Supplementary Figure \ref{fig:S01} illustrates the equivalent definitions for the 
+FTC class \emph{Anti-blood coagulation agent} (FTC\_A0007596).
 
-\section{Motivation}
+\begin{figure}[!tpb]%figureS1
+\centerline{\includegraphics{figS1.png}}
+\caption{Example of equivalent definitions (gray boxes) for the FTC category \emph{Anti-blood coagulation 
+agent} (FTC\_A0007596). After the data integration step, a reasoner will look the definitions and identify 
+which drugs satisfy the FTC definitions. Definitions are expressed using the Web Ontology Language (OWL), 
+serialised here using the Manchester syntax.}\label{fig:S01}
+\end{figure}
 
-\begin{figure*}[!tpb]%figure1
-\centerline{\includegraphics{fig1.png}}
-\caption{Summary of the methodology used to construct the resource. (A) The FTC categories are first created from the existing GO terms
-describing molecular functions and biological processes. (B) Then, each category has a logical equivalent definition assign to it.
-(C) The content of various databases in incorporated and linked using the FTC specific properties. (D) Finally a reasoner runs over the knowledge
-base and assigns drugs to MoA categories based on whether or not a definition can be satisfied. For example here, the drug
-\emph{ximelagatran} will be assigned as member of the category \emph{Anti-blood coagulation agent} because of the logical
-links \emph{ximelagatran negatively-perturbs prothrombin} and \emph{prothrombin involved-in positive regulation of blood coagulation}.
-The taxonomic structure of the
-FTC appears also in the reasoning step, from the entailements of the equivalent defintions.}\label{fig:01}
+\subsection{Functional pattern}
+The FTC categories generated from the GO molecular functions (cf section \ref{catfunc}) are also equivalent to a logical 
+definition. \emph{Anti} FTC categories dealing with molecular activities are asserted as equals to the drugs 
+that negatively perturb the function. \emph{Pro} categories are equivalent to the drugs that positively perturb the 
+function of interest. A summary of the patterns definitions is available on the 
+online wiki at {{https://github.com/loopasam/ftc/wiki/Mode-of-Action}}.
+
+\section{Data integration}
+At this stage, the knowledge base contains the created FTC classes associated with their 
+logical definitions, as well as the GO and the core FTC entities. The knowledge base is then 
+further populated with some information coming from various public databases. Only manually curated 
+information is considered.
+
+\subsection{DrugBank}
+The DrugBank database is a unique bioinformatics and cheminformatics resource 
+that combines detailed drug data with comprehensive drug target information \citep{Knox2011}. The approved 
+drugs (small molecules and biotherapeutics) acting on proteins are extracted from the database and 
+imported in the FTC knowledge base. In order to be selected, a compound must firstly be approved and 
+secondly have a pharmacological action on at least one human protein target present in Uniprot \citep{TheUniprotConsortium2013}. 
+The protein targets all have at least one manually asserted GO annotation \citep{Dimmer2012} for a biological process or a 
+molecular function. DrugBank links compounds to targets via \emph{actions}. The DrugBank actions are somehow 
+structured and consistent: Concepts such as \emph{inhibitor} or \emph{agonist} are reused throughout the database for 
+example, yet they are not strictly formalised as a controlled vocabulary. These actions are manually 
+standardised to the core properties of the FTC according to their biochemical meaning: For instance the action 
+\emph{antagonist} is mapped to the FTC \emph{negatively-perturbs} property.
+Compounds coming from DrugBank are represented as OWL classes and asserted 
+as subclasses of the class \emph{DrugBank compound} (FTC\_C2). Protein targets are described as OWL classes 
+too and subclasses of the core class \emph{Protein}. Each DrugBank compound is then connected to 
+its target via the following axiom pattern: \emph{[drug] SubClassOf perturbs some [protein]}. 
+E.g. \emph{Ximelagatran SubClassOf negatively-perturbs some Prothrombin}.
+
+\subsection{Gene Ontology annotations}
+The GO annotation program aims to provide high-quality GO annotations to 
+proteins in UniProt \citep{Dimmer2012}. In the context of the FTC, such annotations are used 
+to create axioms linking protein targets to molecular functions and biological processes. 
+Each protein annotated with a function creates an axiom such 
+as \emph{[protein] SubClassOf has-function some [molecular function]}. 
+Each protein annotated to a biological process creates an axiom such 
+as \emph{[protein] SubClassOf involved-in some [biological process]}. 
+E.g. \emph{Prothrombin SubClassOf involved-in some positive regulation of blood coagulation}. 
+Each protein can be involved in multiple processes and capable of realizing multiple functions;
+the polypharmacology is captured at this level.
+
+\section{Knowledge base classification}
+The knowledge base is fully built at this step and contains core classes, mode 
+of actions descriptions alongside the actions of approved DrugBank compounds on 
+protein targets in Uniprot. The proteins are linked to their molecular functions and involvement in 
+biological processes via the GO annotations. The logical specifications of the FTC are there to glue 
+the different data together and to explicitly express the logical links between resources. The FTC 
+knowledge base follows an OWL2 EL profile \citep{Motik2009}, which enable the use of fast and parallelised reasoners such 
+as ELK. During the classification process, the reasoner checks whether the mode of action equivalent 
+definitions are satisfied or not and assigns drugs inside the corresponding FTC categories. The tree structure 
+of FTC appears also at this step from the logical definitions.
+
+\section{Evaluation methodology}
+As the classification of therapeutic agents is done in an automated way, it 
+is important to evaluate the results generated against a known resource which will be 
+considered as gold standard. The assessment of the FTC is done against another similar 
+classification, the Anatomical Therapeutic Chemical Classification System (ATC) \citep{world2000anatomical}). 
+The ATC has been developed to serve as a tool for drug utilization research in order to 
+improve quality of drug use. In this resource, the information is manually curated, and drugs 
+are assigned to categories based of their legally approved indications. The goal of the ATC differs 
+from the one of the FTC, yet the two resources are sharing some very similar concepts, which can be 
+used for the evaluation. Categories of both classifications contain approved drugs with a DrugBank 
+identifier, meaning that some of the drugs indexed in the FTC are also present in the ATC. From that, 
+it is possible to define some evaluation points, which will help to assess the automated classification process.
+
+\subsection{Evaluation Points}
+An evaluation point is defined as an equivalence between a class from the FTC with one or 
+more classes from the ATC. The idea is to look at the set of drugs contained in both side of the 
+equivalence and estimate the overlap, as illustrated in the Supplementary Figure \ref{fig:S02}. Evaluation points are defined 
+by hand and not themselves evaluated. The full list of evaluation points as well as a summary of the results 
+are available online at https://www.ebi.ac.uk/chembl/ftc/evaluation/. Each evaluation point has a series of 
+true/false positive and false negative drugs associated with it.
+
+\begin{figure}[!tpb]%figureS2
+\centerline{\includegraphics{figS2.png}}
+\caption{Example of evaluation point. Classes from the ATC are manually 
+mapped to classes from the FTC. The drugs present in each side of the equivalences 
+are compared, and some metrics are derived from it (false positive/negative and true positive). 
+A summary of the evaluation is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.}\label{fig:S02}
+\end{figure}
+
+\subsection{True Positives}
+Drugs that are present in both the FTC and the equivalent ATC class(es) are called true positives. 
+These compounds reflect that the automated classification was capable of retrieving correctly the 
+information present in the gold standard (ATC).
+
+\subsection{False Negatives}
+These drugs are present in the ATC class(es) but not in the corresponding FTC class. 
+The automated classification failed to retrieve these compounds. The smaller the number of 
+false negatives is, the better the FTC is at recalling drugs. A small number of false negatives 
+means that if a drug is present in the ATC (gold standard), then it is likely that the drug will 
+also be correctly categorised in the FTC.
+
+\subsection{False Positives}
+The false positives are the drugs present in the FTC category of the evaluation point but not in the corresponding ATC classes. 
+A high number of false positives means that the FTC is over-assigning compounds to classes.
+The false positives relates to the accuracy of the classification. In the context of this work, 
+some false positives could also be considered as drug repurposing opportunities.
+
+\subsection{Precision}
+Precision is the probability that a randomly selected drug from the FTC is present in the equivalent ATC classes. 
+The value is standardised as a percentage and corresponds to the formula: True Positive / (True Positive + False Positive).
+
+\subsection{Recall}
+Recall is the probability that a randomly selected drug from the ATC has been assigned to the correct 
+corresponding class in the FTC. The value is standardised as a percentage and corresponds 
+to the formula: True Positive / (True Positive + False Negative).
+
+\section{Semantic similarity}
+The semantic similarity measure performed over the FTC is a derivative of the 
+Jaccard index \citep{jaccard1912distribution} \citep{rogers1960computer} (see \cite{pesquita2009semantic} for review). 
+It is probably best understood as an example: If we consider two classes A and B, the semantic similarity 
+between these classes corresponds to the number of OWL superclasses (direct and indirect, obtained with a reasoner) that 
+are shared by A and B (intersection) divided by the number of superclasses of A or B (union). The index ranges 
+from 0 (totally different) to 1 (identical).
+
+\section{Mode of action similarity against indication}
+A statistical analysis was performed over the data presented on Figure 4. 
+When two compounds are randomly taken, they have on average a higher mode of action similarity when 
+they are assigned to the same ATC category (one ATC level).
+In order to estimate whether this observation was 
+due to chance only, we formulated the following null hypothesis (H0): \emph{For a pair of drug A and B, it does not 
+matter to which ATC category they belong to, their similarity is always average.}
+The alternative hypothesis (H1) was: \emph{For a pair of drug A and B, if A and B have the same ATC code, 
+we expect on average to obtain a higher similarity value than if A and B have different ATC codes.}
+
+A permutation test was then performed for each ATC category. For example, we started with the ATC 
+category A (first row on Figure 4), looked at the similarity values when pairs of compounds both belong 
+to the category A (top right corner square) and compared it to the similarity values when the pair of compounds 
+belong to different categories (A and B, A and C and so forth). For each comparison we obtained two distributions of 
+values (not shown). On average the similarity values are always higher when the two compounds 
+belong to the same category (A/A versus A/B for instance). A permutation test (n = 20'000) was 
+performed in order to see whether this observation was due to chance only. We were able to reject the null 
+hypothesis for a significance level of 0.05 all the times. 
+The choice for a permutation test was driven by the fact that MoA similarity values do not follow any type 
+of standard distribution (data not shown). For the significance levels set, the null hypothesis was rejected for 
+all first level ATC categories.
+
+\section{Knowledge base specification}\label{specs}
+This section presents the scaffold of the knowledge base underlying the FTC. 
+The logic structuring the FTC comes essentially from a set of core OWL properties (rich RBox). 
+Some of these properties originate from the GO. When necessary some new ones have also been introduced. 
+In order to understand how these properties interact, first will be presented the fundamental classes present 
+at the top of the FTC classification.
+
+\subsection{Core classes}
+The high level concepts covered by the FTC are represented as OWL classes and enumerated below. 
+Some of these core classes are coming from external ontologies, in which case the original URI is preserved.
+
+\begin{itemize}
+\item \textbf{Molecular Function} \\
+Identifier: http://purl.obolibrary.org/obo/GO\_0003674 \\
+Label: molecular function \\
+Definition: As defined by the Gene Ontology. Elemental activities, such as catalysis or binding, describing 
+the actions of a gene product at the molecular level. A given gene product may exhibit one or more molecular functions.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Biological Process} \\
+Identifier: http://purl.obolibrary.org/obo/GO\_0008150 \\
+Label: biological process \\
+Definition: As defined by the Gene Ontology: Any process specifically pertinent to the functioning of 
+integrated living units: cells, tissues, organs, and organisms. A process is a collection of molecular 
+events with a defined beginning and end.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Protein} \\
+Identifier: http://purl.uniprot.org/core/Protein \\
+Label: protein \\
+Definition: As defined by Uniprot. Description of a protein.
+Comment: Gene products present inside the FTC are all human proteins. Uniprot URIs are used.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Drug} \\
+Identifier: http://schema.org/Drug \\
+Label: drug \\
+Definition: As defined by schema.org: A chemical or biologic substance, used as a medical therapy, that has a 
+physiological effect on an organism. \\
+Comment: In the context of the FTC, DrugBank chemicals are considered for their role as therapeutic agent 
+rather than for their chemical structure. Mode of action classes are representing agent types and are subclasses of this concept.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Therapeutic Agent} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_C1 \\
+Label: therapeutic agent \\
+Definition: Role of a drug capable of producing a therapeutic effect.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{DrugBank Compound} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_C2 \\
+Label: DrugBank compound \\
+Definition: Drug coming from DrugBank.
+\end{itemize}
+
+\subsection{Core properties}
+The expressivity of the FTC comes mostly for the properties (RBox). Below is a list of the 
+OWL object properties present inside the FTC knowledge base.
+
+\begin{itemize}
+\item \textbf{Part Of} \\
+Identifier: http://purl.obolibrary.org/obo/BFO\_0000050 \\
+Characteristic: Transitive \\
+Label: part-of \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Has Part} \\
+Identifier: http://purl.obolibrary.org/obo/BFO\_0000051 \\
+Characteristic: Transitive \\
+Label: has-part \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002211 \\
+Chained property: regulates o part-of -\textgreater regulates \\
+Label: regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Negatively Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002212 \\
+SubPropertyOf: regulates \\
+Label: negatively-regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Positively Regulates} \\
+Identifier: http://purl.obolibrary.org/obo/RO\_0002213 \\
+SubPropertyOf: regulates \\
+Label: positively-regulates \\
+Definition: As defined and used in the Gene Ontology.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Involved In} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R1 \\
+Label: involved-in \\
+Domain: protein \\
+Range: biological process \\
+Definition: Entails the participation of a protein in a biological process. \\
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Has Function} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R2 \\
+Label: has-function \\
+Domain: protein \\
+Range: molecular function \\
+Definition: Describes the molecular function a protein can realize. \\
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R3 \\
+Label: perturbs \\
+Domain: drug \\
+Range: protein \\
+Definition: Specific biochemical interaction through which a drug substance 
+will affect the activity of a protein (mechanism of action). The property refers to the specific molecular 
+targets to which the drug binds, such as an enzyme or receptor.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Negatively Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R4 \\
+Label: negatively-perturbs \\
+SubPropertyOf: perturbs \\
+Definition: Specific biochemical interaction through which a drug substance will decrease the activity of a protein. 
+The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
+\end{itemize}
+
+\begin{itemize}
+\item \textbf{Positively Perturbs} \\
+Identifier: https://www.ebi.ac.uk/chembl/ftc/FTC\_R5 \\
+Label: positively-perturbs \\
+SubPropertyOf: perturbs \\
+Definition: Specific biochemical interaction through which a drug substance will increase the activity of a protein. 
+The property refers to the specific molecular targets to which the drug binds, such as an enzyme or receptor.
+\end{itemize}
+
+\begin{figure*}[!tpb]%figureS3
+\centerline{\includegraphics{figS3.png}}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but with two ATC levels considered. It is possible to observe some clustering around the diagonal line (no
+statistical analysis performed).}\label{fig:S03}
 \end{figure*}
 
-Drug repurposing is the use of known active compounds for new therapeutic indications \citep{Sanseau01072011}. 
-When administered in a living organism, a compound can indeed play various roles and affect different biological processes; 
-accurately identifying these different functions helps to predict the potential side-effects a drug could have and can also lead to 
-interesting repurposing opportunities \citep{Medina-Franco2013}. For instance, \emph{sildenafil} was initially developed 
-against angina and has been repurposed towards erectile dysfunction during the clinical trials \citep{Ashburn2004}. Approved compounds are 
-attractive because they have been extensively studied and have by definition already successfully passed clinical trials, where 
-most drugs fail because of safety or efficacy issues.
-There is an increasing number of approaches to predict repurposing opportunities using computational methods (see \cite{dudley2011exploiting} or \cite{andronis2011literature} for recent reviews).
-Most
-methods operate on the profiles of physicochemical descriptors derived from molecular structures \citep{Haupt2011}. Other methods characterise 
-the drugs on more abstract levels, such as the gene expression signature \citep{Iorio2010} or 
-via the reported side-effects \citep{Campillos2008}. 
-These approaches 
-have in common to look for similarities within existing drugs and forward similar compounds as repurposing hypotheses.
-A feature of particular interest to describe drugs is the Mode of Action (MoA). According to Wikipedia, 
-the MoA describes \emph{a functional or anatomical change, 
-at the cellular level, resulting from the exposure of a living organism to a substance}. For instance 
-terms such as \emph{transcriptional regulation agent} or \emph{anticoagulant} define MoAs and characterise the 
-roles of a certain type of drugs. The 
-MoA abstracts over the relations between molecular functions, protein targets and drug activities; it is the central concept linking a 
-chemical structure to a set of biological effects. Intuitively, the indication of a drug logically depends on its MoA.
-Despite its widespread use in drug discovery, the MoA has not been used yet as a descriptor for repurposing analyses. One reason for this 
-might be the challenge of formally defining MoAs. Indeed, MoAs are terms or categories, it is not possible to represent 
-them straightforwardly with values and numbers like one can do for a 3D molecular structure or for a gene expression profile. 
-Nonetheless, the meaning of a concept can be formalised with controlled vocabularies and ontologies \citep{Gruber1995}; 
-originating from description logics, such frameworks help to define the mathematical meaning of symbols and strings of characters. 
-In an ontology or knowledge base, concepts are organised and linked following the logical type of relation they have among them. 
-In the Gene Ontology (GO) for example \citep{Ashburner2000}, biological processes and molecular functions terms are manually curated  
-and their meaning specified by the relation types linking two GO terms.
-MoA definitions are present in other classifications such as the Medical Subject Headings \citep{Nelson2004} or the Chemical Entities of 
-Biological Interest \citep{Hastings2012} for example. The Anatomical Therapeutic 
-Chemical Classification System (ATC) \citep{world2000anatomical} also describes to some 
-extent the action of drugs at the anatomical level. All these resources are valuable for the community as a source of carefully 
-manually curated information. Moreover, the categories described in these classification systems are sometimes used to annotate 
-drugs: For instance the compound \emph{sildenafil} has been manually annotated as \emph{vasodilator agent} (CHEBI:35620 or MeSH:D27.505.954.411.918).
-The classifications mentioned previously are not specially designed for drug repurposing; they purposefully report only the 
-well-known and major MoAs of chemical compounds. The pharmacological spectrum of each drug is not necessarily well covered, yet 
-it would be the best way to predict new indications. In our context, an ideal knowledge base would feature the known MoAs of a 
-drug as well as some predicted ones to be tested in experiments. The MoA categories should derive and scale over primary molecular 
-evidences exposed in biomedical databases in an automated way.
-To address the lack of systematic MoA annotations, we have implemented the Functional Therapeutic Chemical 
-Classification System (FTC), presented here in this manuscript. The FTC is automatically built by leveraging the content 
-of various biomedical databases using description logics and automated reasoning. Over 20'000 new MoA categories are defined 
-in the resource and further populated with approved drugs using the Web Ontology Language (OWL) in combination with a reasoner. 
-Most of the drugs are present in multiple FTC categories, reflecting the various roles a compound can play inside a biological system which 
-can serve as starting point for repurposing. 
-The resource was evaluated against the ATC, traditional classification scheme introduced before. We present as well some preliminary 
-analyses over the data, by looking at the relation between the MoA and the indication of a compound using semantic similarity. 
-Finally, we illustrate also how the FTC can be used as a pharmacological toolbox to analyse drug repurposing for Alzheimer’s disease.
-
-\section{Results}
-
-The knowledge base behind the FTC is built by integrating information coming from various sources. 
-The GO terms serve as template to create the FTC categories describing the MoAs; DrugBank \citep{Knox2011} provides the 
-known links between drugs and their protein targets and Uniprot \citep{TheUniprotConsortium2013} maps targets to their 
-respective GO annotations \citep{Dimmer2012}. 
-Drugs are further assigned into MoA categories based of the OWL constructs and axioms defined in the 
-FTC. A reasoner, a program capable of understanding such axioms, performed this task (see \emph{Availability and Implementation} for details). 
-The process to build the FTC is summarised in Figure~\ref{fig:01} alongside an example.
-It takes around 5 seconds on a modern 
-laptop (4 cores processor with 4 GB of memory) to classify the knowledge base using the ELK reasoner \citep{Kazakov2011}. 
-Other OWL reasoners have been tried (e.g, Hermit, Pellet, etc.) but none of them were suitable for this purpose; 
-the classification time was always superior to a few minutes (data not shown). The FTC has a taxonomic structure as illustrated 
-on Figure~\ref{fig:02}, which arises when the reasoner classifies the knowledge base. Categories can have multiple parents and multiple children. 
-The reader can browse and access the content of the FTC online at {{https://www.ebi.ac.uk/chembl/ftc}}.
-In total there are 1280 FDA-approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets.
-Only drugs acting on at least one human protein target were selected.
-The FTC introduces 23'353 new categories describing the mode and mechanism of action of therapeutic compounds. 4289 of these 
-categories are related to biological processes and 19'064 to the molecular functions. A summary of the metrics behind the latest 
-build is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.
-Out of all FTC classes, 1432 
-categories (\textgreater 6\%) directly contain at least one approved drug. This number increases up to 2532 (\textgreater 11\%) 
-when direct and 
-indirect drugs are considered. FTC categories not containing drugs (e.g, FTC\_A0001771 - \emph{Anti-immunological synapse 
-formation agent}) represent modes and mechanisms of action for which no approved compounds exist yet or that have not been identified 
-as such in the FTC.
-
-\begin{figure}[!tpb]%figure2
-\centerline{\includegraphics{fig2.png}}
-\caption{Parent categories to the FTC class \emph{Pro-fibrinolysis agent} (FTC\_P0042730). The classification is a direct 
-acyclic graph where categories are describing increasingly specific concepts. Arrows entail subclass relationships 
-between the terms (\emph{is a} relation).}\label{fig:02}
-\end{figure}
-
-\subsection{Evaluation}
-The content of the FTC has been evaluated against the information present in the ATC. The motivation behind 
-this step is to have an estimation of how the automatically-built classification (FTC) compares to a manually curated 
-one (ATC). In the context of this work, we have considered the ATC as a gold standard against which the FTC 
-has been assessed. In practice, these two resources have different goals; therefore the evaluation should be 
-interpreted accordingly (cf section \ref{interpreting}). The full methodology behind the evaluation is 
-described in the section 6 of the supplementary material (\emph{Evaluation methodology}). 
-Briefly, 68 categories present in the FTC have an equivalent meaning to a combination of categories present in the ATC. 
-This semantic equivalence was manually asserted 
-and called \emph{evaluation point} (see supplementary material section 6.1 for details). 
-The set of drugs present in both classes for each evaluation points were then compared and some metrics derived. 
-For example, the FTC category \emph{Anti-hydrogen:potassium-exchanging ATPase activity agent} (FTC\_A0008900) has 
-been manually asserted as equivalent to the ATC category \emph{Proton pump inhibitors} (A02BC). A summary of this 
-evaluation point is furthermore available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/FTC\_A0008900}}.
-Out of the 1280 DrugBank compounds present in the FTC, 1134 are also present in the ATC, therefore only those were 
-considered. The \emph{evaluation points} cover a total of 471 DrugBank compounds or around 41\% of common drugs to both classifications.
-Out of these, 275 compounds are true positives, 
-meaning that given an evaluation point, they are present in both FTC and ATC categories. The \emph{proton pump inhibitor} 
-evaluation point is such an example where all the drugs (\emph{omeprazole, esomeprazole, pantoprazole, lansoprazole, rabeprazole}) are 
-present both in the FTC category and in the corresponding ATC class(es). The total number of compounds present in an ATC class but 
-not in the corresponding FTC category (false negatives) is 35. Finally, 280 compounds are present in a FTC class but not in the 
-corresponding ATC class(es); such agents are the false positives. From these values, it is possible to derive a 
-recall of 89\%; this percentage indicates that the automatic build of the FTC covers a good portion of the content 
-already present in the ATC. The precision of 50\% shows that the FTC contains for a given MoA many more drugs than the 
-equivalent ATC categories. This result was expected and comes the original idea behind the FTC: Representing in a systematic fashion
-the implicit and explicit MoAs of drugs, in particular the ones not already indexed by current classification scheme.
-
-\subsection{Exploration}\label{exploration}
-The FTC was designed to assist drug repurposing analyses by explicitly representing the polypharmacology of 
-approved drugs. In this section we exemplify how the resource can be used to perform different types of analysis.
-
-\subsubsection{Polypharmacology spectrum}
-The more information on a drug’s molecular targets and their physiological roles, the more opportunities 
-exist to re-orient a drug into doing something new. The therapeutic agents described in the FTC can have several MoAs, 
-representing the intrinsic polypharmacology of the approved compounds. Figure \ref{fig:03} illustrates the polypharmacology 
-spectrum by showing the distribution of number of MoAs per compound. When only direct categories are considered, compounds belong on average
-to 13.5 MoA categories. This number increases to 61.2 when parent categories are taken into consideration (super classes).
-Not all the MoAs are disease relevant, 
-some FTC categories are particularly abstract (e.g, \emph{Anti-biological process agent}) yet they represent discrete 
-categories to which the drug belongs with an explicit and clear meaning. These discrete MoAs are a good starting 
-point to understand what a compound can do when administered in a human system, as it will be shown with Alzheimer's disease.
-The human readable definitions 
-associated to each FTC category furthermore help the user to choose the adequate MoA for the desired outcome.
-Compound's polypharmacology is 
-well represented in the FTC, as shown by the numerous MoAs each approved drug can exhibit.
-
-\begin{figure*}[!tpb]%figure3
-\centerline{\includegraphics{fig3.png}}
-\caption{Distribution of the direct (blue) and indirect (red) number of MoAs per drug.
-Means are indicated with a solid line. On average each compound has 13.5 MoAs when only direct classes are considered. The number rises
-to 61.2 when indirect MoAs are included. Indirect MoAs are the ancestor classes in the taxonomy as shown in Figure \ref{fig:02}.
-The distribution range is wider when indirect MoAs are considered (range=299; min=7; max=306) versus direct MoAs only (range=79; min=3; max=82).
-These results emphasises the fact that some drugs
-are well characterised in databases and could be used for a variety of specific biological tasks. Finally some compounds have been assigned
-to a small number of FTC categories; in such cases little is known or reported about their pharmacology and repurposing opportunities
-might be limited.}\label{fig:03}
+\begin{figure*}[!tpb]%figureS3
+\centerline{\includegraphics{figS4.png}}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but with a hierarchical clustering applied on the top of it, based on the manhattan distance.
+The dendrogram reflects the original structure of the FTC: Functionaly similar compounds being grouped together.
+This figure can be used for further analysis over functional clusters of drugs.}\label{fig:S04}
 \end{figure*}
 
-\subsubsection{Drugs with similar MoAs have similar indications}
-The list of MoAs attributed to a drug can be exploited as a descriptor for the therapeutic agent:
-The tree structure of the FTC can be used to derive some similarity metrics over the MoAs. The underlying 
-heuristic is to assume that the closer two entities are in the taxonomy, the more similar they are. 
-We used a straightforward approach derived from the Jaccard index (see supplementary section 7) in order to compare 
-approved drugs based on the similarity of their MoAs. For instance, the similarity between two compounds present in 
-the same FTC category is 1 (maximum). The similarity between an \emph{anti-blood coagulant} and \emph{pro-blood coagulant} is 0.29, 
-reflecting the fact that such compounds are not similar in regards to the outcome of their biological effect. As the MoA is 
-intuitively expected to be the central concept leading to the indication of the drug, we expected that on average, drugs with 
-similar MoAs would be indicated towards similar therapeutic areas. The heat map presented in Figure \ref{fig:04} shows a pairwise 
-comparison of all the drugs of the FTC based on their relative MoA similarity. The compounds are further grouped by 
-therapeutic indications as defined by the ATC.
-The heatmap reveals some square patches around the central diagonal;
-the overall similarity appears higher when compounds from the same ATC group are considered. 
-A significance analysis (details covered in supplementary section 8)
-revealed that the average MoA similarity of compounds belonging to the same ATC category is 
-significantly higher than when compounds belonging to different categories are compared. Indeed, for 
-each category, the p-value was inferior to 0.05 based on 20'000 random permutations over the similarity values. 
-This result supports the idea that drugs with similar MoAs have similar indications. Note that the mean of the 
-similarity values was considered for the statistical analysis; some outliers are also present in the map, which 
-can be interpreted as repositioning hypotheses. These outliers have indeed similar MoAs, yet they belong to totally 
-different therapeutic areas and are used for different purposes according to the ATC. We are currently further analysing 
-such cases in a systematic fashion. Hypotheses have to be manually examined and interpreted, as ATC categories are
-only covering some of the legal usage of the drugs. We expect to find off-label indications in 
-the predictions for instance, as well as false positives.
-
-Supplementary Figure 3 present similar association behaviour when two levels of the ATC are 
-considered (no statistical significance performed). Supplementary Figure 4 re-uses the same data as Figure \ref{fig:04} (one ATC level) 
-but with a clustering function apply to it (hierarchical clustering - manhattan distance) in order to reveal 
-functional clusters of drugs. Finally, Supplementary Figure 5 shows the distribution when compounds are sorted based on 
-their identifiers; no patterns are identifiable in this case. Taken together, these results emphasize that the MoAs as 
-defined in the FTC are indeed on average associated with the therapeutic indication of a drug. This result supports the validity of the
-resource and its potential to computationally address indication discovery.
-
-\begin{figure*}[!tpb]%figure4
-\centerline{\includegraphics{fig4.png}}
-\caption{Pairwise comparison of MoAs similarities. Therapeutic indications are represented by ATC categories which are the colors on the side.
-For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
-which appears as \emph{B} (dark orange) on the plot.
-Only the first ATC level 
-is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
-The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black).
-Some compounds belong to multiple ATC categories (\emph{Mulitple}) and some others do not have
-an ATC code (\emph{NoCategory}).
-The average similarity of drugs present in the same therapeutic category 
-is significantly higher on average when separately compared to all other indications.}\label{fig:04}
+\begin{figure*}[!tpb]%figureS5
+\centerline{\includegraphics{figS5.png}}
+\caption{Pairwise comparison of MoA similarities. This graph presents the same 
+data as Figure 4 but the compounds have been sorted horizontally and vertically only based on their identifiers, which
+we considered as equivalent to a pseudo-random sorting. In this case
+no visual pattern is observable, the similarity values appear homogeneously distributed (no statistical analysis performed).}\label{fig:S05}
 \end{figure*}
-
-\subsection{Drug repurposing hypotheses for Alzheimer's disease}
-In this section we provide examples how actual drug repurposing hypotheses can be derived from the FTC. 
-The approach presented here makes use of the FTC categories as analogs to compartments of a toolbox helping to 
-find drugs to treat Alzheimer's disease. Five FTC categories containing drugs are directly related to the biological 
-processes of the neurodegenerative condition: \emph{Anti-amyloid precursor protein biosynthetic process agent} (FTC\_A0042983), 
-\emph{Anti-tau protein kinase activity agent} (FTC\_A0050321), \emph{Anti-tau protein binding agent} (FTC\_A0048156), 
-\emph{Anti-beta amyloid 
-binding agent} (FTC\_A0001540) and finally \emph{Pro-beta amyloid binding agent} (FTC\_P0001540). We have then considered the drugs present 
-inside each of these classes as potential candidates. Figure \ref{fig:05} shows these drugs, which have been further manually grouped 
-based on the overall similarity of their actions (numbers on Figure \ref{fig:05}).
-
-\begin{figure}[!tpb]%figure5
-\centerline{\includegraphics{fig5.png}}
-\caption{FTC categories describing some of the modes and mechanisms of 
-action that could impact Alzheimer's disease (letters on figure). The categories have been manually picked on the basis that they could directly affect
-the dementia. Drugs classified in these FTC categories further 
-manually grouped based on their MoAs similarities (numbers on figure).}\label{fig:05}
-\end{figure}
-
-The subgroups (1), (2) and (3) are inhibitors of the cholinergic system and some of them, such as \emph{galantamine} (DB00674), are already 
-investigated to treat Alzheimer's disease and other related dementias. This class of agent is in line with the 
-cholinergic hypothesis \citep{Francis1999}, stating that Alzheimer's disease could be caused by dysfunctions in the processing of the
-acetylcholine. The subgroup (4) is exclusively composed of barbiturates (central nervous system depressants). 
-The presence of this pharmacological class of compounds as an Alzheimer's disease treatment is more surprising, as very little 
-literature reports on it. Further investigations reveal that the neuronal acetylcholine receptor subunit alpha-7, a 
-common off-target of barbiturates, binds beta-amyloids with high affinity \citep{Wang2000}. As beta-amyloids are themselves strongly involved 
-in the pathology, barbiturates could affect the state of the condition, similarly to cholinergic inhibitors \citep{Wang2000}.
-The group (B) contains four compounds; \emph{nicotine} and \emph{varenicline} have been further grouped together because of 
-similar pharmacology (group 5).
-\emph{Nicotine} has been shown to improve some of the symptoms of Alzheimer's disease \citep{Jones1992}, it is therefore expected to 
-find this molecule in the predictions. \emph{Varenicline} possess a pharmacology related to that of \emph{nicotine}, which would explain 
-the presence of the drug in this category. The two remaining drugs of group F are respectively \emph{pralidoxime} and \emph{dipivefrin}; 
-little information is available regarding their potential action against the condition, yet these compounds seem linked to the 
-cholinergic hypothesis and could be considered for experimental testing.
-The groups (C) and (D) contain respectively one molecule each. These compounds have been classified as agents perturbing some of 
-the physiological function of the Tau protein, key actor in Alzheimer's 
-disease \citep{Grundke-Iqbal1986}. \emph{Vorinostat} (group C) is currently 
-indicated for the treatment of cutaneous manifestations in patients with T-cell lymphoma, yet a study has 
-shown in vivo (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
-deficit \citep{Kilgore2010}. 
-The presence of the \emph{lithium} (group D) was 
-confirmed by a recent study demonstrating a long-term 
-protective effect for the ion in regards to Alzheimer's disease \citep{Young2011}. The last group (E) contains 
-\emph{ezetimibe} and \emph{hesperetin}. These two compounds are primarily used as cholesterol lowering agents (statins). As cholesterol 
-metabolism in the brain appears to be related to dementia, statins are believed to prevent or improve the symptoms of the 
-patients. Although early studies \citep{Wolozin2004} have failed to clearly show a beneficial effect, the investigation is still open.
-From the examples briefly presented above, reported and confirmed by the literature, the FTC appears to be suitable to 
-identify real repurposing hypotheses tailored to a disease. Correctly 
-identifying MoAs of interest helps to retrieve the compounds which might impact 
-the treatment of a condition.
-
-\section{Discussion}
-
-The FTC is a novel classification for approved drugs, which can be used as a starting point to 
-generate drug repurposing hypotheses. We have guided the reader through some of the ways the resource 
-can be explored. No experimental validations have been presented in this manuscript, yet references to 
-supporting literature was provided in order to legitimate the claims. The main goal of this document is to 
-introduce the resource to the community; deeper analyses are being performed over the classification at the time of writing.
-
-\subsection{Biological assumptions}
-An asset of the FTC is its ability to handle efficiently categorical data: Classes and relationships 
-are accurately defined, in order to classify compounds based on the semantics of their relations. The properties 
-linking drugs to their respective protein targets (\emph{positive} and \emph{negative perturbations}) are however simplistic. At the 
-time being, no consideration is given regarding the binding strength between the drug and the proteins, yet it is a key 
-factor to derive potent and specific activities in the human body. This is also the case for other types of numerical data, 
-such as the dosage; the FTC can predict a role for a drug, yet it cannot provide any information about the concentration or 
-the administration route necessary to obtain the potential effects. The current relations between targets and their involvement 
-in biological processes are also not a fully accurate representation of the biological phenomenon. In a cell, specific domains 
-of the protein could mediate different functions. Only one of such activity types can sometimes be inhibited 
-by a drug \citep{kruger}, yet we are assuming in the FTC that as long as a drug affects a protein, 
-it can therefore alter all its known functions.
-These limitations come from the semantics behind the axioms structuring the classification, themselves based on the 
-information available from the databases. Despite entailing not entirely accurately the 
-biochemical reality, the axioms help to generate a larger number of hypotheses, the primary goal of the FTC. The dosage 
-issue is partially addressed by the \emph{regulator pattern} (see section 3.1 of supplementary material): It should be 
-easier to experimentally adjust the concentration of the compounds classified as \emph{pro} or \emph{anti} biological process agents 
-in order to modulate a physiological effect.
-The predictions generated by the FTC depend on the resolution of the curated information released by 
-the original data providers. Erroneous or missing information will lead to misclassification by the reasoner. 
-Some expected outcomes are also missing from the predictions; \emph{sildenafil} for instance was expected to be classified 
-as \emph{pro-penile erection agent} (FTC\_A0043084), yet the lack of appropriate GO annotation prevents it. After discussion 
-with the GOA curation team, a manual annotation can only be asserted based on published experimental results. No document 
-was found to support the involvement of the cGMP-specific 3',5'-cyclic phosphodiesterase (\emph{sildenafil}'s main target) 
-in the \emph{negative regulation of penile erection} (GO:0060407), therefore no annotation can be made. Further work could be done 
-in this direction, by trying to automatically infer more annotations or by using the electronically generated ones, in order 
-to generate broader yet potentially less plausible repurposing hypotheses.
-
-\subsection{Interpreting the evaluation}\label{interpreting}
-Out of the evaluation, the high recall value (89\%) supports the idea behind the automated build of the FTC: The data from different 
-repositories funded and curated in parallel, can be integrated to automatically create a new resource. This new classification (FTC) 
-contains most of the known information present in an external gold standard (ATC) and relies on description logics to 
-leverage the native information. In the context of this work we have compared the content of 
-the FTC against the ATC, knowing that these two 
-taxonomies have diverging goals. During the evaluation, equivalences have been manually asserted between classes, 
-which are assumed to have fairly similar meaning and containing similar sets of compounds.
-These manual assertions 
-are however a weakness, as they are themselves not evaluated (free parameter). The presence or absence of a link was 
-determined only by one curator and any mistake can influence consequently the recall and precision values.
-The precision of 50\% tells that the FTC tends to over-assigns compounds to MoA categories. The low precision value is acceptable 
-in our case, as one of the underlying motivation of the FTC is to broadly represent polypharmacology, specially the one not present in
-gold standards such as the ATC, referencing only legal usage.
-In this regards, 
-the evaluation should be considered more as a safety control rather than a formal assessment of a predictive method.
-The false positives derived from the evaluation can also be considered as drug repurposing hypotheses: These drugs 
-can indeed be interpreted as suitable for the ATC category, yet not indexed as such. However, these predictions 
-should be interpreted with caution, as it is currently impossible to distinguish a false positive from a 
-reprofiling opportunity. These considerations do not interfere with the repurposing predictions generated 
-based on semantic similarities or discrete categories as presented in the section \ref{exploration}. Finally, 
-note that the ATC/FTC equivalences are open and editable online, any modification will be automatically 
-incorporated in the next release of the resource. It is also possible to evaluate the FTC against a different 
-taxonomy, like the Medical Subject Headings for example, which can be subject to future work.
-
-\section{Conclusion}
-This manuscript introduces the FTC, a free, open and extensible resource, which can assist drug repurposing 
-initiatives or enhance computational studies making use of the concept of \emph{mode of action}.
-The resource annotates the biological functions of drugs, in a similar way the GO serves to annotate gene products.
-The construction of 
-the classification relies on description logics as a means to define the mode and mechanism of action of approved 
-drugs. We shown the validity of the approach by comparing the content of the FTC to another well implemented taxonomy, 
-the ATC. We further illustrate the tight relationship between the MoA and the indication of a drug and shown how the 
-resource can serve as a starting point to formulate hypotheses for Alzheimer's disease. The work leverages the 
-semantics of distinct databases, working in parallel on different thematics. The platform will be further used 
-to generate predictions in a systematic fashion, which can then be experimentally tested in the laboratory for validation.
-
-\section{Availability and Implementation}\label{impl}
-The code behind the creation of the resource is entirely open and available 
-at {{https://github.com/loopasam/ftc}}. The web application and the FTC are built using the Brain library \citep{Croset2013} and can be 
-find at {{https://www.ebi.ac.uk/chembl/ftc}}. The documentation can be accessed at {{https://github.com/loopasam/ftc/wiki}}. The full
-description of the methodology used to generate the classification is available as supplementary material on Bioinformatics website.
-
-\section*{Acknowledgement}
-We would like to thank Rita Santos, Sarah Carl and Felix Kruger for the fruitful input on this manuscript.
-
-\paragraph{Funding\textcolon} EMBL member states. S.C. is a member of Darwin College, University of Cambridge.
 

### 2013-10-10 reply-to-editors.txt
letter to editors done

@@
-Reviewer: 1
-Comments to the Author
-This manuscript describes a classification system for automatic creation MoA annotations and generation of drug repurposing hypotheses.
-The authors have developed an online database and released its source code to disseminate this data.
-The article is clearly laid out and well explained.
-The approach is novel and should be useful both to the bioinformatic community and a wider scientific audience, but several issues certainly require further clarification.
+Dear Editors,
 
-Major revisions
-1) The authors describes FTC is built by integrating information coming from various sources and using ELK algorithm to classify the knowledge.
-However, I cannot find any clear and detail introduction in manuscript or supplement.
-This should be explained in detail, being one of the most distinctive features of the paper.
+We read with care the comments provided by the reviewers.
+We hereby present a modified version of the manuscript, based on their valuable input.
+Reviewer 1 asked for a more detailed presentation of the reasoning algorithm and methodology.
+In this regard, we have provided a supplementary reference to an introductory text.
+It is challenging to add more information to the main document, as reasoning services and tools have been developed by other research groups and are simply used in the context of our work.
+We rely on the W3C OWL international standard to interact with reasoning engines, for which extensive documentation is available.
+Moreover, we specially wrote an introductory text to description logics, available on the wiki for biomedical users non familiar with the domain.
+Ideally description logics would be presented in greater details in the main manuscript, unfortunately and because of space concerns we decided to focus on the presentation of the biomedical novelty brought by our project.
+Understanding in details the methodology requires some knowledge about description logics (as noticed by Reviewer 2 too), yet we believe that Figure 1 provides a clear simplified overview.
+Reviewer 1 did not manage to find the Gene Ontology (GO) annotation "positive regulation of blood coagulation" for the protein Prothrombin (P00734).
+The annotation exists and is present on the GO annotations website: http://www.ebi.ac.uk/QuickGO/GProtein?ac=P00734.
+Reviewer 1 understood correctly the methodology, yet probably missed the link "Complete GO annotation..." on the Uniprot website, leading to the complete list where the mentioned annotation can be found.
+Reviewer 1 also asks for "how much of the different results between ATC and FTC".
+The relation between the two resources is described in section 2.1 "Evaluation" and furthermore on the website of the application (https://www.ebi.ac.uk/chembl/ftc/evaluation/), where we compare equivalent classes and derive metrics (precision and recall) summarising the overlap between these resources.
+It is not clear to us what are the specific requirements of reviewer 1 regarding this revision.
+Reviewer 2 asked us to illustrate the potential of the resource with a known repurposing example.
+We chose the thalidomide to fulfil this legitimate requirement, as it is possibly one of the best known repositioning case.
+We have added a paragraph to the manuscript (end of section 2.2.1) explaining how the teratogenic effect of the molecule is described in the FTC and how the new indications of the drug are captured (anti-cancer and immunomodulatory activities, nowadays routinely prescribed to patients).
+We briefly discussed the relationship between our resource and the work done in Bioinformatics 28.16 (2012): 2169-2175 and the Open PHACTS platform.
+Moreover, the methodology section further references an omitted similar approach presented in Nucleic Acids Res.
+2011 Oct;39(18):e119.
+Reviewer 2 pointed the confusing number of drugs and their origin.
+This section was clarified in the manuscript (section 2); we used the FDA-approved drugs as described in DrugBank and interacting with at least one human protein target.
+Reviewer 2 asked whether the outliers points present in Figure 4 could have different interpretation; the text was extended accordingly, and the meaning of the data points summarily discussed.
+Reviewer 2 expressed concerns regarding the user interaction with the resource, as a large number of new classes are present.
+We argue that the web application greatly improves the accessibility of the FTC's content.
+Each FTC category is described in details on the web application, and is further linked to related terms for the user to browse.
+Moreover, a text search functionality is available (https://www.ebi.ac.uk/chembl/ftc/search/), in order to easily find categories or drugs of interest.
+More search functionality might be added in the future, depending on the adoption of the classification.
+Minor corrections have been incorporated, as the exception of supplementary tables listing all the evaluation points.
+We believe that the website summarises better this aspect of the work.
+The text of the manuscript was also extensively edited and proof-read.
+Finally, Reviewer 3 did not manifest any major or minor objections.
 
-2) Result section 2, 2.1 is mostly methods benchmarking.
-I suggest those section should be renamed as “methodology” section.
-In addition, the detail classification description in supplementary (i.e.
-section 3, 3.1 and 3.2) should be wrote at manuscript.
-It would make it clearer how the methods are developed and tested.
+We believe to have addressed or answered all the major comments raised by the referees with this new improved version of the manuscript. We are looking forward to hear your decision.
 
-3) In Figure 1, part C (Incorporation of databases), authors show ‘Prothrombin’ (P00734) is involved in ‘positive regulation of blood coagulation’ (GO_0030194).
-I tried to check this information from UniProt, but I cannot find any description there (http://www.uniprot.org/uniprot/P00734).
-How is this relationship module structured?
-This should be explained in detail.
+Best regards,
 
-4) How much of the different results between ATC and FTC? It should be listed and accessed via web.
-
-Minor points:
-1) Authors used ATC corpus to evaluate proposed FTC algorithm. The detailed evaluation results were shown on project’s web. I suggest the results should be listed at manuscript, not only at web.
-2) 4 Gb of memory --> 4 GB of memory
-
-Reviewer: 2
-Comments to the Author
-The authors present a taxonomy, the Functional Therapeutic Chemical Classification System, and a knowledge base to assist drug repurposing analyses.
-The mode of action of a drug, represented by Gene Ontology derived terms, forms the basis for the drug similarity.
-The authors evaluate their taxonomy by comparing it against the Anatomical Therapeutic Chemical Classification System, and show the usefulness of their system by applying it in a use case to find new drugs for Alzheimers disease.
-
-The knowledge base and taxonomy is very well-documented, but requires deep knowledge about both the system and description logic in general in order to use.
-Also, presenting a single example use case with only literature evidence to back up the results is not very convincing for a research paper.
-
-Major
-Please show that the system can reproduce one known case of drug repurposing.
-
-The authors should describe differences with their system to related efforts such as work by Hoehndorf, Dumontier, and Gkoutos.
-Bioinformatics 28.16 (2012): 2169-2175, and the Open PHACTS discovery platform http://www.openphacts.org/.
-A reference to a recent review paper such as the one by Andronis et al in Brief Bioinform.
-2011 Jul;12(4):357-68 would be helpful as well.
-
-In addition, the following specific revisions are needed:
-p.
-2, line 39: “In total there are 1280 approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets.” I looked up the current Drugbank figures: the database contains 6811 drug entries including 1528 FDA-approved small molecule drugs, 150 FDA-approved biotech (protein/peptide) drugs, 87 nutraceuticals and 5080 experimental drugs.
-It is therefore not clear to me where the number 1280 comes from.
-Later the authors talk about “approved” drugs.
-Do they refer to FDA-approved?
-
-p.
-4 line 47: “Note that the mean of the similarity values was considered for the statistical analysis; some outliers are also present in the map, which can be interpreted as repurposing hypotheses.” Is this the only interpretation?
-
-p.
-4 line 51: “Five FTC categories containing drugs are directly related to the biological processes of the neurodegenerative condition:..” It seems like a user needs to know a lot of details about the system with > 23000 categories in order to use it.
-How can the user be helped here?
-
-Minor
-
-p. 2 line 26: “Drugs are further assigned into MoA categories based of the OWL constructs and axioms defined in the FTC (see Availability and Implementation).” Provide an example here.
-
-p. 5 line 52: “The presence of lithium (group D) was unexpected and suspected of being a false positive at first glance.” Please explain why this result was unexpected.
-
-Supplemental material:
-“The web application built on the top of the FTC can be find …” should be “FTC can be found”
-
-“The FTC implementation relies mostly on Brain [Croset et al., 2013] and the web application builds on the top of the Play! framework”. Please provide reference for the Play! Framework.
+Samuel Croset

### 2013-10-14 document.tex
modification post-review done
modification DRS

--- 2013-10-10 document.tex
@@
 \section{Motivation:}
-Drug repurposing is the discovery of new indications for compounds already approved and used in a clinical setting. 
-Recently, some computational approaches have been developed to unveil new opportunities in a systematic fashion, by taking 
+Drug repurposing is the discovery of new indications for compounds that have already been approved and used in a clinical setting. 
+Recently, some computational approaches have been suggested to unveil new opportunities in a systematic fashion, by taking 
 into consideration gene expression signatures or chemical features for instance. 
@@
 \section{Results:}
-In order to computationally generate repurposing hypotheses, we used the Web Ontology Language (OWL) to 
-mathematically define the meaning of over 20'000 terms formally describing various modes and mechanisms of action. Based on an 
+In order to computationally generate repositioning hypotheses, we used the Web Ontology Language (OWL) to 
+formally define the semantics of over 20'000 terms with axioms to correctly denote various modes of action (MoA). Based on an 
 integration of public data, we have automatically assigned over a thousand of approved drugs into 
@@
 \centerline{\includegraphics{fig1.png}}
-\caption{Summary of the methodology used to construct the resource. (A) The FTC categories are first created from the existing GO terms
-describing molecular functions and biological processes. (B) Then, each category has a logical equivalent definition assign to it.
-(C) The content of various databases in incorporated and linked using the FTC specific properties. (D) Finally a reasoner runs over the knowledge
-base and assigns drugs to MoA categories based on whether or not a definition can be satisfied. For example here, the drug
+\caption{The diagram gives an overview of the integrated resources. (A) The name of FTC categories representing MoAs are derived from the GO terms representing the molecular 
+functions and biological processes. A 
+prefix \emph{anti} or \emph{pro} is appended to the original 
+GO term, as well as the word \emph{agent} as suffix. (B) Each of the new FTC class has a logical
+equivalent definition assigned to it (axiom), representing the necessary and sufficient conditions for a drug to be classified
+in the corresponding MoA class.
+(C) The content of various databases is incorporated and linked using the FTC specific logical properties. (D) Finally a reasoner classifies the knowledge
+base and assigns drugs to MoA classes based on whether or not a definition can be satisfied. For example, the drug
 \emph{ximelagatran} will be assigned as member of the category \emph{Anti-blood coagulation agent} because of the logical
@@
 The taxonomic structure of the
-FTC appears also in the reasoning step, from the entailements of the equivalent defintions.}\label{fig:01}
+FTC appears also in the reasoning step, from the entailment of the equivalent definitions.}\label{fig:01}
 \end{figure*}
@@
 Drug repurposing is the use of known active compounds for new therapeutic indications \citep{Sanseau01072011}. 
-When administered in a living organism, a compound can indeed play various roles and affect different biological processes; 
-accurately identifying these different functions helps to predict the potential side-effects a drug could have and can also lead to 
+When administered in a living organism, a compound can indeed play 
+various roles and affect different biological processes (called \emph{mode of action}, MoA); 
+accurately identifying these different functions 
+helps to predict the potential side-effects a drug could have and can also lead to 
 interesting repurposing opportunities \citep{Medina-Franco2013}. For instance, \emph{sildenafil} was initially developed 
-against angina and has been repurposed towards erectile dysfunction during the clinical trials \citep{Ashburn2004}. Approved compounds are 
+to relieve angina pectoris symptoms and has been repurposed towards erectile dysfunction during the clinical trials \citep{Ashburn2004} when a new function of the target enzyme was discovered.
+Approved compounds are
 attractive because they have been extensively studied and have by definition already successfully passed clinical trials, where 
@@
 have in common to look for similarities within existing drugs and forward similar compounds as repurposing hypotheses.
-A feature of particular interest to describe drugs is the Mode of Action (MoA). According to Wikipedia, 
+A feature of particular interest to describe drugs is the MoA. According to Wikipedia, 
 the MoA describes \emph{a functional or anatomical change, 
@@
 roles of a certain type of drugs. The 
-MoA abstracts over the relations between molecular functions, protein targets and drug activities; it is the central concept linking a 
-chemical structure to a set of biological effects. Intuitively, the indication of a drug logically depends on its MoA.
+MoA abstracts over the relations between molecular functions, protein targets and drug activities; it is the central concept linking a chemical structure to a set of biological activities.
+Intuitively, the indication of a drug logically depends on its MoA.
 Despite its widespread use in drug discovery, the MoA has not been used yet as a descriptor for repurposing analyses. One reason for this 
@@
 Nonetheless, the meaning of a concept can be formalised with controlled vocabularies and ontologies \citep{Gruber1995}; 
-originating from description logics, such frameworks help to define the mathematical meaning of symbols and strings of characters. 
-In an ontology or knowledge base, concepts are organised and linked following the logical type of relation they have among them. 
-In the Gene Ontology (GO) for example \citep{Ashburner2000}, biological processes and molecular functions terms are manually curated  
-and their meaning specified by the relation types linking two GO terms.
+originating from description logics, such frameworks help to formalise the semantics of symbols and strings of characters with explicit axioms. 
+In an ontology or knowledge base, 
+\emph{concepts} (interchangeable with \emph{category}, \emph{term} and \emph{class} in this manuscript) are 
+organised and linked following the logical type of relation they have among them. 
+In the Gene Ontology (GO) for example \citep{Ashburner2000}, biological processes and molecular functions terms are manually curated  and their meaning specified by the relation types linking two GO terms.
 MoA definitions are present in other classifications such as the Medical Subject Headings \citep{Nelson2004} or the Chemical Entities of 
@@
 drug as well as some predicted ones to be tested in experiments. The MoA categories should derive and scale over primary molecular 
-evidences exposed in biomedical databases in an automated way.
+evidences exposed in biomedical databases, in an automated way.
 To address the lack of systematic MoA annotations, we have implemented the Functional Therapeutic Chemical 
@@
 in the resource and further populated with approved drugs using the Web Ontology Language (OWL) in combination with a reasoner.
-Most of the drugs are present in multiple FTC categories, reflecting the various roles a compound can play inside a biological system which 
-can serve as starting point for repurposing. 
+The population step takes in account the type of pharmacological action, the molecular targets of the drug and their involvement
+into biological processes. Drugs can exhibit several MoAs, and the same MoA can be reached through different mechanisms.
+Most of the drugs are present in multiple FTC categories, reflecting the various roles a compound can play inside a biological system which can serve as starting point for repurposing. 
 The resource was evaluated against the ATC, traditional classification scheme introduced before. We present as well some preliminary 
@@
 respective GO annotations \citep{Dimmer2012}. 
-Drugs are further assigned into MoA categories based of the OWL constructs and axioms defined in the 
+Drugs are further assigned into MoA categories according to the OWL constructs and axioms defined in the 
 FTC. A reasoner, a program capable of understanding such axioms, performed this task (see \emph{Availability and Implementation} for details). 
 The process to build the FTC is summarised in Figure~\ref{fig:01} alongside an example.
-It takes around 5 seconds on a modern 
-laptop (4 cores processor with 4 GB of memory) to classify the knowledge base using the ELK reasoner \citep{Kazakov2011}. 
-Other OWL reasoners have been tried (e.g, Hermit, Pellet, etc.) but none of them were suitable for this purpose; 
-the classification time was always superior to a few minutes (data not shown). The FTC has a taxonomic structure as illustrated 
-on Figure~\ref{fig:02}, which arises when the reasoner classifies the knowledge base. Categories can have multiple parents and multiple children. 
-The reader can browse and access the content of the FTC online at {{https://www.ebi.ac.uk/chembl/ftc}}.
-In total there are 1280 FDA-approved DrugBank compounds (chemical and biotherapeutics) perturbing 1264 human protein targets.
-Only drugs acting on at least one human protein target were selected.
-The FTC introduces 23'353 new categories describing the mode and mechanism of action of therapeutic compounds. 4289 of these 
-categories are related to biological processes and 19'064 to the molecular functions. A summary of the metrics behind the latest 
+The core step is the generation of axiomatic representations of MoAs by decomposing GO types into positive and negative regulations of biomolecular functions and processes. 
+The help of reasoning techniques we can further derive and assign MoA across the knowledge base to given drugs.
+It requires a few seconds (four processing cores, 4 GB RAM) to classify the knowledge base (ELK reasoner, \citep{Kazakov2011}).
+Other OWL reasoners (e.g, Hermit, Pellet, etc.) disqualified mainly due to long processing time (data not shown). 
+The FTC forms a taxonomic structure as illustrated 
+on Figure~\ref{fig:02}, which arises when the reasoner classifies the knowledge base.
+In general, categories may have multiple parents and multiple children (see {{https://www.ebi.ac.uk/chembl/ftc}} for interactive use).
+In total there are 1'280 FDA-approved DrugBank compounds (chemical and biotherapeutics) associated with 1'264 human protein targets, where each drug is acting on at least one human protein target.
+The FTC introduces 23'353 new categories new categories describing the mode and mechanism of action of therapeutic compounds. 
+4'289 of these categories belong to the biological processes in GO and 19'064 to the molecular functions. A summary of the metrics behind the latest 
 build is available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/}}.
-Out of all FTC classes, 1432 
-categories (\textgreater 6\%) directly contain at least one approved drug. This number increases up to 2532 (\textgreater 11\%) 
+Out of all FTC categories, 1'432 
+categories (\textgreater 6\%) directly contain at least one approved drug. This number increases up to 2'532 (\textgreater 11\%) 
 when direct and 
 indirect drugs are considered. FTC categories not containing drugs (e.g, FTC\_A0001771 - \emph{Anti-immunological synapse 
-formation agent}) represent modes and mechanisms of action for which no approved compounds exist yet or that have not been identified 
+formation agent}) represent MoAs for which no approved compounds has qualified yet or that have not been identified  
 as such in the FTC.
@@
 \subsection{Evaluation}
-The content of the FTC has been evaluated against the information present in the ATC. The motivation behind 
-this step is to have an estimation of how the automatically-built classification (FTC) compares to a manually curated 
-one (ATC). In the context of this work, we have considered the ATC as a gold standard against which the FTC 
-has been assessed. In practice, these two resources have different goals; therefore the evaluation should be 
-interpreted accordingly (cf section \ref{interpreting}). The full methodology behind the evaluation is 
-described in the section 6 of the supplementary material (\emph{Evaluation methodology}). 
-Briefly, 68 categories present in the FTC have an equivalent meaning to a combination of categories present in the ATC. 
-This semantic equivalence was manually asserted 
-and called \emph{evaluation point} (see supplementary material section 6.1 for details). 
-The set of drugs present in both classes for each evaluation points were then compared and some metrics derived. 
+The content of the FTC has been evaluated against the drug categorisation of the ATC, which has been produced by manual curation and serves as a gold standard. 
+\emph{A priori}, both resources serve different purposes and as a consequence, the evaluation has to take this into consideration (cf section \ref{interpreting}). 
+The full methodology behind the evaluation is described in the section 6 of the supplementary material (\emph{Evaluation methodology}).
+Briefly, for 68 categories from the FTC we can manually identify a set of semantically equivalent categories in the ATC.
+We call these equivalent categories the \emph{evaluation points} (see supplementary material section 6.1 for details).
+All drugs from each evaluation point were then assessed to determine the quality of the FTC against our gold standard, i.e.\ the ATC.
 For example, the FTC category \emph{Anti-hydrogen:potassium-exchanging ATPase activity agent} (FTC\_A0008900) has 
@@
 evaluation point is furthermore available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/FTC\_A0008900}}.
-Out of the 1280 DrugBank compounds present in the FTC, 1134 are also present in the ATC, therefore only those were 
+For 1'280 DrugBank compounds in the FTC, 1'134 are also present in the ATC, therefore only those were 
 considered. The \emph{evaluation points} cover a total of 471 DrugBank compounds or around 41\% of common drugs to both classifications.
-Out of these, 275 compounds are true positives, 
-meaning that given an evaluation point, they are present in both FTC and ATC categories. The \emph{proton pump inhibitor} 
-evaluation point is such an example where all the drugs (\emph{omeprazole, esomeprazole, pantoprazole, lansoprazole, rabeprazole}) are 
-present both in the FTC category and in the corresponding ATC class(es). The total number of compounds present in an ATC class but 
-not in the corresponding FTC category (false negatives) is 35. Finally, 280 compounds are present in a FTC class but not in the 
-corresponding ATC class(es); such agents are the false positives. From these values, it is possible to derive a 
+Out of these, 275 compounds are true positives, i.e.\ they match both, the FTC and ATC categories for a given evaluation point.
+The \emph{proton pump inhibitor} 
+evaluation point is such a case where all the drugs (\emph{omeprazole, esomeprazole, pantoprazole, lansoprazole, rabeprazole}) are 
+present both in the FTC category and in the corresponding ATC categori(es). The total number of compounds from an ATC category but 
+where we could not identify a corresponding FTC category is 35 (false negatives). Finally, 280 compounds are present in a FTC class but not in any 
+corresponding ATC category (false positives). Overall we derive a 
 recall of 89\%; this percentage indicates that the automatic build of the FTC covers a good portion of the content 
@@
 the implicit and explicit MoAs of drugs, in particular the ones not already indexed by current classification scheme.
-
-\subsection{Exploration}\label{exploration}
-The FTC was designed to assist drug repurposing analyses by explicitly representing the polypharmacology of 
-approved drugs. In this section we exemplify how the resource can be used to perform different types of analysis.
-
-\subsubsection{Polypharmacology spectrum}
-The more information on a drug’s molecular targets and their physiological roles, the more opportunities 
-exist to re-orient a drug into doing something new. The therapeutic agents described in the FTC can have several MoAs, 
-representing the intrinsic polypharmacology of the approved compounds. Figure \ref{fig:03} illustrates the polypharmacology 
-spectrum by showing the distribution of number of MoAs per compound. When only direct categories are considered, compounds belong on average
-to 13.5 MoA categories. This number increases to 61.2 when parent categories are taken into consideration (super classes).
-Not all the MoAs are disease relevant, 
-some FTC categories are particularly abstract (e.g, \emph{Anti-biological process agent}) yet they represent discrete 
-categories to which the drug belongs with an explicit and clear meaning. These discrete MoAs are a good starting 
-point to understand what a compound can do when administered in a human system.
-Compound's polypharmacology is 
-well represented in the FTC, as shown by the numerous MoAs each approved drug can exhibit.
-We decided to further look at a well-known repositioning example, 
-in order to see whether the FTC was suitable to identify the new
-uses of an old drug. We picked the \emph{thalidomide} for 
-this exercise ({{https://www.ebi.ac.uk/chembl/ftc/agent/DB01041}}). The molecule was 
-first indicated to treat morning sickness
-in pregnant women, but got quickly withdrawn from the market after the dramatic identification 
-of teratogenic effect in newborns, caused by the drug. The accepted molecular mechanism behind the side-effect is  
-an impairment of the angiogenic process responsible for the development of members, affecting in particular the limbs \citep{therapontos2009thalidomide}.
-The \emph{thalidomide} was accurately classified as \emph{Anti-cell migration involved in sprouting angiogenesis agent} in the FTC, capturing the known toxicity of the drug.
-The \emph{thalidomide} is currently investigated for a multitude of new usages,
-in particular for anti-cancer and immunomodulatory activities among others \citep{teo2005thalidomide}.
-These new indications are well represented in the FTC too, for 
-example by the categories \emph{Anti-vascular endothelial growth factor production agent} or 
-\emph{Anti-cell division agent} for neoplastic activities, 
-or by the classes \emph{Anti-cytokine secretion agent} and \emph{Anti-I-kappaB kinase/NF-kappaB cascade agent} for its effect
-on the immune system. These observations demonstrate that the FTC can successfully capture the 
-molecular reasons behind the repositioning of an old compound. 
-Moreover the classification can also provide valuable insight regarding
-potential toxicity too.
 
@@
 
+\subsection{Exploration}\label{exploration}
+The FTC was designed to assist drug repositioning analyses by explicitly representing the polypharmacology of 
+approved drugs. In this section we exemplify how the resource can be used to perform different types of analysis.
+
+\subsubsection{Polypharmacology spectrum}
+The more information on a drug’s molecular targets and their physiological roles, the more opportunities 
+exist to re-orient a drug into doing something new. The therapeutic agents described in the FTC can have several MoAs, 
+i.e.\ may be acting on different biomolecular functions or processes, 
+which demonstrates the intrinsic polypharmacology of the approved compounds. Figure \ref{fig:03} illustrates the polypharmacology 
+spectrum by showing the distribution of number of MoAs per compound. When only direct categories are considered, compounds belong on average
+to 13.5 MoA categories. This number increases to 61.2 when parent categories are taken into consideration (super classes).
+Not all the MoAs are relevant to a disease, 
+some FTC categories are particularly abstract (e.g, \emph{Anti-biological process agent}) yet they represent discrete 
+categories to which the drug belongs with an explicit and clear meaning. These discrete MoAs are a good starting 
+point to understand what a compound can do when administered in a human system.
+Compound's polypharmacology is 
+well represented in the FTC, as shown by the numerous MoAs each approved drug can exhibit.
+We decided to further look at a well-known repositioning example, 
+in order to see whether the FTC was suitable to identify the new
+uses of an old drug. We picked the \emph{thalidomide} for 
+this exercise ({{https://www.ebi.ac.uk/chembl/ftc/agent/DB01041}}). The molecule was 
+first indicated to treat morning sickness
+in pregnant women, but has been quickly abandoned after its developmental toxicity has been discovered in newborns. 
+The accepted molecular mechanism behind the side-effect is  
+an impairment of the angiogenic process responsible for the development of members, affecting in particular the limbs \citep{therapontos2009thalidomide}.
+We found that the \emph{thalidomide} was accurately classified as \emph{Anti-cell migration involved in sprouting angiogenesis agent} in the FTC, capturing the known toxicity of the drug.
+Furthermore, the \emph{thalidomide} is currently investigated for a multitude of new usages, in particular for anti-cancer and immunomodulatory activities among others \citep{teo2005thalidomide}.
+These new indications are well represented in the FTC too, for 
+example by the categories \emph{Anti-vascular endothelial growth factor production agent} or 
+\emph{Anti-cell division agent} for neoplastic activities, 
+or by the classes \emph{Anti-cytokine secretion agent} and \emph{Anti-I-kappaB kinase/NF-kappaB cascade agent} for its effect
+on the immune system. These observations demonstrate that the FTC can successfully capture the 
+molecular reasons behind the repositioning of an old compound. 
+Moreover the classification can also provide valuable insight regarding
+potential toxicity too.
+
+\begin{figure*}[!tpb]%figure4
+\centerline{\includegraphics{fig4.png}}
+\caption{Pairwise comparison of MoAs similarities. Therapeutic indications are represented by ATC categories which are the colours on the side.
+For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
+which appears as \emph{B} (dark orange) on the plot.
+Only the first ATC level 
+is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
+The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black).
+Some compounds belong to multiple ATC categories (\emph{Multiple}) and some others do not have
+an ATC code (\emph{NoCategory}).
+The average similarity of drugs present in the same therapeutic category 
+is significantly higher on average when separately compared to all other indications.}\label{fig:04}
+\end{figure*}
+
 \subsubsection{Drugs with similar MoAs have similar indications}
@@
 the same FTC category is 1 (maximum). The similarity between an \emph{anti-blood coagulant} and \emph{pro-blood coagulant} is 0.29, 
-reflecting the fact that such compounds are not similar in regards to the outcome of their biological effect. As the MoA is 
+reflecting the fact that such compounds are dissimilar with regards to the outcome of their biological effect. As the MoA is 
 intuitively expected to be the central concept leading to the indication of the drug, we expected that on average, drugs with 
@@
 the overall similarity appears higher when compounds from the same ATC group are considered. 
-A significance analysis (details covered in supplementary section 8)
+A significance analysis (see supplement, section 8)
 revealed that the average MoA similarity of compounds belonging to the same ATC category is 
@@
 the predictions for instance, as well as false positives.
-
 Supplementary Figure 3 present similar association behaviour when two levels of the ATC are 
@@
 functional clusters of drugs. Finally, Supplementary Figure 5 shows the distribution when compounds are sorted based on 
-their identifiers; no patterns are identifiable in this case. Taken together, these results emphasize that the MoAs as 
+their identifiers; no patterns are identifiable in this case. Taken together, these results emphasise that the MoAs as 
 defined in the FTC are indeed on average associated with the therapeutic indication of a drug. This result supports the validity of the
@@
 
-\begin{figure*}[!tpb]%figure4
-\centerline{\includegraphics{fig4.png}}
-\caption{Pairwise comparison of MoAs similarities. Therapeutic indications are represented by ATC categories which are the colors on the side.
-For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
-which appears as \emph{B} (dark orange) on the plot.
-Only the first ATC level 
-is considered. The reader can refer to Supplementary Figure 1A for a two ATC level granularity. 
-The similarity descriptor ranges from 0 (not similar - white) to 1 (identical - black).
-Some compounds belong to multiple ATC categories (\emph{Mulitple}) and some others do not have
-an ATC code (\emph{NoCategory}).
-The average similarity of drugs present in the same therapeutic category 
-is significantly higher on average when separately compared to all other indications.}\label{fig:04}
-\end{figure*}
+\begin{figure}[!tpb]%figure5
+\centerline{\includegraphics{fig5.png}}
+\caption{FTC categories describing some of the modes of 
+action that could impact Alzheimer's disease (letters on figure). The categories have been manually picked on the basis that they could directly affect
+the dementia. Drugs classified in these FTC categories further 
+manually grouped based on their MoAs similarities (numbers on figure).}\label{fig:05}
+\end{figure}
 
@@
 \emph{Anti-tau protein kinase activity agent} (FTC\_A0050321), \emph{Anti-tau protein binding agent} (FTC\_A0048156), 
-\emph{Anti-beta amyloid 
+\emph{Anti beta amyloid 
 binding agent} (FTC\_A0001540) and finally \emph{Pro-beta amyloid binding agent} (FTC\_P0001540). We have then considered the drugs present 
@@
 based on the overall similarity of their actions (numbers on Figure \ref{fig:05}).
-
-\begin{figure}[!tpb]%figure5
-\centerline{\includegraphics{fig5.png}}
-\caption{FTC categories describing some of the modes and mechanisms of 
-action that could impact Alzheimer's disease (letters on figure). The categories have been manually picked on the basis that they could directly affect
-the dementia. Drugs classified in these FTC categories further 
-manually grouped based on their MoAs similarities (numbers on figure).}\label{fig:05}
-\end{figure}
-
 The subgroups (1), (2) and (3) are inhibitors of the cholinergic system and some of them, such as \emph{galantamine} (DB00674), are already 
@@
 indicated for the treatment of cutaneous manifestations in patients with T-cell lymphoma, yet a study has 
-shown in vivo (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
+shown \emph{in vivo} (mouse model) the potential of the drug and other histone deacetylase inhibitors in regards to memory 
 deficit \citep{Kilgore2010}. 
@@
 The FTC is a novel classification for approved drugs, which can be used as a starting point to 
-generate drug repurposing hypotheses. We have guided the reader through some of the ways the resource 
-can be explored. No experimental validations have been presented in this manuscript, yet references to 
-supporting literature was provided in order to legitimate the claims. The main goal of this document is to 
-introduce the resource to the community; deeper analyses are being performed over the classification at the time of writing.
+generate drug repurposing hypotheses. 
 This classification leverages the information present in various databases and ontologies, similarly to the 
@@
 \section{Conclusion}
-This manuscript introduces the FTC, a free, open and extensible resource, which can assist drug repurposing 
-initiatives or enhance computational studies making use of the concept of \emph{mode of action}.
-The resource annotates the biological functions of drugs, in a similar way the GO serves to annotate gene products.
-The construction of 
-the classification relies on description logics as a means to define the mode and mechanism of action of approved 
-drugs. We shown the validity of the approach by comparing the content of the FTC to another well implemented taxonomy, 
-the ATC. We further illustrate the tight relationship between the MoA and the indication of a drug and shown how the 
-resource can serve as a starting point to formulate hypotheses for Alzheimer's disease. The work leverages the 
+The FTC is public resource, which should assist drug repurposing 
+initiatives or enhance computational studies that judge drugs according to their \emph{mode of action}.
+The resource attributes biomolecular functions and processes to drugs, the same way as GO types have been assigned to gene products.
+The construction of FTC relies on axiomatic representations of MoA as the core means to attribute and derive the MoA for approved 
+drugs. We shown the validity of the approach by comparing the content of the FTC to a well established gold standard, 
+the ATC. We further illustrate the tight relationship between the MoA and the indication of a drug and demonstrate using Alzheimer's disease as an example how the 
+resource helps to formulate drug repositioning hypotheses. The work leverages the 
 semantics of distinct databases, working in parallel on different thematics. The platform will be further used 
@@
 \section*{Acknowledgement}
-We would like to thank Rita Santos, Sarah Carl and Felix Kruger for the fruitful input on this manuscript.
+We would like to thank Rita Santos, Sarah Carl and Felix Kruger, as well 
+as the reviewers for the fruitful input on this manuscript.
 

### 2013-10-14 document.tex
correction DRS

--- 2013-10-10 document.tex
@@
 This section describes the building mechanism behind the FTC.
-First, a list of categories describing the mode and mechanism of action of drugs is defined. 
+First, a list of \emph{categories} describing the mode and mechanism of action of drugs is defined (the word \emph{category} is interchangeable with \emph{term}, \emph{class} or \emph{concept} in this manuscript).
 Then in a second step the newly created categories are automatically populated with approved compounds. 
@@
 are only tokens with a human readable label as well as an identifier. The next step is going to 
-assign equivalent definitions to each FTC class. A reasoner can understand such definitions and will automatically 
-classify the knowledge base accordingly; drugs will then be assigned to FTC categories and the taxonomic structure 
+assign equivalent definitions to each FTC class. An OWL reasoner can understand such definitions and will automatically 
+classify the knowledge base accordingly, following standard description logics reasoning services (see wiki for introduction).
+Drugs will then be assigned to FTC categories and the taxonomic structure 
 arises after this reasoning step. Equivalent definitions are written as OWL class expressions using the entities of 
@@
 E.g. \emph{Prothrombin SubClassOf involved-in some positive regulation of blood coagulation}. 
-Each protein can be involved in multiple processes and capable of realizing multiple functions;
-the polypharmacology is captured at this level.
+Each protein can be involved in multiple processes and capable of realising multiple functions;
+some of the polypharmacology is captured at this level.
 
@@
 are shared by A and B (intersection) divided by the number of superclasses of A or B (union). The index ranges 
-from 0 (totally different) to 1 (identical).
+from 0 (totally different) to 1 (identical). A similar approach was successfully implemented by \citep{hoehndorf2011phenomenet}, for similarity computations over phenotypic traits.
 

### 2013-10-15 reply-to-editors.txt
ready for resubmission

@@
 In this regard, we have provided a supplementary reference to an introductory text.
-It is challenging to add more information to the main document, as reasoning services and tools have been developed by other research groups and are simply used in the context of our work.
-We rely on the W3C OWL international standard to interact with reasoning engines, for which extensive documentation is available.
-Moreover, we specially wrote an introductory text to description logics, available on the wiki for biomedical users non familiar with the domain.
+It is challenging to add more information to the main document, as reasoning services and tools have been developed by other research groups and are simply used for standard tasks in the context of our work.
+We rely on the W3C OWL2 international standard to interact with reasoning engines, for which extensive documentation is available.
+Moreover, we specially wrote an introductory text to description logics, available on the wiki, for biomedical users non familiar with the domain.
 Ideally description logics would be presented in greater details in the main manuscript, unfortunately and because of space concerns we decided to focus on the presentation of the biomedical novelty brought by our project.
@@
 The relation between the two resources is described in section 2.1 "Evaluation" and furthermore on the website of the application (https://www.ebi.ac.uk/chembl/ftc/evaluation/), where we compare equivalent classes and derive metrics (precision and recall) summarising the overlap between these resources.
-It is not clear to us what are the specific requirements of reviewer 1 regarding this revision.
+The section 3.2 also discusses the interpretation of the evaluation and the differences/similarity between FTC/ATC.
+It is unfortunately not clear to us what are the specific requirements of reviewer 1 regarding this revision.
 Reviewer 2 asked us to illustrate the potential of the resource with a known repurposing example.
 We chose the thalidomide to fulfil this legitimate requirement, as it is possibly one of the best known repositioning case.
-We have added a paragraph to the manuscript (end of section 2.2.1) explaining how the teratogenic effect of the molecule is described in the FTC and how the new indications of the drug are captured (anti-cancer and immunomodulatory activities, nowadays routinely prescribed to patients).
+We have added a paragraph to the manuscript (end of section 2.2.1) explaining how the transgenic effect of the molecule is described in the FTC and how the new indications of the drug are captured (anti-cancer and immunomodulatory activities, nowadays routinely prescribed to patients).
 We briefly discussed the relationship between our resource and the work done in Bioinformatics 28.16 (2012): 2169-2175 and the Open PHACTS platform.

### 2013-10-15 document.tex
ready for resubmission

--- 2013-10-14 document.tex
@@
 \section{Motivation:}
-Drug repurposing is the discovery of new indications for compounds that have already been approved and used in a clinical setting. 
+Drug repositioning is the discovery of new indications for compounds that have already been approved and used in a clinical setting. 
 Recently, some computational approaches have been suggested to unveil new opportunities in a systematic fashion, by taking 
@@
 \centerline{\includegraphics{fig1.png}}
-\caption{The diagram gives an overview of the integrated resources. (A) The name of FTC categories representing MoAs are derived from the GO terms representing the molecular 
+\caption{The diagram gives an overview of the integrated resources. (A) The name of FTC categories representing MoAs 
+are directly derived from the GO terms representing the molecular 
 functions and biological processes. A 
 prefix \emph{anti} or \emph{pro} is appended to the original 
-GO term, as well as the word \emph{agent} as suffix. (B) Each of the new FTC class has a logical
+GO term, as well as the word \emph{agent} as suffix (e.g, the GO term \emph{blood coagulation} serves as a template to create the
+\emph{Anti-blood coagulation agent} and \emph{Pro-blood coagulation agent} categories). (B) Each of the new FTC class has a logical
 equivalent definition assigned to it (axiom), representing the necessary and sufficient conditions for a drug to be classified
@@
 have in common to look for similarities within existing drugs and forward similar compounds as repurposing hypotheses.
+
 A feature of particular interest to describe drugs is the MoA. According to Wikipedia, 
@@
 well-known and major MoAs of chemical compounds. The pharmacological spectrum of each drug is not necessarily well covered, yet 
-it would be the best way to predict new indications. In our context, an ideal knowledge base would feature the known MoAs of a 
+it would be the best way to predict new indications.
+
+In our context, an ideal knowledge base would feature the known MoAs of a 
 drug as well as some predicted ones to be tested in experiments. The MoA categories should derive and scale over primary molecular 
@@
 in the resource and further populated with approved drugs using the Web Ontology Language (OWL) in combination with a reasoner.
-The population step takes in account the type of pharmacological action, the molecular targets of the drug and their involvement
+The population step takes in account the type of pharmacological action, the molecular targets of the drugs and their involvement
 into biological processes. Drugs can exhibit several MoAs, and the same MoA can be reached through different mechanisms.
@@
 Other OWL reasoners (e.g, Hermit, Pellet, etc.) disqualified mainly due to long processing time (data not shown).
+
 The FTC forms a taxonomic structure as illustrated 
@@
 The full methodology behind the evaluation is described in the section 6 of the supplementary material (\emph{Evaluation methodology}).
+
 Briefly, for 68 categories from the FTC we can manually identify a set of semantically equivalent categories in the ATC.
@@
 evaluation point is furthermore available online at {{https://www.ebi.ac.uk/chembl/ftc/evaluation/FTC\_A0008900}}.
+
 For 1'280 DrugBank compounds in the FTC, 1'134 are also present in the ATC, therefore only those were 
@@
 evaluation point is such a case where all the drugs (\emph{omeprazole, esomeprazole, pantoprazole, lansoprazole, rabeprazole}) are 
-present both in the FTC category and in the corresponding ATC categori(es). The total number of compounds from an ATC category but 
+present both in the FTC category and in the corresponding ATC category. The total number of compounds from an ATC category but 
 where we could not identify a corresponding FTC category is 35 (false negatives). Finally, 280 compounds are present in a FTC class but not in any 
@@
 well represented in the FTC, as shown by the numerous MoAs each approved drug can exhibit.
+
 We decided to further look at a well-known repositioning example, 
@@
 example by the categories \emph{Anti-vascular endothelial growth factor production agent} or 
-\emph{Anti-cell division agent} for neoplastic activities, 
+\emph{Anti-cell division agent} for antineoplastic activities, 
 or by the classes \emph{Anti-cytokine secretion agent} and \emph{Anti-I-kappaB kinase/NF-kappaB cascade agent} for its effect
@@
 \caption{Pairwise comparison of MoAs similarities. Therapeutic indications are represented by ATC categories which are the colours on the side.
-For instance, the compound reteplase (DB00015) has the ATC code B01AD07, 
+For instance, the compound \emph{reteplase} (DB00015) has the ATC code B01AD07, 
 which appears as \emph{B} (dark orange) on the plot.
@@
 intuitively expected to be the central concept leading to the indication of the drug, we expected that on average, drugs with 
-similar MoAs would be indicated towards similar therapeutic areas. The heat map presented in Figure \ref{fig:04} shows a pairwise 
+similar MoAs would be indicated towards similar therapeutic areas.
+
+The heat map presented in Figure \ref{fig:04} shows a pairwise 
 comparison of all the drugs of the FTC based on their relative MoA similarity. The compounds are further grouped by 
@@
 the predictions for instance, as well as false positives.
+
 Supplementary Figure 3 present similar association behaviour when two levels of the ATC are 
